### 2022-01-05 paper.tex
Add paper

% Created 2022-01-04 Tue 17:13
% Intended LaTeX compiler: pdflatex
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\usepackage{indentfirst}
\author{Teddy Groves and Areti Tsigkinopoulou}
\date{\today}
\title{Statistical analysis of kinetic parameter data from online databases}
\hypersetup{
 pdfauthor={Teddy Groves and Areti Tsigkinopoulou},
 pdftitle={Statistical analysis of kinetic parameter data from online databases},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 29.0.50 (Org mode 9.5)}, 
 pdflang={English}}
\usepackage[notquote]{hanging}
\begin{document}\makeatletter
\newcommand{\citeprocitem}[2]{\hyper@linkstart{cite}{citeproc_bib_item_#1}#2\hyper@linkend}
\makeatother

\maketitle
\tableofcontents

\begin{abstract}
We fit a range of Bayesian regression models to kinetic parameter data from the
BRENDA and SABIO-RK databases, report the results and provide tools for
reproducing the results and using them in a systems biology workflow. We
illustrate the intended use of our results with case studies.
\end{abstract}

\section{Introduction}
\label{sec:org10d6a8f}
The choice of appropriate parameters for biological models has been a recurring
issue for computational biologists in the recent years. As computational models
become increasingly prominent in the current research \citeprocitem{1}{[1]}–\citeprocitem{3}{[3]}
and lead the way to biological breakthroughs, the use of high quality parameter
values is becoming essential. In view of these developments, there have been
various efforts to address this issue by employing different
strategies. Although the most popular approach is the estimation of parameters
through optimization methods \citeprocitem{4}{[4]}, \citeprocitem{5}{[5]},
ensemble modelling strategies are rapidly gaining ground in the field of systems
biology \citeprocitem{6}{[6]}–\citeprocitem{9}{[9]}
along with the development of tools to define informative parameter priors \citeprocitem{10}{[10]}, \citeprocitem{11}{[11]}. At
the same time, machine learning methods have recently been gaining popularity as
a tool for interpreting the large amount of existing biological data \citeprocitem{12}{[12]}–\citeprocitem{14}{[14]}
and have also been employed for the prediction of parameter values (mostly Km
values) in biological models \citeprocitem{15}{[15]}–\citeprocitem{17}{[17]}.

By taking into the account the existing issues with model parameterisation and
in an effort to address them, we have previously developed a standardized
protocol for the definition of appropriate parameter lognormal distributions
based on information retrieved from different sources and in accordance with the
modeller’s beliefs about the reliability of each experimental value \citeprocitem{18}{[18]}. This protocol was a
significant step towards making ensemble modelling more accessible and promoting
interdisciplinary collaborations. However, there were some remaining issues to
be addressed. Firstly, the protocol should take into account the surrounding
parameter landscape (phylogenetically related species, enzyme sub-classes etc.)
in order to avoid having the prior distribution being too narrow or focused only
on one particular area. Furthermore, it should avoid over-reliance on a limited
set of experimental reports.

In order to resolve these problems and additionally make the protocol fully
automated, we devised a novel method for generating appropriate prior
distributions for kinetic parameters. Using previous work by Liebermeister and
Klipp as a starting point \citeprocitem{19}{[19]}, we
developed a hierarchical Bayesian regression model in order to perform a
statistical analysis of parameter reports from the BRENDA \citeprocitem{20}{[20]} and SABIO-RK \citeprocitem{21}{[21]} databases. The results of this
analysis can be used in our existing protocol to rationalise parameter weights,
or as part of another protocol.  In order to make it easier to access and review
the information in BRENDA, we also provided an online interface through which
our model’s results can easily be reviewed and extracted.

\section{State of the art}
\label{sec:orgdd26fbe}
The problem of modelling kinetic parameter data from online databases has
previously been addressed in several studies.

\citeprocitem{19}{[19]} models logarithmic-scale measurements
\(y_{ijk}\) of a Km value for substrate \(i\), ec number \(j\) and organism \(k\) using
the following ANOVA-style regression model:

\begin{equation}
y_ijk \sim N(\mu_i + \alpha_ij + \beta_ik, \sigma)
\end{equation}

In this expression the term \(\mu\) represents substrate-specific effects, whereas
the terms \(\alpha\) and \(\beta\) respectively represent substrate-ec number and
substrate-organism interaction effects. \(\sigma\) represents the accuracy of the
measurement apparatus.

The authors fit this measurement model to a subset of BRENDA data using maximum
likelihood estimation, obtain out of sample predictions using leave-one-out
cross validation and investigate patterns in the results.

This approach is limited by inability to incorporate prior information about the
plausible values of the main effects or their general trends. In addition, the
analysis produces point estimates: it does not attempt to fully capture the
available information about Kms. This issue is particularly pertinent for
systems biologists who need to construct informative prior distributions for
kinetic models. For these users it is more important to know which possible Km
values are ruled out by the data in BRENDA than it is to know which make that
data most likely. Below we show that both of these issues with the approach in
[19] can be addressed by incorporating a hierarchical Bayesian component.

\citeprocitem{16}{[16]} and \citeprocitem{17}{[17]} predict
kinetic parameters from BRENDA and other sources by taking into account protein
and substrate structure information using neural networks. Exploiting this
information allows for improved out of sample predictive performance compared to
models that do not use it.

However, this approach does not avoid the issues with lack of prior information
and inaptness for downstream prior modelling that we highlight above. In
addition, the need for information about protein structure limits the amount of
kinetic parameters for which predictions can be obtained. Again taking the point
of view of a systems biologist attempting to construct informative prior
distributions, this is a severe problem as coverage is at least as important a
consideration for this application as precision.

\section{Method}
\label{sec:org4bebd6f}
\subsection{Data fetching}
\label{sec:orgba54ca3}
We fetched data from the SABIO-RK \citeprocitem{21}{[21]}
and BRENDA \citeprocitem{22}{[22]} databases using their publicly
available APIs. We also fetched a list of all EC numbers from the Expasy
database \citeprocitem{23}{[23]}. In the project repository,
see the script \texttt{fetch\_data.py} and the library file \texttt{fetching.py} for code used
to fetch data and the directory \texttt{data/raw/} for the results.

\subsection{Data processing}
\label{sec:orgdfed31c}
We made several significant data processing choices. See the library file
\texttt{data\_preparation.py} for code used to implement these choices.

\subsubsection{Filtering}
\label{sec:org8326023}
For each dataset and kinetic parameter, we removed all reports which failed to
satisfy any of the following conditions:

\begin{itemize}
\item The kinetic parameter value must be a number
\item The literature reference must not be missing
\item The substrate must be catalysed naturally by the enzyme
\item The enzyme must be from a wild organism
\item The temperature, if recorded, must be between 5 and 50 degrees C
\item The pH, if recorded, must be between 4 and 9
\item The organism must be have data from at least 50 separate study/biology
(i.e. organism:substrate:ec4 for BRENDA or organism:substrate:enzyme for
SABIO-RK) combinations that satisfy all the other conditions.
\end{itemize}

\subsubsection{Grouping}
\label{sec:orgc5b09e3}
Instead of modelling reports directly, we chose to group together reports with
the same biology and study, treating the median log-scale km as a single
observation. We took this decision because of the presence in both datasets of
different kinds of study. In some cases - presumably when the aim of a study was
to discover the sensitivity of a kinetic parameter to changes in conditions -
many reports with the same enzyme, organism, substrate and study are available,
with a range of different kinetic parameter values and different experimental
conditions recorded in the \texttt{commentary} field. In other cases a study will
report only a single value for one kinetic parameter.

Due to this discrepancy it seemed wrong to treat reports from better populated
studies as equivalent to reports from more concise studies. While taking the
median for a given study/biology combination before modelling destroys
information, we judged that it would lead to more realistic results than
treating each report as an observation, especially since we chose not to
attempt to model the effects of experimental conditions.

\subsection{Statistical model}
\label{sec:org9e76746}
We used a Bayesian hierarchical regression model to describe all data. Since
both Km and Kcat parameters are constrained to be positive, we modelled them on
natural logarithmic scale, using the same approach taken in \citeprocitem{19}{[19]}. Our model includes a global mean
parameter \(\mu\), hierarchical substrate-specific intercept parameters \(a^{sub}\)
and hierarchical intercept parameters \(a^{enz:sub}\), \(a^{ec4:sub}\) and
\(a^{org:sub}\) specific to interactions of substrate and enzyme, ec4 number and
organism respectively. In addition we used a student-T measurement model with
latent standard deviation and degrees of freedom \(\sigma\) and \(\nu\).

We chose assigned semi-informative prior distributions based on the
pre-experimental information. In particular, the prior for the measurement
distribution degrees of freedom parameter \(\nu\) follows the recommendation in \citeprocitem{24}{[24]} and is truncated below at
zero, reflecting our view that the measurement distribution should not be
excessively heavy-tailed.

The full model specification in tilde notation is as follows:

\begin{align*}
\ln y &\sim ST(\nu,
               \mu + a^{sub} + a^{enz:sub} + a^{ec4:sub} + a^{org:sub},
               \sigma) \\
\nu &\sim \Gamma(2, 0.1)[1, \infty] \\
\sigma &\sim N(0, 2)[0, \infty] \\
\mu &\sim N(-2, 1) \\
a^{sub} &\sim N(0, \tau^{sub}) \\
a^{enz:sub} &\sim N(0, \tau^{enz:sub}) \\
a^{ec4:sub} &\sim N(0, \tau^{ec4:sub}) \\
a^{org:sub} &\sim N(0, \tau^{org:sub}) \\
\tau^{sub} &\sim N(0, 1) \\
\tau^{enz:sub} &\sim N(0, 1) \\
\tau^{ec4:sub} &\sim N(0, 1) \\
\tau^{org:sub} &\sim N(0, 1) \\
\end{align*}

This model can be expressed as the following Stan program (see \citeprocitem{25}{[25]}):

\begin{verbatim}
/* Extends the BLK model for more specific data from the SABIO-rk database */

data {
  int<lower=1> N_train;
  int<lower=1> N_test;
  int<lower=1> N_biology;
  int<lower=1> N_substrate;
  int<lower=1> N_ec4_sub;
  int<lower=1> N_enz_sub;
  int<lower=1> N_org_sub;
  int<lower=1,upper=N_ec4_sub> ec4_sub[N_biology];
  int<lower=1,upper=N_enz_sub> enz_sub[N_biology];
  int<lower=1,upper=N_org_sub> org_sub[N_biology];
  int<lower=1,upper=N_substrate> substrate[N_biology];
  array[N_train] int<lower=1,upper=N_biology> biology_train;
  vector[N_train] y_train;
  array[N_test] int<lower=1,upper=N_biology> biology_test;
  vector[N_test] y_test;
  int<lower=0,upper=1> likelihood;
}
parameters {
  real<lower=1> nu;
  real mu;
  real<lower=0> sigma;
  real<lower=0> tau_substrate;
  real<lower=0> tau_ec4_sub;
  real<lower=0> tau_enz_sub;
  real<lower=0> tau_org_sub;
  vector<multiplier=tau_substrate>[N_substrate] a_substrate;
  vector<multiplier=tau_ec4_sub>[N_ec4_sub] a_ec4_sub;
  vector<multiplier=tau_enz_sub>[N_enz_sub] a_enz_sub;
  vector<multiplier=tau_org_sub>[N_org_sub] a_org_sub;
}
transformed parameters {
  vector[N_biology] log_km =
    mu
    + a_substrate[substrate]
    + a_ec4_sub[ec4_sub]
    + a_enz_sub[enz_sub]
    + a_org_sub[org_sub];
}
model {
  if (likelihood){y_train ~ student_t(nu, log_km[biology_train], sigma);}
  nu ~ gamma(2, 0.1);
  sigma ~ normal(0, 2);
  mu ~ normal(-2, 1);
  a_substrate ~ normal(0, tau_substrate);
  a_ec4_sub ~ normal(0, tau_ec4_sub);
  a_enz_sub ~ normal(0, tau_enz_sub);
  a_org_sub ~ normal(0, tau_org_sub);
  tau_org_sub ~ normal(0, 1);
  tau_ec4_sub ~ normal(0, 1);
  tau_enz_sub ~ normal(0, 1);
  tau_substrate ~ normal(0, 1);
}
generated quantities {
  vector[N_test] llik;
  vector[N_test] yrep;
  for (n in 1:N_test){
    llik[n] = student_t_lpdf(y_test[n] | nu, log_km[biology_test[n]], sigma);
    yrep[n] = student_t_rng(nu, log_km[biology_test[n]], sigma);
  }
}
\end{verbatim}

For the BRENDA data, we used the same model but removed the enzyme-substrate
interaction parameters as BRENDA does not provide enzyme-specific information.

\subsection{Model validation}
\label{sec:org903f537}
We used a number of standard methods to test our models' validity, aiming to
follow the method described in \citeprocitem{26}{[26]}. To verify the
computation we used standard MCMC convergence metrics and fit our model to fake
data. To assess model specification we performed graphical prior and posterior
predictive checks and both approximate and exact cross validation, using some
simple test models for comparison.

\subsubsection{Computation}
\label{sec:orge25051a}
For each MCMC run we used arviz \citeprocitem{27}{[27]} to
calculate the improved \hat{R} statistic following the method in \citeprocitem{28}{[28]} and to check for
divergent transitions. If the \hat{R} statistic was sufficiently close to 1
(i.e. +/- 0.03) and there were no post-warmup divergent transitions we judged
that the computation was likely to have been successful in the sense that the
draws can be treated as samples from the target distribution.

To further validate the computation we also fit the model to fake data generated
using the model assumptions and a plausible configuration of parameters. We used
graphical posterior predictive checks to assess whether the true parameters were
approximately recovered in the posterior distribution. See supplementary
material for graphical posterior predictive checks with fake data.

\subsubsection{Comparison models}
\label{sec:org091a4ae}
Ideally we would compare the results of our models with previously published
attempts to model the same data. However this was not practical in our case for
two reasons. First, other results are typically derived from different raw
datasets, and rely on different data filtering and summarising decisions. In
particular, we have not previously seen the problem of whether and how to
aggregate results from the same study addressed in detail. Second, the other
results typically assessed model specification by comparing observed data to
point predictions generated by their models. Applying this procedure to our
models would give an incomplete picture, since a single point cannot adequately
summarise the full posterior predictive distribution. In addition, it would not
align with the priorities of our intended users, namely systems biologists
constructing prior distributions for the purpose of ensemble modelling, who we
imagine care more about the extremes of the distributions than central
estimates.

Due to this lack, we constructed two simple models purely for comparison with
our main model. The first model, which we called the "really simple" model,
removes all latent random variables from the main model's predictor, except for
the global mean parameter \(\mu\). The second model, which we called the "simple"
model, includes a single vector of hierarchical parameters at the finest
available granularity, i.e. organism:substrate:ec4 for BRENDA data and
organism:substrate:enzyme for SABIO-RK data. See supplemental information for
full model specifications in the form of tilde notation and Stan programs.

Our thinking was that the really simple model and simple models represent
relative extremes of underfitting and overfitting. A well-specified model should
be able to make better predictions than both, unless the data are very
surprisingly uninformative.

\subsubsection{Graphical posterior predictive checks}
\label{sec:org9b7eca4}
Figure \ref{fig:org829db47} below shows our main model's marginal posterior
predictive 1\%-99\% interval for each observation, alongside the observed
value. Ideally exactly 98\% of observations should be covered, and whether or not
an observation is covered should not be systematically predictable.

\begin{figure}[htbp]
\centering
\includegraphics[width=.9\linewidth]{./results/plots/ppc_km_sabio_enz.png}
\caption{\label{fig:org829db47}Marginal posterior predictive distributions for main model fit to SABIO Km data}
\end{figure}

In contrast figure \ref{fig:org96eb4db} are graphical posterior predictive checks for the really simple
and simple models.

\begin{figure}[htbp]
\centering
\includegraphics[width=.9\linewidth]{./results/plots/ppc_km_sabio_enz_simple.png}
\caption{\label{fig:org96eb4db}Marginal posterior predictive distributions for simple and really simple models fit to SABIO Km data}
\end{figure}

\subsubsection{Cross validation}
\label{sec:org7c3b35d}
For a quantitative compliment to the graphical posterior predictive checks, we
used cross-validation to assess our models' out-of-sample predictive
performance. For exploratory comparison we calculated each model's approximate
leave-one-out expected log predictive density using the method set out in \citeprocitem{29}{[29]} and implemented in \citeprocitem{27}{[27]}. To address cases where diagnostics
suggested that the approximate leave-one-out algorithm was likely to be
unreliable we also carried out exact tenfold cross-validation.

The results were as follows:

\subsection{Web app}
\label{sec:orge6374d2}
\section{Results}
\label{sec:org21ac059}
\subsection{Marginal distributions of interesting parameters}
\label{sec:org8bc318e}
\subsection{Comparison of estimated Kms with physiological metabolite concentrations}
\label{sec:orgbc6eea3}
\begin{figure}[htbp]
\centering
\includegraphics[width=.9\linewidth]{./results/plots/hmdb_comparison.png}
\caption{\label{fig:org83a902f}Comparison of model predictions with physiological metabolite concentrations from HMDB}
\end{figure}

\subsection{NADH vs NADPH}
\label{sec:org1fdc6c3}

\begin{figure}[htbp]
\centering
\includegraphics[width=.9\linewidth]{./results/plots/nadh_comparison.png}
\caption{\label{fig:org3ae5622}Comparison of modelled and measured Km parameters for NADH and NADPH}
\end{figure}

\subsection{BRENDA vs SABIO-RK}
\label{sec:org5f86a62}

\section{Case studies}
\label{sec:orgd3eb7c0}
\section{Discussion}
\label{sec:org599ba7b}
\section{References}
\label{sec:orgf7f9da0}

\hypertarget{citeproc_bib_item_1}{[1] A. J. Lopatkin and J. J. Collins, “Predictive Biology: Modelling, Understanding and Harnessing Microbial Complexity,” \textit{Nature Reviews. Microbiology}, vol. 18, no. 9, pp. 507–520, Sep. 2020, doi: \href{https://doi.org/10.1038/s41579-020-0372-5}{10.1038/s41579-020-0372-5}.}

\hypertarget{citeproc_bib_item_2}{[2] S. Khoshnaw, R. Salih, and S. Sulaimany, “Mathematical Modelling for Coronavirus Disease (COVID-19) in Predicting Future Behaviours and Sensitivity Analysis,” \textit{Mathematical Modelling of Natural Phenomena}, vol. 15, May 2020, doi: \href{https://doi.org/10.1051/mmnp/2020020}{10.1051/mmnp/2020020}.}

\hypertarget{citeproc_bib_item_3}{[3] M. Tokic, V. Hatzimanikatis, and L. Miskovic, “Large-Scale Kinetic Metabolic Models of Pseudomonas Putida KT2440 for Consistent Design of Metabolic Engineering Strategies,” \textit{Biotechnology for Biofuels}, vol. 13, p. 33, 2020, doi: \href{https://doi.org/10.1186/s13068-020-1665-7}{10.1186/s13068-020-1665-7}.}

\hypertarget{citeproc_bib_item_4}{[4] A. Khodayari and C. D. Maranas, “A Genome-Scale Escherichia Coli Kinetic Metabolic Model k-Ecoli457 Satisfying Flux Data for Multiple Mutant Strains,” \textit{Nature Communications}, vol. 7, no. 1, Dec. 2016, doi: \href{https://doi.org/10.1038/ncomms13806}{10.1038/ncomms13806}.}

\hypertarget{citeproc_bib_item_5}{[5] P. A. Saa and L. K. Nielsen, “Formulation, Construction and Analysis of Kinetic Models of Metabolism: A Review of Modelling Frameworks,” \textit{Biotechnology Advances}, vol. 35, no. 8, pp. 981–1003, Dec. 2017, doi: \href{https://doi.org/10.1016/j.biotechadv.2017.09.005}{10.1016/j.biotechadv.2017.09.005}.}

\hypertarget{citeproc_bib_item_6}{[6] A. Tsigkinopoulou, S. M. Baker, and R. Breitling, “Respectful Modeling: Addressing Uncertainty in Dynamic System Models for Molecular Biology,” \textit{Trends in Biotechnology}, vol. 35, no. 6, pp. 518–529, Jun. 2017, doi: \href{https://doi.org/10.1016/j.tibtech.2016.12.008}{10.1016/j.tibtech.2016.12.008}.}

\hypertarget{citeproc_bib_item_7}{[7] F. Hussain, C. J. Langmead, Q. Mi, J. Dutta-Moscato, Y. Vodovotz, and S. K. Jha, “Automated Parameter Estimation for Biological Models Using Bayesian Statistical Model Checking,” \textit{BMC Bioinformatics}, vol. 16, no. 17, p. S8, Dec. 2015, doi: \href{https://doi.org/10.1186/1471-2105-16-S17-S8}{10.1186/1471-2105-16-S17-S8}.}

\hypertarget{citeproc_bib_item_8}{[8] X. Zhu, M. Welling, F. Jin, and J. Lowengrub, “Predicting Simulation Parameters of Biological Systems Using a Gaussian Process Model,” \textit{Statistical Analysis and Data Mining: The ASA Data Science Journal}, vol. 5, no. 6, pp. 509–522, 2012, doi: \href{https://doi.org/10.1002/sam.11163}{10.1002/sam.11163}.}

\hypertarget{citeproc_bib_item_9}{[9] S. van Mourik, C. ter Braak, H. Stigter, and J. Molenaar, “Prediction Uncertainty Assessment of a Systems Biology Model Requires a Sample of the Full Probability Distribution of Its Parameters,” \textit{PeerJ}, vol. 2, p. e433, Jun. 2014, doi: \href{https://doi.org/10.7717/peerj.433}{10.7717/peerj.433}.}

\hypertarget{citeproc_bib_item_10}{[10] P. Saa and L. K. Nielsen, “A General Framework for Thermodynamically Consistent Parameterization and Efficient Sampling of Enzymatic Reactions,” \textit{PLOS Computational Biology}, vol. 11, no. 4, p. e1004195, Apr. 2015, doi: \href{https://doi.org/10.1371/journal.pcbi.1004195}{10.1371/journal.pcbi.1004195}.}

\hypertarget{citeproc_bib_item_11}{[11] S. Mukherjee and T. P. Speed, “Network Inference Using Informative Priors,” \textit{Proceedings of the National Academy of Sciences}, vol. 105, no. 38, pp. 14313–14318, Sep. 2008, doi: \href{https://doi.org/10.1073/pnas.0802272105}{10.1073/pnas.0802272105}.}

\hypertarget{citeproc_bib_item_12}{[12] R. E. Baker, J.-M. Peña, J. Jayamohan, and A. Jérusalem, “Mechanistic Models versus Machine Learning, a Fight Worth Fighting for the Biological Community?,” \textit{Biology Letters}, vol. 14, no. 5, p. 20170660, May 2018, doi: \href{https://doi.org/10.1098/rsbl.2017.0660}{10.1098/rsbl.2017.0660}.}

\hypertarget{citeproc_bib_item_13}{[13] C. Xu and S. A. Jackson, “Machine Learning and Complex Biological Data,” \textit{Genome Biology}, vol. 20, no. 1, p. 76, Apr. 2019, doi: \href{https://doi.org/10.1186/s13059-019-1689-0}{10.1186/s13059-019-1689-0}.}

\hypertarget{citeproc_bib_item_14}{[14] M. Zitnik, F. Nguyen, B. Wang, J. Leskovec, A. Goldenberg, and M. M. Hoffman, “Machine Learning for Integrating Data in Biology and Medicine: Principles, Practice, and Opportunities,” \textit{An International Journal on Information Fusion}, vol. 50, pp. 71–91, Oct. 2019, doi: \href{https://doi.org/10.1016/j.inffus.2018.09.012}{10.1016/j.inffus.2018.09.012}.}

\hypertarget{citeproc_bib_item_15}{[15] S.-M. Yan, D.-Q. Shi, H. Nong, and G. Wu, “Predicting Km Values of Beta-Glucosidases Using Cellobiose as Substrate,” \textit{Interdisciplinary Sciences, Computational Life Sciences}, vol. 4, no. 1, pp. 46–53, Mar. 2012, doi: \href{https://doi.org/10.1007/s12539-012-0115-z}{10.1007/s12539-012-0115-z}.}

\hypertarget{citeproc_bib_item_16}{[16] A. Kroll, M. K. M. Engqvist, D. Heckmann, and M. J. Lercher, “Deep Learning Allows Genome-Scale Prediction of Michaelis Constants from Structural Features,” \textit{PLoS biology}, vol. 19, no. 10, p. e3001402, Oct. 2021, doi: \href{https://doi.org/10.1371/journal.pbio.3001402}{10.1371/journal.pbio.3001402}.}

\hypertarget{citeproc_bib_item_17}{[17] F. Li \textit{et al.}, “Deep Learning Based Kcat Prediction Enables Improved Enzyme Constrained Model Reconstruction.” Cold Spring Harbor Laboratory, pp. 2021.08.06.455417, Aug. 2021. doi: \href{https://doi.org/10.1101/2021.08.06.455417}{10.1101/2021.08.06.455417}.}

\hypertarget{citeproc_bib_item_18}{[18] A. Tsigkinopoulou, A. Hawari, M. Uttley, and R. Breitling, “Defining Informative Priors for Ensemble Modeling in Systems Biology,” \textit{Nature Protocols}, vol. 13, no. 11, pp. 2643–2663, Nov. 2018, doi: \href{https://doi.org/10.1038/s41596-018-0056-z}{10.1038/s41596-018-0056-z}.}

\hypertarget{citeproc_bib_item_19}{[19] S. Borger, W. Liebermeister, and E. Klipp, “Prediction of Enzyme Kinetic Parameters Based on Statistical Learning,” \textit{Genome Informatics}, vol. 17, no. 1, pp. 80–87, 2006.}

\hypertarget{citeproc_bib_item_20}{[20] “BRENDA - SOAP Access Help.” https://www.brenda-enzymes.org/soap.php\\#Turnover\\_Number.}

\hypertarget{citeproc_bib_item_21}{[21] U. Wittig \textit{et al.}, “SABIO-RK\textemdash Database for Biochemical Reaction Kinetics,” \textit{Nucleic Acids Research}, vol. 40, no. D1, pp. D790–D796, Jan. 2012, doi: \href{https://doi.org/10.1093/nar/gkr1046}{10.1093/nar/gkr1046}.}

\hypertarget{citeproc_bib_item_22}{[22] A. Chang \textit{et al.}, “BRENDA, the ELIXIR Core Data Resource in 2021: New Developments and Updates,” \textit{Nucleic Acids Research}, vol. 49, no. D1, pp. D498–D508, Jan. 2021, doi: \href{https://doi.org/10.1093/nar/gkaa1025}{10.1093/nar/gkaa1025}.}

\hypertarget{citeproc_bib_item_23}{[23] S. Duvaud, C. Gabella, F. Lisacek, H. Stockinger, V. Ioannidis, and C. Durinx, “Expasy, the Swiss Bioinformatics Resource Portal, as Designed by Its Users,” \textit{Nucleic Acids Research}, vol. 49, no. W1, pp. W216–W227, Jul. 2021, doi: \href{https://doi.org/10.1093/nar/gkab225}{10.1093/nar/gkab225}.}

\hypertarget{citeproc_bib_item_24}{[24] M. A. Juárez and M. F. J. Steel, “Model-Based Clustering of Non-Gaussian Panel Data Based on Skew-t Distributions,” \textit{Journal of Business \& Economic Statistics}, vol. 28, no. 1, pp. 52–66, Jan. 2010, doi: \href{https://doi.org/10.1198/jbes.2009.07145}{10.1198/jbes.2009.07145}.}

\hypertarget{citeproc_bib_item_25}{[25] B. Carpenter \textit{et al.}, “Stan: A Probabilistic Programming Language,” \textit{Journal of Statistical Software}, vol. 76, no. 1, pp. 1–32, Jan. 2017, doi: \href{https://doi.org/10.18637/jss.v076.i01}{10.18637/jss.v076.i01}.}

\hypertarget{citeproc_bib_item_26}{[26] A. Gelman \textit{et al.}, “Bayesian Workflow,” \textit{arXiv:2011.01808 [stat]}, Nov. 2020,Available: \href{https://arxiv.org/abs/2011.01808}{https://arxiv.org/abs/2011.01808}}

\hypertarget{citeproc_bib_item_27}{[27] R. Kumar, C. Carroll, A. Hartikainen, and O. Martin, “ArviZ a Unified Library for Exploratory Analysis of Bayesian Models in Python,” \textit{Journal of Open Source Software}, vol. 4, no. 33, p. 1143, Jan. 2019, doi: \href{https://doi.org/10.21105/joss.01143}{10.21105/joss.01143}.}

\hypertarget{citeproc_bib_item_28}{[28] A. Vehtari, A. Gelman, D. Simpson, B. Carpenter, and P.-C. Bürkner, “Rank-Normalization, Folding, and Localization: An Improved R\textasciicircum for Assessing Convergence of MCMC (with Discussion),” \textit{Bayesian Analysis}, vol. 16, no. 2, pp. 667–718, Jun. 2021, doi: \href{https://doi.org/10.1214/20-BA1221}{10.1214/20-BA1221}.}

\hypertarget{citeproc_bib_item_29}{[29] A. Vehtari, A. Gelman, and J. Gabry, “Practical Bayesian Model Evaluation Using Leave-One-out Cross-Validation and WAIC,” \textit{Statistics and Computing}, vol. 27, no. 5, pp. 1413–1432, Sep. 2017, doi: \href{https://doi.org/10.1007/s11222-016-9696-4}{10.1007/s11222-016-9696-4}.}
\end{document}

### 2022-01-06 generate_results.py
Fixed cross validation directory-not-existing bug

@@ -48,7 +48,9 @@ def main():
                 print(f"\n***Writing inference data to {idata_file}***\n")
                 idata.to_netcdf(idata_file)
             if mc.run_cross_validation:
-                splits_dir = os.path.join(mc.data_dir, "splits")
+                cv_dir = os.path.join(run_dir, "splits")
+                if not os.path.exists(cv_dir):
+                    os.mkdir(cv_dir)
                 if mc.sample_kwargs_cross_validation is None:
                     sample_kwargs = mc.sample_kwargs
                 else:
@@ -57,11 +59,11 @@ def main():
                         **mc.sample_kwargs_cross_validation,
                     }
 
-                for f in sorted(os.listdir(splits_dir)):
-                    output_dir = os.path.join(
-                        run_dir, "splits", f.split(".")[0]
-                    )
-                    input_json = os.path.join(splits_dir, f)
+                for f in sorted(os.listdir(cv_dir)):
+                    output_dir = os.path.join(cv_dir, f.split(".")[0])
+                    if not os.path.exists(output_dir):
+                        os.mkdir(output_dir)
+                    input_json = os.path.join(cv_dir, f)
                     idata = sample(
                         stan_file=mc.stan_file,
                         input_json=input_json,

### 2022-01-06 generate_results.py
Simplify output files

@@ -3,7 +3,10 @@
 import json
 import os
 
+import arviz as az
 import toml
+import xarray
+from xarray.core.dataset import Dataset
 
 from src.model_configuration import ModelConfiguration
 from src.sampling import sample
@@ -33,9 +36,6 @@ def main():
                 input_json = os.path.join(
                     mc.data_dir, f"stan_input_{mode}.json"
                 )
-                output_dir = os.path.join(run_dir, mode)
-                if not os.path.exists(output_dir):
-                    os.mkdir(output_dir)
                 print(f"\n***Fitting model {mc.name} in {mode} mode...***\n")
                 idata = sample(
                     stan_file=mc.stan_file,
@@ -44,13 +44,10 @@ def main():
                     dims=dims,
                     sample_kwargs=mc.sample_kwargs,
                 )
-                idata_file = os.path.join(output_dir, f"idata.nc")
+                idata_file = os.path.join(run_dir, f"{mode}.nc")
                 print(f"\n***Writing inference data to {idata_file}***\n")
                 idata.to_netcdf(idata_file)
             if mc.run_cross_validation:
-                cv_dir = os.path.join(run_dir, "splits")
-                if not os.path.exists(cv_dir):
-                    os.mkdir(cv_dir)
                 if mc.sample_kwargs_cross_validation is None:
                     sample_kwargs = mc.sample_kwargs
                 else:
@@ -58,22 +55,23 @@ def main():
                         **mc.sample_kwargs,
                         **mc.sample_kwargs_cross_validation,
                     }
-
-                for f in sorted(os.listdir(cv_dir)):
-                    output_dir = os.path.join(cv_dir, f.split(".")[0])
-                    if not os.path.exists(output_dir):
-                        os.mkdir(output_dir)
-                    input_json = os.path.join(cv_dir, f)
-                    idata = sample(
+                lliks = []
+                cv_input_dir = os.path.join(mc.data_dir, "stan_inputs_cv")
+                for f in sorted(os.listdir(cv_input_dir)):
+                    input_json_file = os.path.join(cv_input_dir, f)
+                    llik = sample(
                         stan_file=mc.stan_file,
-                        input_json=input_json,
+                        input_json=input_json_file,
                         coords=coords,
                         dims=dims,
                         sample_kwargs=sample_kwargs,
-                    )
-                    idata_file = os.path.join(output_dir, f"idata.nc")
[…]

### 2022-01-06 prepare_data.py

@@ -75,10 +75,10 @@ def generate_prepared_data():
     prepare_data_outputs = [brenda_km_data, sabio_km_data]
     for po in prepare_data_outputs:
         output_dir = os.path.join(PREPARED_DIR, po.name)
-        splits_dir = os.path.join(output_dir, "splits")
+        cv_dir = os.path.join(output_dir, "stan_inputs_cv")
         if not os.path.exists(output_dir):
             os.mkdir(output_dir)
-            os.mkdir(splits_dir)
+            os.mkdir(cv_dir)
         input_file_prior, input_file_posterior = (
             os.path.join(output_dir, f"stan_input_{s}.json")
             for s in ["prior", "posterior"]
@@ -88,14 +88,12 @@ def generate_prepared_data():
         write_stan_json(input_file_posterior, po.standict_posterior)
         write_stan_json(input_file_prior, po.standict_prior)
         for i, si in enumerate(po.standicts_cv):
-            f = os.path.join(splits_dir, f"split_{str(i)}.json")
+            f = os.path.join(cv_dir, f"split_{str(i)}.json")
             write_stan_json(f, si)
         with open(os.path.join(output_dir, "coords.json"), "w") as f:
             json.dump(po.coords, f)
         with open(os.path.join(output_dir, "dims.json"), "w") as f:
             json.dump(po.dims, f)
-        with open(os.path.join(output_dir, "split_ixs.json"), "w") as f:
-            json.dump([(a.tolist(), b.tolist()) for a, b in po.splits], f)
     return prepare_data_outputs
 
 
@@ -133,8 +131,7 @@ def generate_fake_data(pos: List[PrepareDataOutput]):
             print(f"min {param}:", min(params[param]))
             print(f"max {param}:", max(params[param]))
         input_fake = input_orig.copy()
-        input_fake["y_train"] = generated_y
-        input_fake["y_test"] = generated_y
+        input_fake["y"] = generated_y
         write_stan_json(fake_input_file, input_fake)
         write_stan_json(fake_param_file, params)
 

### 2022-01-06 src/data_preparation.py

@@ -3,7 +3,7 @@
 import os
 from dataclasses import dataclass, field
 from functools import partial
-from typing import Any, Dict, Iterable, List, Tuple, Union
+from typing import Any, Dict, List, Union
 
 import numpy as np
 import pandas as pd
@@ -32,6 +32,8 @@ DIMS = {
     "a_enz_sub": ["enz_sub"],
     "a_org_sub": ["org_sub"],
     "log_km": ["biology"],
+    "llik": ["ix_test"],
+    "yrep": ["ix_test"],
 }
 
 THIS_FILE = os.path.dirname(os.path.abspath(__file__))
@@ -52,13 +54,16 @@ class PrepareDataOutput:
     standict_prior: StanDict = field(init=False)
     standict_posterior: StanDict = field(init=False)
     standicts_cv: List[StanDict] = field(init=False)
-    splits: Iterable[Tuple] = field(init=False)
 
     def __post_init__(self):
-        ix_all = range(len(self.lits))
-        self.splits = list(
-            KFold(self.number_of_cv_folds, shuffle=True).split(self.lits)
-        )
+        ix_all = list(range(len(self.lits)))
+        splits = []
+        for train, test in KFold(self.number_of_cv_folds, shuffle=True).split(
+            self.lits
+        ):
+            assert isinstance(train, np.ndarray)
+            assert isinstance(test, np.ndarray)
+            splits.append([list(train), list(test)])
         get_standict_xv = partial(
             get_standict, lits=self.lits, coords=self.coords, likelihood=True
         )
@@ -73,10 +78,10 @@ class PrepareDataOutput:
             get_standict_main(likelihood=likelihood)
             for likelihood in (False, True)
         )
-        self.standicts_cv = list(
+        self.standicts_cv = [
             get_standict_xv(train_ix=train_ix, test_ix=test_ix)
-            for train_ix, test_ix in self.splits
-        )
+            for train_ix, test_ix in splits
+        ]
 
 
 def replace_nulls_with_empty_set(s: pd.Series) -> pd.Series:
@@ -245,10 +250,12 @@ def get_standict(
             "org_sub": lits.groupby("biology")["org_sub_stan"].first(),
             "N_train": len(train_ix),
             "N_test": len(test_ix),
+            "N": len(list(set(train_ix + test_ix))),
             "biology_train": lits.loc[train_ix, "biology_stan"],
             "biology_test": lits.loc[test_ix, "biology_stan"],
[…]

### 2022-01-06 src/sampling.py

@@ -23,6 +23,8 @@ def sample(
     model = CmdStanModel(stan_file=stan_file)
     with open(input_json, "r") as f:
         stan_input = json.load(f)
+    coords["ix_train"] = [i - 1 for i in stan_input["ix_train"]]
+    coords["ix_test"] = [i - 1 for i in stan_input["ix_test"]]
     mcmc = model.sample(data=stan_input, **sample_kwargs)
     return az.from_cmdstanpy(
         posterior=mcmc,

### 2022-01-14 paper.tex
Update paper and analysis

@@
-% Created 2022-01-04 Tue 17:13
+% Created 2022-01-14 Fri 16:43
 % Intended LaTeX compiler: pdflatex
@@
 \section{Introduction}
-\label{sec:org10d6a8f}
+\label{sec:org48c3960}
 The choice of appropriate parameters for biological models has been a recurring
@@
 \section{State of the art}
-\label{sec:orgdd26fbe}
+\label{sec:org1c4397a}
 The problem of modelling kinetic parameter data from online databases has
@@
 \section{Method}
-\label{sec:org4bebd6f}
+\label{sec:orgfced3ba}
 \subsection{Data fetching}
-\label{sec:orgba54ca3}
+\label{sec:orga8ffce1}
 We fetched data from the SABIO-RK \citeprocitem{21}{[21]}
@@
 \subsection{Data processing}
-\label{sec:orgdfed31c}
+\label{sec:org0fe8979}
 We made several significant data processing choices. See the library file
@@
 \subsubsection{Filtering}
-\label{sec:org8326023}
+\label{sec:org6cbac9f}
 For each dataset and kinetic parameter, we removed all reports which failed to
@@
 \subsubsection{Grouping}
-\label{sec:orgc5b09e3}
+\label{sec:orgd279533}
 Instead of modelling reports directly, we chose to group together reports with
@@
 \subsection{Statistical model}
-\label{sec:org9e76746}
+\label{sec:org2076f31}
 We used a Bayesian hierarchical regression model to describe all data. Since
@@
 \subsection{Model validation}
-\label{sec:org903f537}
+\label{sec:org6c0b905}
 We used a number of standard methods to test our models' validity, aiming to
@@
 \subsubsection{Computation}
-\label{sec:orge25051a}
+\label{sec:org0216e2f}
 For each MCMC run we used arviz \citeprocitem{27}{[27]} to
@@
 \subsubsection{Comparison models}
-\label{sec:org091a4ae}
+\label{sec:org7036961}
 Ideally we would compare the results of our models with previously published
@@
 
+In addition to these comparison models, we also compared our final model with a
+Bayesian version of the model in \citeprocitem{19}{[19]},
+which we obtained simply by removing the parameters \(a^{enz:sub}\) and
+\(\tau^{enz:sub}\) from our final model. We included this extra test in order to
+verify whether including enzyme-specific parameters was worthwhile.
+
 \subsubsection{Graphical posterior predictive checks}
-\label{sec:org9b7eca4}
-Figure \ref{fig:org829db47} below shows our main model's marginal posterior
+\label{sec:org32f6ee5}
+Figure \ref{fig:orge1e57a1} below shows our main model's marginal posterior
 predictive 1\%-99\% interval for each observation, alongside the observed
@@
 \includegraphics[width=.9\linewidth]{./results/plots/ppc_km_sabio_enz.png}
-\caption{\label{fig:org829db47}Marginal posterior predictive distributions for main model fit to SABIO Km data}
+\caption{\label{fig:orge1e57a1}Marginal posterior predictive distributions for main model fit to SABIO Km data}
 \end{figure}
 
-In contrast figure \ref{fig:org96eb4db} are graphical posterior predictive checks for the really simple
+In contrast figure \ref{fig:orgf769900} are graphical posterior predictive checks for the really simple
 and simple models.
@@
 \includegraphics[width=.9\linewidth]{./results/plots/ppc_km_sabio_enz_simple.png}
-\caption{\label{fig:org96eb4db}Marginal posterior predictive distributions for simple and really simple models fit to SABIO Km data}
+\caption{\label{fig:orgf769900}Marginal posterior predictive distributions for simple and really simple models fit to SABIO Km data}
 \end{figure}
@@
 \subsubsection{Cross validation}
-\label{sec:org7c3b35d}
+\label{sec:orgdeecae1}
 For a quantitative compliment to the graphical posterior predictive checks, we
@@
 \subsection{Web app}
-\label{sec:orge6374d2}
+\label{sec:org7c35f5e}
 \section{Results}
-\label{sec:org21ac059}
+\label{sec:org71a5b7b}
 \subsection{Marginal distributions of interesting parameters}
-\label{sec:org8bc318e}
+\label{sec:orgc2caead}
+We begin by looking at the marginal distributions of the model's standard
+deviation parameters. Comparing these indicates which effects varied the
+most. The comparison in figure \ref{fig:org1671fb6} shows that the most important
+hierarchical parameter is the substrate effect, followed by the ec4/substrate
+interaction, the enzyme/substrate interaction and finally the organism/substrate
+interaction. The three larger taus are all clearly greater than the estimated
+measurement error \texttt{sigma}, whereas there is overlap between the marginal
+distributions for \texttt{sigma} and \texttt{tau\_org\_sub}.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.9\linewidth]{./results/plots/sd_posteriors.png}
+\caption{\label{fig:org1671fb6}Comparison of hierarchical standard deviation parameters}
+\end{figure}
+
 \subsection{Comparison of estimated Kms with physiological metabolite concentrations}
-\label{sec:orgbc6eea3}
+\label{sec:org3c7fea9}
+Several sources suggest that the Km parameter for a particular organism,
+substrate and enzyme should have the same order of magnitude as the typical
+physiological concentration of that substrate. [ANOTHER REFERENCE HERE?] For example \citeprocitem{19}{[19]} write
+
+\begin{quote}
+\ldots{}we may hypothesize that KM values are adjusted to the order of magnitude of
+the substrate concentration. If this is the case and if a metabolite exhibits a
+particularly high concentration in a certain organism, then all corresponding KM
+values should also tend to be increased.
+\end{quote}
+
+In order to test whether this is the case, we obtained data about some
+physiological metabolite concentrations and compared these with the
+corresponding Km measurements and marginal posterior distributions from our
+model. Figure \ref{fig:org3352b54} shows the results.
+
 \begin{figure}[htbp]
@@
 \includegraphics[width=.9\linewidth]{./results/plots/hmdb_comparison.png}
-\caption{\label{fig:org83a902f}Comparison of model predictions with physiological metabolite concentrations from HMDB}
+\caption{\label{fig:org3352b54}Comparison of model predictions with physiological metabolite concentrations from HMDB}
 \end{figure}
 
+From the graph it is clear that the first part of the hypothesis is correct, at
+least for the metabolites for which we were able to obtain both physiological
+concentrations and Km measurements: both are generally in the same order of
+magnitude. However, while enzymes with higher physiological concentrations tend
+to have higher Km values, this relationship is fairly weak.
+
+This is reflected in our model's marginal posterior distribution for the
+parameter \(\tau^{org:sub}\), which concentrates closer to zero than any of the
+other hierarchical standard deviation parameters [CHECK THIS IS TRUE], as can be
+seen in figure \ref{fig:org1671fb6}
+
 \subsection{NADH vs NADPH}
-\label{sec:org1fdc6c3}
+\label{sec:org394d869}
+
+[What is the expected result??]
+
+Figure \ref{fig:org7976aa5} shows histograms of measured Km parameters from the SABIO-RK
+dataset for NADH and NADPH binding enzymes alongside histograms of our final
+model's posterior samples for these parameters. We can see that the measurements
+for NADPH tended to be higher, and this difference is also present in the posterior
+distributions.
 
@@
 \centering
-\includegraphics[width=.9\linewidth]{./results/plots/nadh_comparison.png}
-\caption{\label{fig:org3ae5622}Comparison of modelled and measured Km parameters for NADH and NADPH}
+\includegraphics[width=.9\linewidth]{./results/plots/nadh.png}
+\caption{\label{fig:org7976aa5}Comparison of modelled and measured Km parameters for NADH and NADPH}
 \end{figure}
@@
 \subsection{BRENDA vs SABIO-RK}
-\label{sec:org5f86a62}
+\label{sec:orgfd9da54}
+
+Another use for our analysis is to compare the information in the two databases.
+
+It was not clear in advance which would be more useful. On the one hand, the
+SABIO-RK data provides more specific information about the measured enzyme -
+UNIPROT ids are available for most Km measurements. BRENDA provides only the EC4
+number, which is much less specific. On the other hand, experimental conditions
+like temperature and pH are available for BRENDA measurements but not, to our
+knowledge, from the SABIO-RK database. Both of these factors could affect the
+model performance - more specific information could lead to better predictions,
+whereas missing temperature and pH information could lead to bias due to
+anomalous data not being excluded.
+
+Dirctly comparing the results for BRENDA and SABIO-RK is tricky, however,
+because the two databases use different, not necessarily compatible labelling
+strategies.
+
+Figure \ref{fig:orgb72f936} compares the marginal distributions of km
+parameters from our best SABIO-RK and BRENDA models, in order to try and
+determine which provides more information.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.9\linewidth]{./results/plots/log_km_comparison.png}
+\caption{\label{fig:orgb72f936}Comparison of measured and modelled Km parameter distributions in BRENDA and SABIO-RK}
+\end{figure}
 
 \section{Case studies}
-\label{sec:orgd3eb7c0}
+\label{sec:org181351d}
 \section{Discussion}
-\label{sec:org599ba7b}
+\label{sec:org6b53c87}
 \section{References}
-\label{sec:orgf7f9da0}
+\label{sec:org9dc0a8d}
 

### 2022-01-14 analyse.py

@@ -1,3 +1,5 @@
+"""Script for drawing graphs after running sample.py"""
+
 import os
 
 import arviz as az
@@ -8,9 +10,9 @@ from matplotlib import ticker
 from xarray.core.dataset import Dataset
 
 from prepare_data import PREPARED_DIR as PD
+from src.analysis import plot_log_km_comparison, plot_nadh_comparison, plot_vars
 
 RUN_DIR_PARENT = os.path.join("results", "runs")
-IDATA_RELPATH = os.path.join("posterior", "idata.nc")
 RUN_DIRS = list(
     filter(
         os.path.isdir,
@@ -48,8 +50,8 @@ def flatten_columns(df: pd.DataFrame, sep="_") -> pd.DataFrame:
 
 def main():
     # load data
-    idatas = {
-        os.path.basename(d): az.from_netcdf(os.path.join(d, IDATA_RELPATH))
+    posteriors = {
+        os.path.basename(d): az.from_netcdf(os.path.join(d, "posterior.nc"))
         for d in RUN_DIRS
     }
     hmdb = pd.read_csv(HMDB_PATH)
@@ -57,217 +59,26 @@ def main():
         "sabio-km": pd.read_csv(os.path.join(PD, "sabio_km", "lits.csv")),
         "brenda-km": pd.read_csv(os.path.join(PD, "brenda_km", "lits.csv")),
     }
-
-    # Check out of sample predictions
-    loo = {k: az.loo(v, pointwise=True) for k, v in idatas.items()}
-    print(loo)
-    comparison = az.compare(idatas)
-    print(comparison)
-    posterior_km = idatas["sabio-km-enz"].get("posterior")
+    posterior_km = posteriors["sabio-km-enz"].get("posterior")
+    posterior_km_brenda = posteriors["brenda-km-blk"].get("posterior")
 
     # Ensure that posterior_km has the right type (mainly to shut up the checker...)
     assert isinstance(posterior_km, Dataset)
+    assert isinstance(posterior_km_brenda, Dataset)
 
     # compare the sd parameters
-    az.plot_forest(
+    f = plot_vars(posterior_km, vars=["tau", "sigma"])
+    f.savefig(os.path.join("results", "plots", "sd_posteriors.png"))
+
+    # nadh comparison
+    f = plot_nadh_comparison(posterior_km, input_dfs["sabio-km"])
+    f.savefig(os.path.join("results", "plots", "nadh.png"))
+
+    # log km comparison
[…]

### 2022-01-14 prepare_data.py

@@ -1,4 +1,4 @@
-"""Read the data in RAW_DATA_DIR and save prepared data to PREPARED_DATA_DIR."""
+"""Read the data in RAW_DIR and save prepared data to PREPARED_DIR."""
 
 import json
 import os

### 2022-01-14 src/analysis.py

@@ -1,0 +1,99 @@
+"""Functions for analysis some results."""
+
+from typing import List
+
+import numpy as np
+import pandas as pd
+from arviz.plots.forestplot import plot_forest
+from matplotlib import pyplot as plt
+from matplotlib.figure import Figure
+from xarray.core.dataset import Dataset
+
+
+def plot_vars(posterior: Dataset, vars: List[str]) -> Figure:
+    var_regex = "|".join(vars)
+    plot_forest(
+        posterior,
+        filter_vars="regex",
+        var_names=var_regex,
+        ridgeplot_kind="hist",
+        combined=True,
+        kind="ridgeplot",
+        hdi_prob=0.999,
+    )
+    f = plt.gcf()
+    ax = plt.gca()
+    ax.set_xlabel("Parameter value")
+    ax.set_ylabel("Posterior density")
+    f.suptitle(
+        "Comparison of marginal distributions for standard deviation parameters"
+    )
+    return f
+
+
+def plot_nadh_comparison(posterior: Dataset, lits: pd.DataFrame) -> Figure:
+    samples = posterior["log_km"].to_series().reset_index()
+    is_nadh = lambda df: df["biology"].str.endswith("NADH")
+    is_nadph = lambda df: df["biology"].str.endswith("NADPH")
+    f, axes = plt.subplots(1, 2, figsize=[12, 6], sharex=True)
+    axes = axes.ravel()
+    f.suptitle(
+        "Comparison of measured and modelled Km parameters for NADH and NADPH"
+    )
+    axes[0].set_title("Measurements")
+    axes[1].set_title("Posterior samples")
+    for ax, df, bins, title, ycol in zip(
+        axes,
+        (lits, samples),
+        (25, 100),
+        ("Measurements", "Posterior samples"),
+        ("y", "log_km"),
+    ):
+        ax.set_title(title)
+        ax.hist(df.loc[is_nadh, ycol], alpha=0.6, label="nadh", bins=bins)
+        ax.hist(df.loc[is_nadph, ycol], alpha=0.6, label="nadph", bins=bins)
+        ax.set_xlabel("Km value ($\ln$ scale)")
+        ax.set_ylabel("Frequency")
+        ax.legend(frameon=False)
+    plt.tight_layout()
+    return f
+
+
+def plot_log_km_comparison(
+    posterior_sabio: Dataset,
+    posterior_brenda: Dataset,
+    lits_sabio: pd.DataFrame,
[…]

### 2022-01-26 paper.tex
Lots of changes

--- 2022-01-05 paper.tex
@@
-% Created 2022-01-04 Tue 17:13
+% Created 2022-01-20 Thu 12:40
 % Intended LaTeX compiler: pdflatex
@@
 \section{Introduction}
-\label{sec:org10d6a8f}
+\label{sec:org9123447}
 The choice of appropriate parameters for biological models has been a recurring
@@
 \section{State of the art}
-\label{sec:orgdd26fbe}
+\label{sec:org502960c}
 The problem of modelling kinetic parameter data from online databases has
@@
 \section{Method}
-\label{sec:org4bebd6f}
+\label{sec:org21ec188}
 \subsection{Data fetching}
-\label{sec:orgba54ca3}
+\label{sec:orgb834723}
 We fetched data from the SABIO-RK \citeprocitem{21}{[21]}
@@
 \subsection{Data processing}
-\label{sec:orgdfed31c}
+\label{sec:org29c1dbe}
 We made several significant data processing choices. See the library file
@@
 \subsubsection{Filtering}
-\label{sec:org8326023}
+\label{sec:orgeb3fb80}
 For each dataset and kinetic parameter, we removed all reports which failed to
@@
 \subsubsection{Grouping}
-\label{sec:orgc5b09e3}
+\label{sec:org4f0c10d}
 Instead of modelling reports directly, we chose to group together reports with
@@
 \subsection{Statistical model}
-\label{sec:org9e76746}
+\label{sec:org6de2f17}
 We used a Bayesian hierarchical regression model to describe all data. Since
@@
 \subsection{Model validation}
-\label{sec:org903f537}
+\label{sec:org9a4e03e}
 We used a number of standard methods to test our models' validity, aiming to
@@
 \subsubsection{Computation}
-\label{sec:orge25051a}
+\label{sec:org13b81c3}
 For each MCMC run we used arviz \citeprocitem{27}{[27]} to
@@
 \subsubsection{Comparison models}
-\label{sec:org091a4ae}
+\label{sec:org44dd353}
 Ideally we would compare the results of our models with previously published
@@
 
+In addition to these comparison models, we also compared our final model with a
+Bayesian version of the model in \citeprocitem{19}{[19]},
+which we obtained simply by removing the parameters \(a^{enz:sub}\) and
+\(\tau^{enz:sub}\) from our final model. We included this extra test in order to
+verify whether including enzyme-specific parameters was worthwhile.
+
 \subsubsection{Graphical posterior predictive checks}
-\label{sec:org9b7eca4}
-Figure \ref{fig:org829db47} below shows our main model's marginal posterior
+\label{sec:org802ecca}
+Figure \ref{fig:org169524e} below shows our main model's marginal posterior
 predictive 1\%-99\% interval for each observation, alongside the observed
@@
 \includegraphics[width=.9\linewidth]{./results/plots/ppc_km_sabio_enz.png}
-\caption{\label{fig:org829db47}Marginal posterior predictive distributions for main model fit to SABIO Km data}
+\caption{\label{fig:org169524e}Marginal posterior predictive distributions for main model fit to SABIO Km data}
 \end{figure}
 
-In contrast figure \ref{fig:org96eb4db} are graphical posterior predictive checks for the really simple
+In contrast figure \ref{fig:orgfb00340} are graphical posterior predictive checks for the really simple
 and simple models.
@@
 \includegraphics[width=.9\linewidth]{./results/plots/ppc_km_sabio_enz_simple.png}
-\caption{\label{fig:org96eb4db}Marginal posterior predictive distributions for simple and really simple models fit to SABIO Km data}
+\caption{\label{fig:orgfb00340}Marginal posterior predictive distributions for simple and really simple models fit to SABIO Km data}
 \end{figure}
@@
 \subsubsection{Cross validation}
-\label{sec:org7c3b35d}
+\label{sec:orgef84cd3}
 For a quantitative compliment to the graphical posterior predictive checks, we
@@
 \subsection{Web app}
-\label{sec:orge6374d2}
+\label{sec:org2b5a253}
 \section{Results}
-\label{sec:org21ac059}
+\label{sec:orgb55063c}
 \subsection{Marginal distributions of interesting parameters}
-\label{sec:org8bc318e}
+\label{sec:orgb4ce400}
+We begin by looking at the marginal distributions of the model's standard
+deviation parameters. Comparing these indicates which effects varied the
+most. The comparison in figure \ref{fig:org2ba4151} shows that the most important
+hierarchical parameter is the substrate effect, followed by the ec4/substrate
+interaction, the enzyme/substrate interaction and finally the organism/substrate
+interaction. The three larger taus are all clearly greater than the estimated
+measurement error \texttt{sigma}, whereas there is overlap between the marginal
+distributions for \texttt{sigma} and \texttt{tau\_org\_sub}.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.9\linewidth]{./results/plots/sd_posteriors.png}
+\caption{\label{fig:org2ba4151}Comparison of hierarchical standard deviation parameters}
+\end{figure}
+
 \subsection{Comparison of estimated Kms with physiological metabolite concentrations}
-\label{sec:orgbc6eea3}
+\label{sec:orgf7b0675}
+Several sources suggest that the Km parameter for a particular organism,
+substrate and enzyme should have the same order of magnitude as the typical
+physiological concentration of that substrate. [ANOTHER REFERENCE HERE?] For example \citeprocitem{19}{[19]} write
+
+\begin{quote}
+\ldots{}we may hypothesize that KM values are adjusted to the order of magnitude of
+the substrate concentration. If this is the case and if a metabolite exhibits a
+particularly high concentration in a certain organism, then all corresponding KM
+values should also tend to be increased.
+\end{quote}
+
+In order to test whether this is the case, we obtained data about some
+physiological metabolite concentrations and compared these with the
+corresponding Km measurements and marginal posterior distributions from our
+model. Figure \ref{fig:org5701bf2} shows the results.
+
 \begin{figure}[htbp]
@@
 \includegraphics[width=.9\linewidth]{./results/plots/hmdb_comparison.png}
-\caption{\label{fig:org83a902f}Comparison of model predictions with physiological metabolite concentrations from HMDB}
+\caption{\label{fig:org5701bf2}Comparison of model predictions with physiological metabolite concentrations from HMDB}
 \end{figure}
 
+From the graph it is clear that the first part of the hypothesis is correct, at
+least for the metabolites for which we were able to obtain both physiological
+concentrations and Km measurements: both are generally in the same order of
+magnitude. However, while enzymes with higher physiological concentrations tend
+to have higher Km values, this relationship is fairly weak.
+
+This is reflected in our model's marginal posterior distribution for the
+parameter \(\tau^{org:sub}\), which concentrates closer to zero than any of the
+other hierarchical standard deviation parameters [CHECK THIS IS TRUE], as can be
+seen in figure \ref{fig:org2ba4151}
+
 \subsection{NADH vs NADPH}
-\label{sec:org1fdc6c3}
+\label{sec:org92a3f6d}
+
+[What is the expected result??]
+
+Figure \ref{fig:org7b8e3d0} shows histograms of measured Km parameters from the SABIO-RK
+dataset for NADH and NADPH binding enzymes alongside histograms of our final
+model's posterior samples for these parameters. We can see that the measurements
+for NADPH tended to be higher, and this difference is also present in the posterior
+distributions.
 
@@
 \centering
-\includegraphics[width=.9\linewidth]{./results/plots/nadh_comparison.png}
-\caption{\label{fig:org3ae5622}Comparison of modelled and measured Km parameters for NADH and NADPH}
+\includegraphics[width=.9\linewidth]{./results/plots/nadh.png}
+\caption{\label{fig:org7b8e3d0}Comparison of modelled and measured Km parameters for NADH and NADPH}
 \end{figure}
@@
 \subsection{BRENDA vs SABIO-RK}
-\label{sec:org5f86a62}
+\label{sec:orgcae16de}
+
+Our results make it possible to compare the SABIO-RK and BRENDA datasets,
+allowing practitioners to make more informed modelling decisions.
+
+The SABIO-RK dataset provides more specific information than the BRENDA
+datase. UNIPROT ids are available for most Km measurements, whereas the most
+specific enzyme-level information in the BRENDA dataset is the EC4 number. In
+addition, in contrast to BRENDA, the SABIO-RK dataset provides information about
+temperature, substrate, literature reference, enzyme, pH and strain type in
+interoperable formats and without the need for potentially error-prone string
+operations.
+
+As a result of this extra information, the model fit to the SABIO-RK data is
+more specific, allowing modellers to obtain priors that reflect measurements of
+the exact enzyme they wish to model while still appropriately taking into
+account measurements of related enyzmes. In addition, the SABIO model excludes
+some sources of bias, such as reports from non-wildtype strains and atypical
+conditions that are not accounted for in the BRENDA model.
+
+The SABIO dataset is worse than the BRENDA dataset in one respect, however: at
+the time of writing there are far more reports in the BRENDA dataset. [EXACT
+NUMBERS HERE]. The model fit to the BRENDA dataset is therefore able to take
+into account more information than is available to the SABIO-RK model.
+
+There are therefore potentially conflicting factors to take into account when
+deciding which dataset to use in a given case. SABIO should be favoured in cases
+where it provides a lot of information for a given parameter or conversely when
+the information in BRENDA is too coarse or likely to be biased. In other cases
+BRENDA is likely preferable.
+
+Our results make it easier to assess which of these two scenarios apply. The
+information available for a given km is captured by the marginal posterior
+distribution for that parameter - if the marginal posterior distribution for a
+parameter in the SABIO model is very narrow, this means that the SABIO dataset
+contains a lot of information about that parameter. In order to assess the
+liklihood of the BRENDA dataset being too coarse for a parameter whose enzyme
+shares an EC number, the marginal prior distributions of the common parameters
+can be compared. If they are very different, it is likely better to use the more
+specific SABIO results. Finally, in order to assess the likelihood of bias in
+the BRENDA data, our webapp provides links to the source documents for any
+parameter measurement.
+
+Figure \ref{fig:orgdc7a843} compares the marginal distributions of km
+parameters from our best SABIO-RK and BRENDA models. Clearly there is a general
+tendency for the BRENDA kms to be higher than the SABIO kms, reflecting the fact
+that the SABIO dataset was curated to exclude non-wildtype measurements. The
+marginal posterior distributions in the BRENDA model tend to be narrower,
+reflecting the larger input dataset.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.9\linewidth]{./results/plots/log_km_comparison.png}
+\caption{\label{fig:orgdc7a843}Comparison of measured and modelled Km parameter distributions in BRENDA and SABIO-RK}
+\end{figure}
 
 \section{Case studies}
-\label{sec:orgd3eb7c0}
+\label{sec:org09980f5}
 \section{Discussion}
-\label{sec:org599ba7b}
+\label{sec:orgc78dc85}
 \section{References}
-\label{sec:orgf7f9da0}
+\label{sec:org53c82c6}
 

### 2022-01-26 analyse.py

@@ -3,10 +3,7 @@
 import os
 
 import arviz as az
-import numpy as np
 import pandas as pd
-from matplotlib import pyplot as plt
-from matplotlib import ticker
 from xarray.core.dataset import Dataset
 
 from prepare_data import PREPARED_DIR as PD
@@ -19,7 +16,6 @@ RUN_DIRS = list(
         [os.path.join(RUN_DIR_PARENT, d) for d in os.listdir(RUN_DIR_PARENT)],
     )
 )
-HMDB_PATH = os.path.join(PD, "hmdb_concs.csv")
 
 
 def flatten_columns(df: pd.DataFrame, sep="_") -> pd.DataFrame:
@@ -54,7 +50,6 @@ def main():
         os.path.basename(d): az.from_netcdf(os.path.join(d, "posterior.nc"))
         for d in RUN_DIRS
     }
-    hmdb = pd.read_csv(HMDB_PATH)
     input_dfs = {
         "sabio-km": pd.read_csv(os.path.join(PD, "sabio_km", "lits.csv")),
         "brenda-km": pd.read_csv(os.path.join(PD, "brenda_km", "lits.csv")),
@@ -75,10 +70,9 @@ def main():
     f.savefig(os.path.join("results", "plots", "nadh.png"))
 
     # log km comparison
-    f = plot_log_km_comparison(
-        posterior_km,
-        posterior_km_brenda,
-        input_dfs["sabio-km"],
-        input_dfs["brenda-km"],
-    )
+    f = plot_log_km_comparison(posterior_km, posterior_km_brenda)
     f.savefig(os.path.join("results", "plots", "log_km_comparison.png"))
+
+
+if __name__ == "__main__":
+    main()

### 2022-01-26 fetch_data.py

@@ -112,8 +112,8 @@ def main():
         sabio = pd.concat([sabio, ec_df], ignore_index=True)
         sabio.to_csv(fpath_sabio)
         sleep(0.2)
-    n_sabio = sabio["parameter.startValue"].notnull().sum()
-    print(f"Saved {n_sabio} sabio reports to {fpath_sabio}")
+    sabio.to_csv(fpath_sabio)
+    print(f"Saved {len(sabio)} sabio reports to {fpath_sabio}")
 
     if not os.path.exists(OUTPUT_FILEPATHS["hmdb_metabolite_concentrations"]):
         fpath = OUTPUT_FILEPATHS["hmdb_metabolite_concentrations"]

### 2022-01-26 generate_results.py

@@ -6,7 +6,6 @@ import os
 import arviz as az
 import toml
 import xarray
-from xarray.core.dataset import Dataset
 
 from src.model_configuration import ModelConfiguration
 from src.sampling import sample
@@ -31,7 +30,8 @@ def main():
             run_dir = os.path.join(RESULTS_DIR, mc.name)
             if not os.path.exists(run_dir):
                 os.mkdir(run_dir)
-            modes = ["posterior", "fake", "prior"]
+            modes = ["posterior"]
+            # modes = ["posterior", "fake", "prior"]
             for mode in modes:
                 input_json = os.path.join(
                     mc.data_dir, f"stan_input_{mode}.json"
@@ -43,6 +43,7 @@ def main():
                     coords=coords,
                     dims=dims,
                     sample_kwargs=mc.sample_kwargs,
+                    diagnose=True,
                 )
                 idata_file = os.path.join(run_dir, f"{mode}.nc")
                 print(f"\n***Writing inference data to {idata_file}***\n")
@@ -65,6 +66,7 @@ def main():
                         coords=coords,
                         dims=dims,
                         sample_kwargs=sample_kwargs,
+                        diagnose=False,
                     ).get("log_likelihood")
                     lliks.append(llik)
                 full_llik = xarray.concat(lliks, dim="ix_test")

### 2022-01-26 src/analysis.py

@@ -60,40 +60,45 @@ def plot_nadh_comparison(posterior: Dataset, lits: pd.DataFrame) -> Figure:
 
 
 def plot_log_km_comparison(
-    posterior_sabio: Dataset,
-    posterior_brenda: Dataset,
-    lits_sabio: pd.DataFrame,
-    lits_brenda: pd.DataFrame,
+    posterior_sabio: Dataset, posterior_brenda: Dataset
 ) -> Figure:
     posteriors = posterior_sabio, posterior_brenda
-    datasets = lits_sabio, lits_brenda
-    titles = ["SABIO-RK", "BRENDA"]
-    f, axes = plt.subplots(1, 2, figsize=[15, 5], sharey=True)
+    names = ["SABIO-RK", "BRENDA"]
+    colors = ["tab:blue", "tab:orange"]
+    f, axes = plt.subplots(1, 2, figsize=[15, 5])
     axes = axes.ravel()
-    for posterior, df, ax, title in zip(posteriors, datasets, axes, titles):
+    bins = np.linspace(0, 20, 50)
+    xlow, xhigh = axes[1].get_xlim()
+    for posterior, name, color in zip(posteriors, names, colors):
         qs = (
             posterior["log_km"]
             .quantile([0.01, 0.5, 0.99], dim=["chain", "draw"])
             .to_series()
             .unstack("quantile")
             .rename(columns={0.01: "low", 0.5: "med", 0.99: "high"})
+            .assign(width=lambda df: df["high"].subtract(df["low"]))
             .sort_values("med")
         )
-        xlow, xhigh = ax.get_xlim()
+        axes[0].hist(
+            qs["width"], bins=bins, label=name, density=True, alpha=0.6
+        )
         x = np.linspace(xlow, xhigh, len(qs))
-        df_x = pd.Series(x, index=qs.index).reindex(df["biology"])
-        ax.vlines(
+        axes[1].vlines(
             x,
             qs["low"],
             qs["high"],
-            color="tab:blue",
+            color=color,
             zorder=0,
-            label="1%-99% posterior interval",
+            label=name,
+            linewidths=0.05,
         )
-        # ax.scatter(df_x, df["y"], color="black", s=5, label="Measured value")
-        ax.set_title(title)
-        ax.set_xticks([])
-        ax.set_ylabel("Km value, $\ln$ scale")
-        ax.legend(frameon=False)
+    axes[0].set_title(
[…]

### 2022-01-26 src/data_preparation.py

@@ -1,9 +1,8 @@
 """Provides functions prepare_data_x and dataclass PrepareDataOutput."""
 
-import os
 from dataclasses import dataclass, field
 from functools import partial
-from typing import Any, Dict, List, Union
+from typing import Any, Callable, Dict, List, Union
 
 import numpy as np
 import pandas as pd
@@ -35,8 +34,8 @@ DIMS = {
     "llik": ["ix_test"],
     "yrep": ["ix_test"],
 }
-
-THIS_FILE = os.path.dirname(os.path.abspath(__file__))
+TEMPERATURE_RANGE = 10.0, 45.0
+PH_RANGE = 5.0, 9.0
 
 # types
 StanDict = Dict[str, Union[float, int, List[float], List[int]]]
@@ -51,6 +50,7 @@ class PrepareDataOutput:
     lits: pd.DataFrame
     dims: Dict[str, Any]
     number_of_cv_folds: int
+    standict_function: Callable
     standict_prior: StanDict = field(init=False)
     standict_posterior: StanDict = field(init=False)
     standicts_cv: List[StanDict] = field(init=False)
@@ -65,10 +65,13 @@ class PrepareDataOutput:
             assert isinstance(test, np.ndarray)
             splits.append([list(train), list(test)])
         get_standict_xv = partial(
-            get_standict, lits=self.lits, coords=self.coords, likelihood=True
+            self.standict_function,
+            lits=self.lits,
+            coords=self.coords,
+            likelihood=True,
         )
         get_standict_main = partial(
-            get_standict,
+            self.standict_function,
             lits=self.lits,
             coords=self.coords,
             train_ix=ix_all,
@@ -82,11 +85,6 @@ class PrepareDataOutput:
             get_standict_xv(train_ix=train_ix, test_ix=test_ix)
             for train_ix, test_ix in splits
         ]
-
-
-def replace_nulls_with_empty_set(s: pd.Series) -> pd.Series:
-    """Work around not being able to pass sets to pd.Series.fillna."""
-    return pd.Series(np.where(s.notnull(), s, frozenset()), index=s.index)
 
 
 def process_temperature_column(t: pd.Series) -> pd.Series:
@@ -114,12 +112,16 @@ def correct_brenda_dtypes(r: pd.DataFrame):
 
 
 def get_natural_ligands_col(r: pd.DataFrame, nat: pd.DataFrame):
-    return r.join(
[…]

### 2022-01-26 src/fetching.py

@@ -26,10 +26,10 @@ SABIO_FIELDS = [
     "UniprotID",
     "ECNumber",
     "Parameter",
+    "Temperature",
+    "pH",
 ]
 SABIO_PARAMETER_TYPES = ["Km", "kcat"]
-SABIO_TEMPERATURE_RANGE = "[15 TO 45]"
-SABIO_PH_RANGE = "[5 TO 9]"
 
 
 def fetch_expasy_ec_numbers() -> List[str]:
@@ -115,34 +115,18 @@ def fetch_brenda_reports(
 
 
 def fetch_sabio_data_for_ec_number(ec_number: str) -> pd.DataFrame:
-    query_string = (
-        f"ECNumber:{ec_number}"
-        f" AND TemperatureRange:{SABIO_TEMPERATURE_RANGE}"
-        f" AND pHValueRange:{SABIO_PH_RANGE}"
-    )
-    query = {"fields[]": SABIO_FIELDS, "q": query_string}
-    request = requests.post(SABIO_URL, params=query)
-    request.raise_for_status()  # raise if 404 error
-    out = (
-        pd.read_csv(StringIO(request.text), sep="\t")
-        .loc[lambda df: df["parameter.type"].isin(SABIO_PARAMETER_TYPES)]
-        .copy()
-    )
-    if len(out) == 0:
-        out = pd.DataFrame({"ECNumber": [ec_number]})
-    return out
-
-
-def fetch_sabio_reports(ec_numbers) -> pd.DataFrame:
-    out = pd.DataFrame([])
-    pbar = tqdm(ec_numbers)
-    for i, ec_number in enumerate(pbar):
-        pbar.set_description(f"Fetching data for EC {ec_number}")
-        ec_df = fetch_sabio_data_for_ec_number(ec_number)
-        out = pd.concat([out, ec_df], ignore_index=True)
-        if i % SABIO_SLEEP_INTERVAL == 0:
-            time.sleep(0.5)
-    return out
+    query = {
+        "format": "tsv",
+        "fields[]": SABIO_FIELDS,
+        "q": f'EnzymeType:"wildtype" AND IsRecombinant:false AND ECNumber:{ec_number}',
+    }
+    s = requests.Session()
+    with s.get(SABIO_URL, params=query, stream=True) as r:
+        r.raise_for_status()  # raise if 404 error
+        out = pd.read_csv(StringIO(r.text), sep="\t")
+        if len(out) == 0:
[…]

### 2022-01-26 src/sampling.py

@@ -17,6 +17,7 @@ def sample(
     input_json: str,
     coords: dict,
     dims: dict,
+    diagnose: bool,
     sample_kwargs: dict,
 ) -> InferenceData:
     """Run cmdstanpy.CmdStanModel.sample and return an InferenceData."""
@@ -26,6 +27,8 @@ def sample(
     coords["ix_train"] = [i - 1 for i in stan_input["ix_train"]]
     coords["ix_test"] = [i - 1 for i in stan_input["ix_test"]]
     mcmc = model.sample(data=stan_input, **sample_kwargs)
+    if diagnose:
+        print(mcmc.diagnose())
     return az.from_cmdstanpy(
         posterior=mcmc,
         log_likelihood="llik",

### 2022-01-27 analyse.py
Lots of changes

@@ -44,6 +44,12 @@ def flatten_columns(df: pd.DataFrame, sep="_") -> pd.DataFrame:
     return new
 
 
+def load_df(f):
+    out = pd.read_csv(f)
+    assert isinstance(out, pd.DataFrame)
+    return out
+
+
 def main():
     # load data
     posteriors = {
@@ -51,8 +57,8 @@ def main():
         for d in RUN_DIRS
     }
     input_dfs = {
-        "sabio-km": pd.read_csv(os.path.join(PD, "sabio_km", "lits.csv")),
-        "brenda-km": pd.read_csv(os.path.join(PD, "brenda_km", "lits.csv")),
+        "sabio-km": load_df(os.path.join(PD, "sabio_km", "lits.csv")),
+        "brenda-km": load_df(os.path.join(PD, "brenda_km", "lits.csv")),
     }
     posterior_km = posteriors["sabio-km-enz"].get("posterior")
     posterior_km_brenda = posteriors["brenda-km-blk"].get("posterior")

### 2022-01-27 app.py

@@ -9,16 +9,24 @@ import numpy as np
 import pandas as pd
 import streamlit as st
 from arviz.data.inference_data import InferenceData
+from genericpath import exists
 from scipy.stats.kde import gaussian_kde
 from xarray import DataArray
-
-from src.data_preparation import BIOLOGY_FCTS
+from xarray.core.dataset import Dataset
 
 SUMMARY_CSV_FILE = os.path.join("results", "app_summary.csv")
-IDATA_FILE = os.path.join("results", "runs", "blk", "posterior", "idata.nc")
-DATA_DIR = os.path.join("data", "prepared", "tenfold")
-PREPARED_DATA_FILE = os.path.join(DATA_DIR, "input_df.csv")
-COORDS_FILE = os.path.join(DATA_DIR, "coords.json")
+RUN_DIRS = {
+    "BRENDA": os.path.join("results", "runs", "brenda-km-blk"),
+    "SABIO-RK": os.path.join("results", "runs", "sabio-km-enz"),
+}
+DATA_DIRS = {
+    "BRENDA": os.path.join("data", "prepared", "brenda_km"),
+    "SABIO-RK": os.path.join("data", "prepared", "sabio_km"),
+}
+BIOLOGY_COLS = {
+    "BRENDA": ["organism", "substrate", "ec4"],
+    "SABIO-RK": ["organism", "substrate", "ec4", "uniprot_id"],
+}
 
 
 def get_lit_link(e: Any, l: Any) -> str:
@@ -27,13 +35,11 @@ def get_lit_link(e: Any, l: Any) -> str:
     return f'<a href = "{url}">{lit}</a>'
 
 
-def get_obs(
-    msmts: pd.DataFrame, organism: str, ec4: str, substrate: str
-) -> pd.DataFrame:
-    biology = "|".join([ec4, organism, substrate])
-    obs = msmts.loc[msmts["biology"] == biology].rename(
-        columns={"log_km": "log km"}
-    )
+def get_obs(reports: pd.DataFrame, biology: str) -> pd.DataFrame:
+    obs = reports.query(f"biology == '{biology}'")
+    assert isinstance(obs, pd.DataFrame)
+    obs = obs.rename(columns={"log_km": "log km"})
+    assert isinstance(obs, pd.DataFrame)
     obs["Posterior density"] = 0
     obs["reference"] = [
         get_lit_link(row["ec4"], row["literature"])
@@ -47,58 +53,52 @@ def get_obs(
[…]

### 2022-01-27 generate_results.py

@@ -4,9 +4,11 @@ import json
 import os
 
 import arviz as az
+import pandas as pd
 import toml
 import xarray
 
+from src.analysis import generate_summary_df
 from src.model_configuration import ModelConfiguration
 from src.sampling import sample
 
@@ -48,6 +50,15 @@ def main():
                 idata_file = os.path.join(run_dir, f"{mode}.nc")
                 print(f"\n***Writing inference data to {idata_file}***\n")
                 idata.to_netcdf(idata_file)
+                # Generate summary table
+                lits = pd.read_csv(os.path.join(mc.data_dir, "lits.csv"))
+                assert isinstance(lits, pd.DataFrame)
+                summary_file = os.path.join(run_dir, f"{mode}_summary.csv")
+                is_enz = "enz" in mc.data_dir
+                summary = generate_summary_df(idata, lits, is_enz)
+                print(f"\n***Writing summary table to {summary_file}***\n")
+                summary.to_csv(summary_file)
+
             if mc.run_cross_validation:
                 if mc.sample_kwargs_cross_validation is None:
                     sample_kwargs = mc.sample_kwargs

### 2022-01-27 src/analysis.py

@@ -2,6 +2,7 @@
 
 from typing import List
 
+import arviz as az
 import numpy as np
 import pandas as pd
 from arviz.plots.forestplot import plot_forest
@@ -102,3 +103,27 @@ def plot_log_km_comparison(
     axes[1].legend(frameon=False)
     plt.tight_layout()
     return f
+
+
+def generate_summary_df(
+    idata: az.InferenceData, lits: pd.DataFrame, is_enz: bool
+):
+    posterior = idata.get("posterior")
+    cd = ["chain", "draw"]
+    assert isinstance(posterior, Dataset)
+    quantiles = (
+        posterior["log_km"]
+        .quantile([0.01, 0.5, 0.99], dim=cd)
+        .to_series()
+        .unstack("quantile")
+        .add_prefix("q_")
+    )
+    mean = posterior["log_km"].mean(dim=cd).to_series().rename("posterior_mean")
+    assert isinstance(mean, pd.Series)
+    biofcts = ["organism", "substrate", "ec4"]
+    if is_enz:
+        biofcts += ["uniprot_id"]
+    out = quantiles.join(mean)
+    for fct in biofcts:
+        out = out.join(lits.groupby("biology")[fct].first())
+    return out.reset_index()

### 2022-01-27 src/data_preparation.py

@@ -201,12 +201,15 @@ def prepare_data_brenda_km(
         & reports["is_natural"]
     )
     reports["y"] = np.log(reports["km"].values)
+    reports["biology"] = (
+        reports[biology_cols].fillna("").apply("|".join, axis=1)
+    )
     lits = (
         reports.loc[cond]
-        .groupby(lit_cols)["y"]
-        .median()
+        .groupby(lit_cols)
+        .agg({"y": "median", "biology": "first"})
         .reset_index()
-        .loc[lambda df: df.groupby("organism")["y"].transform("size") > 100]
+        .loc[lambda df: df.groupby("organism")["y"].transform("size") > 50]
         .reset_index()
     )
     coords = {}
@@ -352,9 +355,10 @@ def prepare_data_sabio_km(
     raw_reports: pd.DataFrame,
     number_of_cv_folds: int = 10,
 ) -> PrepareDataOutput:
+    assert isinstance(raw_reports, pd.DataFrame)
     pp = raw_reports.rename(
         columns={
-            "Substrate": "natural_substrates",
+            "Substrate": "reaction_substrates",
             "EnzymeType": "enzyme_type",
             "PubMedID": "literature",
             "Organism": "organism",
@@ -370,6 +374,7 @@ def prepare_data_sabio_km(
             "pH": "ph",
         }
     ).replace("-", np.nan)
+    assert isinstance(pp, pd.DataFrame)
     biology_cols = ["organism", "ec4", "uniprot_id", "substrate"]
     lit_cols = biology_cols + ["literature"]
     cond = (
@@ -377,28 +382,32 @@ def prepare_data_sabio_km(
         & pp["enzyme_type"].str.contains("wildtype")
         & pp["start_value"].notnull()
         & pp["start_value"].gt(0)
+        & pp["unit"].eq("M")
         & (
             pp["temperature"].isnull()
             | pp["temperature"].astype(float).between(*TEMPERATURE_RANGE)
         )
         & (pp["ph"].isnull() | pp["ph"].astype(float).between(*PH_RANGE))
         & pp["literature"].notnull()
-        & _contains(
-            pp["natural_substrates"].fillna(""), pp["substrate"].fillna("")
-        )
     )
     reports = pp.loc[cond].copy()
-    reports["y"] = np.log(reports[["start_value", "end_value"]]).mean(axis=1)
+    reports["y"] = np.log(
+        reports[["start_value", "end_value"]]
[…]

### 2022-01-28 prepare_data.py
Update fake data generation

@@ -28,22 +28,13 @@ PREPARED_DIR = os.path.join("data", "prepared")
 # Used to generate fake data
 TRUE_MODEL_FILE = os.path.join("src", "stan", "blk.stan")
 HARDCODED_PARAMS = {
-    "km": {
-        "nu": 4,
-        "mu": -2,
-        "sigma": 0.8,
-        "tau_substrate": 1.6,
-        "tau_ec4_sub": 0.6,
-        "tau_org_sub": 1.1,
-    },
-    "kcat": {
-        "nu": 4,
-        "mu": -2,
-        "sigma": 0.8,
-        "tau_substrate": 1.6,
-        "tau_ec4_sub": 0.6,
-        "tau_org_sub": 1.1,
-    },
+    "nu": 4,
+    "mu": -2,
+    "sigma": 0.8,
+    "tau_substrate": 1.6,
+    "tau_org_sub": 1.1,
+    "tau_ec4_sub": 0.6,
+    "tau_enz_sub": 0.5,
 }
 SAMPLE_KWARGS_SIM = {
     "chains": 1,
@@ -94,6 +85,8 @@ def generate_prepared_data():
             json.dump(po.coords, f)
         with open(os.path.join(output_dir, "dims.json"), "w") as f:
             json.dump(po.dims, f)
+        with open(os.path.join(output_dir, "biology_maps.json"), "w") as f:
+            json.dump(po.biology_maps, f)
     return prepare_data_outputs
 
 
@@ -115,15 +108,13 @@ def generate_fake_data(pos: List[PrepareDataOutput]):
         fake_param_file = os.path.join(directory, "fake_data_params.json")
         with open(original_input_file, "r") as f:
             input_orig = json.load(f)
-        model_type = "km" if "km" in po.name else "kcat"
-        hardcoded_params = HARDCODED_PARAMS[model_type]
         rng_params = {
             f"a_{suff}": np.random.normal(
-                0, hardcoded_params[f"tau_{suff}"], input_orig[f"N_{suff}"]
+                0, HARDCODED_PARAMS[f"tau_{suff}"], input_orig[f"N_{suff}"]
             ).tolist()
             for suff in ["substrate", "ec4_sub", "org_sub"]
         }
-        params = {**hardcoded_params, **rng_params}
+        params = {**HARDCODED_PARAMS, **rng_params}
         model = CmdStanModel(stan_file=TRUE_MODEL_FILE)
[…]

### 2022-01-28 src/data_preparation.py
Update data preparation

@@ -51,6 +51,7 @@ class PrepareDataOutput:
     dims: Dict[str, Any]
     number_of_cv_folds: int
     standict_function: Callable
+    biology_maps: Dict[str, List[str]]
     standict_prior: StanDict = field(init=False)
     standict_posterior: StanDict = field(init=False)
     standicts_cv: List[StanDict] = field(init=False)
@@ -218,11 +219,18 @@ def prepare_data_brenda_km(
     lits["ec4_sub"] = lits["ec4"].str.cat(lits["substrate"], sep="|")
     lits["org_sub"] = lits["organism"].str.cat(lits["substrate"], sep="|")
     fcts = biology_cols + ["ec4_sub", "org_sub", "literature", "biology"]
+    fcts_with_unknowns = ["substrate", "ec4_sub", "org_sub"]
     for fct in fcts:
-        lits[fct + "_stan"] = pd.factorize(lits[fct])[0] + 1
-        coords[fct] = pd.factorize(lits[fct])[1].tolist()
-    for coord in ["substrate", "ec4_sub", "org_sub"]:
-        coords[coord] += [f"unknown {coord}"]
+        if fct in fcts_with_unknowns:
+            lits[fct + "_stan"] = pd.factorize(lits[fct])[0] + 2
+            coords[fct] = ["unknown"] + pd.factorize(lits[fct])[1].tolist()
+        else:
+            lits[fct + "_stan"] = pd.factorize(lits[fct])[0] + 1
+            coords[fct] = pd.factorize(lits[fct])[1].tolist()
+    biology_maps = {
+        col: lits.groupby("biology")[col].first().tolist()
+        for col in biology_cols
+    }
     return PrepareDataOutput(
         name=name,
         lits=lits,
@@ -231,6 +239,7 @@ def prepare_data_brenda_km(
         dims=DIMS,
         number_of_cv_folds=number_of_cv_folds,
         standict_function=get_standict_brenda,
+        biology_maps=biology_maps,
     )
 
 
@@ -356,7 +365,7 @@ def prepare_data_sabio_km(
     number_of_cv_folds: int = 10,
 ) -> PrepareDataOutput:
     assert isinstance(raw_reports, pd.DataFrame)
-    pp = raw_reports.rename(
+    reports = raw_reports.rename(
         columns={
             "Substrate": "reaction_substrates",
             "EnzymeType": "enzyme_type",
@@ -374,28 +383,34 @@ def prepare_data_sabio_km(
             "pH": "ph",
         }
[…]

### 2022-01-28 generate_results.py
Update sampling and model

@@ -28,6 +28,8 @@ def main():
             coords = json.load(f)
         with open(os.path.join(mc.data_dir, "dims.json"), "r") as f:
             dims = json.load(f)
+        with open(os.path.join(mc.data_dir, "biology_maps.json"), "r") as f:
+            biology_maps = json.load(f)
         if not mc.do_not_run:
             run_dir = os.path.join(RESULTS_DIR, mc.name)
             if not os.path.exists(run_dir):
@@ -46,6 +48,7 @@ def main():
                     dims=dims,
                     sample_kwargs=mc.sample_kwargs,
                     diagnose=True,
+                    biology_maps=biology_maps,
                 )
                 idata_file = os.path.join(run_dir, f"{mode}.nc")
                 print(f"\n***Writing inference data to {idata_file}***\n")
@@ -54,8 +57,7 @@ def main():
                 lits = pd.read_csv(os.path.join(mc.data_dir, "lits.csv"))
                 assert isinstance(lits, pd.DataFrame)
                 summary_file = os.path.join(run_dir, f"{mode}_summary.csv")
-                is_enz = "enz" in mc.data_dir
-                summary = generate_summary_df(idata, lits, is_enz)
+                summary = generate_summary_df(idata)
                 print(f"\n***Writing summary table to {summary_file}***\n")
                 summary.to_csv(summary_file)
 
@@ -78,6 +80,7 @@ def main():
                         dims=dims,
                         sample_kwargs=sample_kwargs,
                         diagnose=False,
+                        biology_maps=biology_maps,
                     ).get("log_likelihood")
                     lliks.append(llik)
                 full_llik = xarray.concat(lliks, dim="ix_test")

### 2022-01-28 src/sampling.py

@@ -19,6 +19,7 @@ def sample(
     dims: dict,
     diagnose: bool,
     sample_kwargs: dict,
+    biology_maps: dict,
 ) -> InferenceData:
     """Run cmdstanpy.CmdStanModel.sample and return an InferenceData."""
     model = CmdStanModel(stan_file=stan_file)
@@ -29,10 +30,13 @@ def sample(
     mcmc = model.sample(data=stan_input, **sample_kwargs)
     if diagnose:
         print(mcmc.diagnose())
-    return az.from_cmdstanpy(
+    idata = az.from_cmdstanpy(
         posterior=mcmc,
         log_likelihood="llik",
         observed_data=stan_input,
         coords=coords,
         dims=dims,
     )
+    for k, v in biology_maps.items():
+        idata.assign_coords({"biology_" + k: ("biology", v)}, inplace=True)
+    return idata

### 2022-01-28 src/analysis.py
Small analysis tweak

@@ -33,9 +33,14 @@ def plot_vars(posterior: Dataset, vars: List[str]) -> Figure:
 
 
 def plot_nadh_comparison(posterior: Dataset, lits: pd.DataFrame) -> Figure:
-    samples = posterior["log_km"].to_series().reset_index()
-    is_nadh = lambda df: df["biology"].str.endswith("NADH")
-    is_nadph = lambda df: df["biology"].str.endswith("NADPH")
+    samples = (
+        posterior["log_km"]
+        .to_dataframe()
+        .rename(columns={"biology_substrate": "substrate"})
+    )
+    assert isinstance(samples, pd.DataFrame)
+    is_nadh = lambda df: df["substrate"] == "NADH"
+    is_nadph = lambda df: df["substrate"] == "NADPH"
     f, axes = plt.subplots(1, 2, figsize=[12, 6], sharex=True)
     axes = axes.ravel()
     f.suptitle(
@@ -105,9 +110,7 @@ def plot_log_km_comparison(
     return f
 
 
-def generate_summary_df(
-    idata: az.InferenceData, lits: pd.DataFrame, is_enz: bool
-):
+def generate_summary_df(idata: az.InferenceData):
     posterior = idata.get("posterior")
     cd = ["chain", "draw"]
     assert isinstance(posterior, Dataset)
@@ -118,12 +121,11 @@ def generate_summary_df(
         .unstack("quantile")
         .add_prefix("q_")
     )
-    mean = posterior["log_km"].mean(dim=cd).to_series().rename("posterior_mean")
-    assert isinstance(mean, pd.Series)
-    biofcts = ["organism", "substrate", "ec4"]
-    if is_enz:
-        biofcts += ["uniprot_id"]
-    out = quantiles.join(mean)
-    for fct in biofcts:
-        out = out.join(lits.groupby("biology")[fct].first())
-    return out.reset_index()
+    mean = (
+        posterior["log_km"]
+        .mean(dim=cd)
+        .to_dataframe()
+        .rename(columns={"log_km": "posterior_mean"})
+    )
+    assert isinstance(mean, pd.DataFrame)
+    return mean.join(quantiles).reset_index()

### 2022-01-28 app.py
Fix app

@@ -1,5 +1,4 @@
 import base64
-import json
 import os
 from typing import Any
 
@@ -8,8 +7,6 @@ import arviz as az
 import numpy as np
 import pandas as pd
 import streamlit as st
-from arviz.data.inference_data import InferenceData
-from genericpath import exists
 from scipy.stats.kde import gaussian_kde
 from xarray import DataArray
 from xarray.core.dataset import Dataset
@@ -63,17 +60,17 @@ def get_table_download_link(df: pd.DataFrame, text: str, filename: str):
 def get_log_km_brenda(
     posterior: Dataset, ec4: str, organism: str, substrate: str
 ) -> DataArray:
-    if substrate not in posterior.coords["substrate"]:
-        substrate = "unknown substrate"
-    ec_sub, org_sub = (
-        f"{s}|{substrate}"
-        if f"{s}|{substrate}" in posterior.coords[coord_name]
-        else "unknown " + coord_name
-        for s, coord_name in [(ec4, "ec4_sub"), (organism, "org_sub")]
-    )
+    # if substrate not in posterior.coords["substrate"]:
+    # substrate = "unknown substrate"
+    ec4_sub = f"{ec4}|{substrate}"
+    if ec4_sub not in posterior.coords["ec4_sub"]:
+        ec4_sub = "unknown"
+    org_sub = f"{organism}|{substrate}"
+    if org_sub not in posterior.coords["org_sub"]:
+        org_sub = "unknown"
     mu = posterior["mu"]
     a_substrate = posterior["a_substrate"].sel({"substrate": substrate})
-    a_ec_sub = posterior["a_ec4_sub"].sel({"ec4_sub": ec_sub})
+    a_ec_sub = posterior["a_ec4_sub"].sel({"ec4_sub": ec4_sub})
     a_org_sub = posterior["a_org_sub"].sel({"org_sub": org_sub})
     return mu + a_substrate + a_ec_sub + a_org_sub
 
@@ -83,7 +80,7 @@ def get_log_km_sabio(
 ):
     enz_sub = f"{uniprot_id}|{substrate}"
     if enz_sub not in posterior.coords["enz_sub"]:
-        enz_sub = "unknown enz_sub"
+        enz_sub = "unknown"
     a_enz_sub = posterior["a_enz_sub"].sel({"enz_sub": enz_sub})
     return get_log_km_brenda(posterior, ec4, organism, substrate) + a_enz_sub
 
[…]

### 2022-01-28 paper.tex
Update paper

@@
-% Created 2022-01-20 Thu 12:40
+% Created 2022-01-28 Fri 15:03
 % Intended LaTeX compiler: pdflatex
@@
 \usepackage{hyperref}
-\usepackage{indentfirst}
+\usepackage{parskip}
 \author{Teddy Groves and Areti Tsigkinopoulou}
@@
  pdfsubject={},
- pdfcreator={Emacs 29.0.50 (Org mode 9.5)}, 
+ pdfcreator={Emacs 28.0.91 (Org mode 9.5.2)}, 
  pdflang={English}}
-\usepackage[notquote]{hanging}
-\begin{document}\makeatletter
+\makeatletter
 \newcommand{\citeprocitem}[2]{\hyper@linkstart{cite}{citeproc_bib_item_#1}#2\hyper@linkend}
 \makeatother
+
+\usepackage[notquote]{hanging}
+\begin{document}
 
@@
 \section{Introduction}
-\label{sec:org9123447}
+\label{sec:orge672a8a}
 The choice of appropriate parameters for biological models has been a recurring
 issue for computational biologists in the recent years. As computational models
-become increasingly prominent in the current research \citeprocitem{1}{[1]}–\citeprocitem{3}{[3]}
+become increasingly prominent in the current research
+\citeprocitem{1}{[1]}–\citeprocitem{3}{[3]}
 and lead the way to biological breakthroughs, the use of high quality parameter
@@
 strategies. Although the most popular approach is the estimation of parameters
-through optimization methods \citeprocitem{4}{[4]}, \citeprocitem{5}{[5]},
+through optimization methods
+\citeprocitem{4}{[4]}, \citeprocitem{5}{[5]},
 ensemble modelling strategies are rapidly gaining ground in the field of systems
-biology \citeprocitem{6}{[6]}–\citeprocitem{9}{[9]}
-along with the development of tools to define informative parameter priors \citeprocitem{10}{[10]}, \citeprocitem{11}{[11]}. At
+biology
+\citeprocitem{6}{[6]}–\citeprocitem{9}{[9]}
+along with the development of tools to define informative parameter priors
+\citeprocitem{10}{[10]}, \citeprocitem{11}{[11]}. At
 the same time, machine learning methods have recently been gaining popularity as
-a tool for interpreting the large amount of existing biological data \citeprocitem{12}{[12]}–\citeprocitem{14}{[14]}
+a tool for interpreting the large amount of existing biological data
+\citeprocitem{12}{[12]}–\citeprocitem{14}{[14]}
 and have also been employed for the prediction of parameter values (mostly Km
-values) in biological models \citeprocitem{15}{[15]}–\citeprocitem{17}{[17]}.
+values) in biological models
+\citeprocitem{15}{[15]}–\citeprocitem{17}{[17]}.
 
@@
 based on information retrieved from different sources and in accordance with the
-modeller’s beliefs about the reliability of each experimental value \citeprocitem{18}{[18]}. This protocol was a
+modeller’s beliefs about the reliability of each experimental value
+\citeprocitem{18}{[18]}. This protocol was a
 significant step towards making ensemble modelling more accessible and promoting
@@
 developed a hierarchical Bayesian regression model in order to perform a
-statistical analysis of parameter reports from the BRENDA \citeprocitem{20}{[20]} and SABIO-RK \citeprocitem{21}{[21]} databases. The results of this
+statistical analysis of parameter reports from the BRENDA
+\citeprocitem{20}{[20]} and SABIO-RK
+\citeprocitem{21}{[21]} databases. The results of this
 analysis can be used in our existing protocol to rationalise parameter weights,
@@
 \section{State of the art}
-\label{sec:org502960c}
+\label{sec:org03911f4}
 The problem of modelling kinetic parameter data from online databases has
@@
 \section{Method}
-\label{sec:org21ec188}
+\label{sec:org39772cb}
 \subsection{Data fetching}
-\label{sec:orgb834723}
+\label{sec:orgb256582}
 We fetched data from the SABIO-RK \citeprocitem{21}{[21]}
@@
 \subsection{Data processing}
-\label{sec:org29c1dbe}
+\label{sec:org892ca55}
 We made several significant data processing choices. See the library file
@@
 \subsubsection{Filtering}
-\label{sec:orgeb3fb80}
+\label{sec:orgc974c1b}
 For each dataset and kinetic parameter, we removed all reports which failed to
@@
 \item The enzyme must be from a wild organism
-\item The temperature, if recorded, must be between 5 and 50 degrees C
-\item The pH, if recorded, must be between 4 and 9
+\item The temperature, if recorded, must be between 10 and 45 degrees C
+\item The pH, if recorded, must be between 5 and 9
 \item The organism must be have data from at least 50 separate study/biology
@@
 \subsubsection{Grouping}
-\label{sec:org4f0c10d}
+\label{sec:orga809439}
 Instead of modelling reports directly, we chose to group together reports with
@@
 \subsection{Statistical model}
-\label{sec:org6de2f17}
+\label{sec:org880e178}
 We used a Bayesian hierarchical regression model to describe all data. Since
 both Km and Kcat parameters are constrained to be positive, we modelled them on
-natural logarithmic scale, using the same approach taken in \citeprocitem{19}{[19]}. Our model includes a global mean
+natural logarithmic scale, using the same approach taken in
+\citeprocitem{19}{[19]}. Our model includes a global mean
 parameter \(\mu\), hierarchical substrate-specific intercept parameters \(a^{sub}\)
@@
 pre-experimental information. In particular, the prior for the measurement
-distribution degrees of freedom parameter \(\nu\) follows the recommendation in \citeprocitem{24}{[24]} and is truncated below at
-zero, reflecting our view that the measurement distribution should not be
+distribution degrees of freedom parameter \(\nu\) follows the recommendation in
+\citeprocitem{24}{[24]} and is truncated below at
+one, reflecting our view that the measurement distribution should not be
 excessively heavy-tailed.
@@
 \sigma &\sim N(0, 2)[0, \infty] \\
-\mu &\sim N(-2, 1) \\
+\mu &\sim N(-1, 2) \\
 a^{sub} &\sim N(0, \tau^{sub}) \\
@@
 
-This model can be expressed as the following Stan program (see \citeprocitem{25}{[25]}):
+This model can be expressed as the following Stan program (see
+\citeprocitem{25}{[25]}):
 
 \begin{verbatim}
-/* Extends the BLK model for more specific data from the SABIO-rk database */
-
 data {
+  int<lower=1> N;
   int<lower=1> N_train;
@@
   int<lower=1,upper=N_ec4_sub> ec4_sub[N_biology];
-  int<lower=1,upper=N_enz_sub> enz_sub[N_biology];
   int<lower=1,upper=N_org_sub> org_sub[N_biology];
+  int<lower=0,upper=N_enz_sub> enz_sub[N_biology];
   int<lower=1,upper=N_substrate> substrate[N_biology];
   array[N_train] int<lower=1,upper=N_biology> biology_train;
-  vector[N_train] y_train;
+  array[N_train] int<lower=1,upper=N> ix_train;
   array[N_test] int<lower=1,upper=N_biology> biology_test;
-  vector[N_test] y_test;
+  array[N_test] int<lower=1,upper=N> ix_test;
+  vector[N] y;
   int<lower=0,upper=1> likelihood;
@@
 transformed parameters {
-  vector[N_biology] log_km =
-    mu
-    + a_substrate[substrate]
-    + a_ec4_sub[ec4_sub]
-    + a_enz_sub[enz_sub]
-    + a_org_sub[org_sub];
+  vector[N_biology] log_km = mu + a_substrate[substrate] + a_ec4_sub[ec4_sub] + a_org_sub[org_sub];
+  for (b in 1:N_biology){
+    log_km[b] += enz_sub[b] > 0 ? a_enz_sub[enz_sub[b]] : 0;
+  }
 }
 model {
-  if (likelihood){y_train ~ student_t(nu, log_km[biology_train], sigma);}
+  if (likelihood){
+    y[ix_train] ~ student_t(nu, log_km[biology_train], sigma);
+  }
   nu ~ gamma(2, 0.1);
   sigma ~ normal(0, 2);
-  mu ~ normal(-2, 1);
+  mu ~ normal(-1, 2);
   a_substrate ~ normal(0, tau_substrate);
@@
   for (n in 1:N_test){
-    llik[n] = student_t_lpdf(y_test[n] | nu, log_km[biology_test[n]], sigma);
+    llik[n] = student_t_lpdf(y[ix_test[n]] | nu, log_km[biology_test[n]], sigma);
     yrep[n] = student_t_rng(nu, log_km[biology_test[n]], sigma);
@@
 \subsection{Model validation}
-\label{sec:org9a4e03e}
+\label{sec:org9deef6f}
 We used a number of standard methods to test our models' validity, aiming to
@@
 \subsubsection{Computation}
-\label{sec:org13b81c3}
+\label{sec:orga7a514d}
 For each MCMC run we used arviz \citeprocitem{27}{[27]} to
-calculate the improved \hat{R} statistic following the method in \citeprocitem{28}{[28]} and to check for
+calculate the improved \hat{R} statistic following the method in
+\citeprocitem{28}{[28]} and to check for
 divergent transitions. If the \hat{R} statistic was sufficiently close to 1
@@
 \subsubsection{Comparison models}
-\label{sec:org44dd353}
+\label{sec:org658e9aa}
 Ideally we would compare the results of our models with previously published
@@
 \subsubsection{Graphical posterior predictive checks}
-\label{sec:org802ecca}
-Figure \ref{fig:org169524e} below shows our main model's marginal posterior
+\label{sec:org817e718}
+Figure \ref{fig:org0bd517d} below shows our main model's marginal posterior
 predictive 1\%-99\% interval for each observation, alongside the observed
@@
 \includegraphics[width=.9\linewidth]{./results/plots/ppc_km_sabio_enz.png}
-\caption{\label{fig:org169524e}Marginal posterior predictive distributions for main model fit to SABIO Km data}
+\caption{\label{fig:org0bd517d}Marginal posterior predictive distributions for main model fit to SABIO Km data}
 \end{figure}
 
-In contrast figure \ref{fig:orgfb00340} are graphical posterior predictive checks for the really simple
+In contrast figure \ref{fig:org0eeb67b} are graphical posterior predictive checks for the really simple
 and simple models.
@@
 \includegraphics[width=.9\linewidth]{./results/plots/ppc_km_sabio_enz_simple.png}
-\caption{\label{fig:orgfb00340}Marginal posterior predictive distributions for simple and really simple models fit to SABIO Km data}
+\caption{\label{fig:org0eeb67b}Marginal posterior predictive distributions for simple and really simple models fit to SABIO Km data}
 \end{figure}
@@
 \subsubsection{Cross validation}
-\label{sec:orgef84cd3}
+\label{sec:org545da5c}
 For a quantitative compliment to the graphical posterior predictive checks, we
@@
 performance. For exploratory comparison we calculated each model's approximate
-leave-one-out expected log predictive density using the method set out in \citeprocitem{29}{[29]} and implemented in \citeprocitem{27}{[27]}. To address cases where diagnostics
+leave-one-out expected log predictive density using the method set out in
+\citeprocitem{29}{[29]} and implemented in
+\citeprocitem{27}{[27]}. To address cases where diagnostics
 suggested that the approximate leave-one-out algorithm was likely to be
@@
 \subsection{Web app}
-\label{sec:org2b5a253}
+\label{sec:org90ac148}
 \section{Results}
-\label{sec:orgb55063c}
+\label{sec:org0dac0a5}
 \subsection{Marginal distributions of interesting parameters}
-\label{sec:orgb4ce400}
+\label{sec:org7d66e87}
 We begin by looking at the marginal distributions of the model's standard
 deviation parameters. Comparing these indicates which effects varied the
-most. The comparison in figure \ref{fig:org2ba4151} shows that the most important
+most. The comparison in figure \ref{fig:org87fd0cc} shows that the most important
 hierarchical parameter is the substrate effect, followed by the ec4/substrate
@@
 \includegraphics[width=.9\linewidth]{./results/plots/sd_posteriors.png}
-\caption{\label{fig:org2ba4151}Comparison of hierarchical standard deviation parameters}
+\caption{\label{fig:org87fd0cc}Comparison of hierarchical standard deviation parameters}
+\end{figure}
+
+Figure \ref{fig:org9c32565} compares the marginal distributions of km
+parameters from our best SABIO-RK and BRENDA models. The marginal posterior
+distributions in the BRENDA model tend to be narrower, reflecting the larger
+input dataset.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.9\linewidth]{./results/plots/log_km_comparison.png}
+\caption{\label{fig:org9c32565}Comparison of measured and modelled Km parameter distributions in BRENDA and SABIO-RK}
 \end{figure}
@@
 \subsection{Comparison of estimated Kms with physiological metabolite concentrations}
-\label{sec:orgf7b0675}
+\label{sec:orgf96003e}
 Several sources suggest that the Km parameter for a particular organism,
 substrate and enzyme should have the same order of magnitude as the typical
-physiological concentration of that substrate. [ANOTHER REFERENCE HERE?] For example \citeprocitem{19}{[19]} write
+physiological concentration of that substrate. [ANOTHER REFERENCE HERE?] For example
+\citeprocitem{19}{[19]} write
 
@@
 corresponding Km measurements and marginal posterior distributions from our
-model. Figure \ref{fig:org5701bf2} shows the results.
+model. Figure \ref{fig:orga8d7e8c} shows the results.
 
@@
 \includegraphics[width=.9\linewidth]{./results/plots/hmdb_comparison.png}
-\caption{\label{fig:org5701bf2}Comparison of model predictions with physiological metabolite concentrations from HMDB}
+\caption{\label{fig:orga8d7e8c}Comparison of model predictions with physiological metabolite concentrations from HMDB}
 \end{figure}
@@
 other hierarchical standard deviation parameters [CHECK THIS IS TRUE], as can be
-seen in figure \ref{fig:org2ba4151}
+seen in figure \ref{fig:org87fd0cc}
 
 \subsection{NADH vs NADPH}
-\label{sec:org92a3f6d}
+\label{sec:orgbf6cfd7}
 
-[What is the expected result??]
-
-Figure \ref{fig:org7b8e3d0} shows histograms of measured Km parameters from the SABIO-RK
+Figure \ref{fig:org5160100} shows histograms of measured Km parameters from the SABIO-RK
 dataset for NADH and NADPH binding enzymes alongside histograms of our final
@@
 \includegraphics[width=.9\linewidth]{./results/plots/nadh.png}
-\caption{\label{fig:org7b8e3d0}Comparison of modelled and measured Km parameters for NADH and NADPH}
+\caption{\label{fig:org5160100}Comparison of modelled and measured Km parameters for NADH and NADPH}
 \end{figure}
 
-\subsection{BRENDA vs SABIO-RK}
-\label{sec:orgcae16de}
+\subsection{Cofactor vs substrate}
+\label{sec:org17699d8}
+We tested whether there tended to be a systematic difference between the km of
+an enzyme's main substrate and that of a cofactor.
+
+We XXX as
+
+\subsection{Which results to use?}
+\label{sec:org604b259}
 
@@
 The SABIO-RK dataset provides more specific information than the BRENDA
-datase. UNIPROT ids are available for most Km measurements, whereas the most
+datase. UNIPROT ids are available for many Km measurements, whereas the most
 specific enzyme-level information in the BRENDA dataset is the EC4 number. In
 addition, in contrast to BRENDA, the SABIO-RK dataset provides information about
-temperature, substrate, literature reference, enzyme, pH and strain type in
+substrate, literature reference, enzyme, temperature, pH and strain type in
 interoperable formats and without the need for potentially error-prone string
@@
 account measurements of related enyzmes. In addition, the SABIO model excludes
-some sources of bias, such as reports from non-wildtype strains and atypical
-conditions that are not accounted for in the BRENDA model.
+some potential sources of bias, such as reports from non-wildtype strains and
+atypical conditions that are not accounted for in the BRENDA model.
 
@@
 There are therefore potentially conflicting factors to take into account when
-deciding which dataset to use in a given case. SABIO should be favoured in cases
-where it provides a lot of information for a given parameter or conversely when
-the information in BRENDA is too coarse or likely to be biased. In other cases
-BRENDA is likely preferable.
+deciding which dataset to use in a given case. A further complication is that,
+due to the lack of interoperable identifiers in the BRENDA dataset, it is
+difficult to directly compare the which yields better predctions.
+
+In our judgment the best approach given these circumstances is to favour the
+SABIO-RK results in cases where they provides a lot of information for a given
+parameter or when the information in BRENDA is too coarse or likely to be
+biased. In other cases the BRENDA results should be preferred.
 
@@
 
-Figure \ref{fig:orgdc7a843} compares the marginal distributions of km
-parameters from our best SABIO-RK and BRENDA models. Clearly there is a general
-tendency for the BRENDA kms to be higher than the SABIO kms, reflecting the fact
-that the SABIO dataset was curated to exclude non-wildtype measurements. The
-marginal posterior distributions in the BRENDA model tend to be narrower,
-reflecting the larger input dataset.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{./results/plots/log_km_comparison.png}
-\caption{\label{fig:orgdc7a843}Comparison of measured and modelled Km parameter distributions in BRENDA and SABIO-RK}
-\end{figure}
-
 \section{Case studies}
-\label{sec:org09980f5}
+\label{sec:org0d6b013}
 \section{Discussion}
-\label{sec:orgc78dc85}
+\label{sec:orgf4e5edd}
 \section{References}
-\label{sec:org53c82c6}
+\label{sec:orgb6ee6b6}
 
@@
 J.
-Collins, “Predictive Biology: Modelling, Understanding and Harnessing Microbial Complexity,” \textit{Nature Reviews.
-Microbiology}, vol.
+Collins, “Predictive biology: Modelling, understanding and harnessing microbial complexity,” \textit{Nature reviews.
+microbiology}, vol.
 18, no.
@@
 Salih, and S.
-Sulaimany, “Mathematical Modelling for Coronavirus Disease (COVID-19) in Predicting Future Behaviours and Sensitivity Analysis,” \textit{Mathematical Modelling of Natural Phenomena}, vol.
+Sulaimany, “Mathematical Modelling for Coronavirus Disease (COVID-19) in Predicting Future Behaviours and Sensitivity Analysis,” \textit{Mathematical modelling of natural phenomena}, vol.
 15, May 2020, doi: \href{https://doi.org/10.1051/mmnp/2020020}{10.1051/mmnp/2020020}.}
@@
 Hatzimanikatis, and L.
-Miskovic, “Large-Scale Kinetic Metabolic Models of Pseudomonas Putida KT2440 for Consistent Design of Metabolic Engineering Strategies,” \textit{Biotechnology for Biofuels}, vol.
+Miskovic, “Large-scale kinetic metabolic models of Pseudomonas putida KT2440 for consistent design of metabolic engineering strategies,” \textit{Biotechnology for biofuels}, vol.
 13, p.
@@
 D.
-Maranas, “A Genome-Scale Escherichia Coli Kinetic Metabolic Model k-Ecoli457 Satisfying Flux Data for Multiple Mutant Strains,” \textit{Nature Communications}, vol.
+Maranas, “A genome-scale Escherichia coli kinetic metabolic model k-ecoli457 satisfying flux data for multiple mutant strains,” \textit{Nature communications}, vol.
 7, no.
@@
 K.
-Nielsen, “Formulation, Construction and Analysis of Kinetic Models of Metabolism: A Review of Modelling Frameworks,” \textit{Biotechnology Advances}, vol.
+Nielsen, “Formulation, construction and analysis of kinetic models of metabolism: A review of modelling frameworks,” \textit{Biotechnology advances}, vol.
 35, no.
@@
 Baker, and R.
-Breitling, “Respectful Modeling: Addressing Uncertainty in Dynamic System Models for Molecular Biology,” \textit{Trends in Biotechnology}, vol.
+Breitling, “Respectful Modeling: Addressing Uncertainty in Dynamic System Models for Molecular Biology,” \textit{Trends in biotechnology}, vol.
 35, no.
@@
 K.
-Jha, “Automated Parameter Estimation for Biological Models Using Bayesian Statistical Model Checking,” \textit{BMC Bioinformatics}, vol.
+Jha, “Automated parameter estimation for biological models using Bayesian statistical model checking,” \textit{Bmc bioinformatics}, vol.
 16, no.
@@
 Jin, and J.
-Lowengrub, “Predicting Simulation Parameters of Biological Systems Using a Gaussian Process Model,” \textit{Statistical Analysis and Data Mining: The ASA Data Science Journal}, vol.
+Lowengrub, “Predicting simulation parameters of biological systems using a Gaussian process model,” \textit{Statistical analysis and data mining: The asa data science journal}, vol.
 5, no.
@@
 Stigter, and J.
-Molenaar, “Prediction Uncertainty Assessment of a Systems Biology Model Requires a Sample of the Full Probability Distribution of Its Parameters,” \textit{PeerJ}, vol.
+Molenaar, “Prediction uncertainty assessment of a systems biology model requires a sample of the full probability distribution of its parameters,” \textit{Peerj}, vol.
 2, p.
@@
 K.
-Nielsen, “A General Framework for Thermodynamically Consistent Parameterization and Efficient Sampling of Enzymatic Reactions,” \textit{PLOS Computational Biology}, vol.
+Nielsen, “A General Framework for Thermodynamically Consistent Parameterization and Efficient Sampling of Enzymatic Reactions,” \textit{Plos computational biology}, vol.
 11, no.
@@
 P.
-Speed, “Network Inference Using Informative Priors,” \textit{Proceedings of the National Academy of Sciences}, vol.
+Speed, “Network inference using informative priors,” \textit{Proceedings of the national academy of sciences}, vol.
 105, no.
@@
 Jayamohan, and A.
-Jérusalem, “Mechanistic Models versus Machine Learning, a Fight Worth Fighting for the Biological Community?,” \textit{Biology Letters}, vol.
+Jérusalem, “Mechanistic models versus machine learning, a fight worth fighting for the biological community?,” \textit{Biology letters}, vol.
 14, no.
@@
 A.
-Jackson, “Machine Learning and Complex Biological Data,” \textit{Genome Biology}, vol.
+Jackson, “Machine learning and complex biological data,” \textit{Genome biology}, vol.
 20, no.
@@
 M.
-Hoffman, “Machine Learning for Integrating Data in Biology and Medicine: Principles, Practice, and Opportunities,” \textit{An International Journal on Information Fusion}, vol.
+Hoffman, “Machine Learning for Integrating Data in Biology and Medicine: Principles, Practice, and Opportunities,” \textit{An international journal on information fusion}, vol.
 50, pp.
@@
 Nong, and G.
-Wu, “Predicting Km Values of Beta-Glucosidases Using Cellobiose as Substrate,” \textit{Interdisciplinary Sciences, Computational Life Sciences}, vol.
+Wu, “Predicting Km values of beta-glucosidases using cellobiose as substrate,” \textit{Interdisciplinary sciences, computational life sciences}, vol.
 4, no.
@@
 J.
-Lercher, “Deep Learning Allows Genome-Scale Prediction of Michaelis Constants from Structural Features,” \textit{PLoS biology}, vol.
+Lercher, “Deep learning allows genome-scale prediction of Michaelis constants from structural features,” \textit{Plos biology}, vol.
 19, no.
@@
 \hypertarget{citeproc_bib_item_17}{[17] F.
-Li \textit{et al.}, “Deep Learning Based Kcat Prediction Enables Improved Enzyme Constrained Model Reconstruction.” Cold Spring Harbor Laboratory, pp.
+Li \textit{et al.}, “Deep learning based kcat prediction enables improved enzyme constrained model reconstruction.” Cold Spring Harbor Laboratory, p.
 2021.08.06.455417, Aug.
@@
 Uttley, and R.
-Breitling, “Defining Informative Priors for Ensemble Modeling in Systems Biology,” \textit{Nature Protocols}, vol.
+Breitling, “Defining informative priors for ensemble modeling in systems biology,” \textit{Nature protocols}, vol.
 13, no.
@@
 Liebermeister, and E.
-Klipp, “Prediction of Enzyme Kinetic Parameters Based on Statistical Learning,” \textit{Genome Informatics}, vol.
+Klipp, “Prediction of enzyme kinetic parameters based on statistical learning,” \textit{Genome informatics}, vol.
 17, no.
 1, pp.
-80–87, 2006.}
+80–87, 2006, Accessed: Oct.
+15, 2021.
+[Online].
+Available: \url{https://pure.mpg.de/pubman/faces/ViewItemOverviewPage.jsp?itemId=item_1585340}}
 
-\hypertarget{citeproc_bib_item_20}{[20] “BRENDA - SOAP Access Help.” https://www.brenda-enzymes.org/soap.php\\#Turnover\\_Number.}
+\hypertarget{citeproc_bib_item_20}{[20] “BRENDA - SOAP access help.” Accessed: Nov. 18, 2021. [Online]. Available: \url{https://www.brenda-enzymes.org/soap.php#Turnover_Number}}
 
 \hypertarget{citeproc_bib_item_21}{[21] U.
-Wittig \textit{et al.}, “SABIO-RK\textemdash Database for Biochemical Reaction Kinetics,” \textit{Nucleic Acids Research}, vol.
+Wittig \textit{et al.}, “SABIO-RK database for biochemical reaction kinetics,” \textit{Nucleic acids research}, vol.
 40, no.
@@
 \hypertarget{citeproc_bib_item_22}{[22] A.
-Chang \textit{et al.}, “BRENDA, the ELIXIR Core Data Resource in 2021: New Developments and Updates,” \textit{Nucleic Acids Research}, vol.
+Chang \textit{et al.}, “BRENDA, the ELIXIR core data resource in 2021: New developments and updates,” \textit{Nucleic acids research}, vol.
 49, no.
@@
 Ioannidis, and C.
-Durinx, “Expasy, the Swiss Bioinformatics Resource Portal, as Designed by Its Users,” \textit{Nucleic Acids Research}, vol.
+Durinx, “Expasy, the Swiss Bioinformatics Resource Portal, as designed by its users,” \textit{Nucleic acids research}, vol.
 49, no.
@@
 J.
-Steel, “Model-Based Clustering of Non-Gaussian Panel Data Based on Skew-t Distributions,” \textit{Journal of Business \& Economic Statistics}, vol.
+Steel, “Model-Based Clustering of Non-Gaussian Panel Data Based on Skew-t Distributions,” \textit{Journal of business \& economic statistics}, vol.
 28, no.
@@
 \hypertarget{citeproc_bib_item_25}{[25] B.
-Carpenter \textit{et al.}, “Stan: A Probabilistic Programming Language,” \textit{Journal of Statistical Software}, vol.
+Carpenter \textit{et al.}, “Stan: A Probabilistic Programming Language,” \textit{Journal of statistical software}, vol.
 76, no.
@@
 \hypertarget{citeproc_bib_item_26}{[26] A.
-Gelman \textit{et al.}, “Bayesian Workflow,” \textit{arXiv:2011.01808 [stat]}, Nov.
-2020,Available: \href{https://arxiv.org/abs/2011.01808}{https://arxiv.org/abs/2011.01808}}
+Gelman \textit{et al.}, “Bayesian Workflow,” \textit{Arxiv:2011.01808 [stat]}, Nov.
+2020, Accessed: Nov.
+05, 2020.
+[Online].
+Available: \url{http://arxiv.org/abs/2011.01808}}
 
@@
 Hartikainen, and O.
-Martin, “ArviZ a Unified Library for Exploratory Analysis of Bayesian Models in Python,” \textit{Journal of Open Source Software}, vol.
+Martin, “ArviZ a unified library for exploratory analysis of Bayesian models in Python,” \textit{Journal of open source software}, vol.
 4, no.
@@
 Carpenter, and P.-C.
-Bürkner, “Rank-Normalization, Folding, and Localization: An Improved R\textasciicircum for Assessing Convergence of MCMC (with Discussion),” \textit{Bayesian Analysis}, vol.
+Bürkner, “Rank-Normalization, Folding, and Localization: An Improved R for Assessing Convergence of MCMC (with Discussion),” \textit{Bayesian analysis}, vol.
 16, no.
@@
 Gelman, and J.
-Gabry, “Practical Bayesian Model Evaluation Using Leave-One-out Cross-Validation and WAIC,” \textit{Statistics and Computing}, vol.
+Gabry, “Practical Bayesian model evaluation using leave-one-out cross-validation and WAIC,” \textit{Statistics and computing}, vol.
 27, no.

### 2022-01-29 app.py
refactor df checking, correct modes

@@ -11,6 +11,8 @@ from scipy.stats.kde import gaussian_kde
 from xarray import DataArray
 from xarray.core.dataset import Dataset
 
+from src.data_preparation import check_is_df
+
 SUMMARY_CSV_FILE = os.path.join("results", "app_summary.csv")
 RUN_DIRS = {
     "BRENDA": os.path.join("results", "runs", "brenda-km-blk"),
@@ -83,11 +85,6 @@ def get_log_km_sabio(
         enz_sub = "unknown"
     a_enz_sub = posterior["a_enz_sub"].sel({"enz_sub": enz_sub})
     return get_log_km_brenda(posterior, ec4, organism, substrate) + a_enz_sub
-
-
-def check_is_df(maybe_df) -> pd.DataFrame:
-    assert isinstance(maybe_df, pd.DataFrame)
-    return maybe_df
 
 
 # Start of app

### 2022-01-29 generate_results.py

@@ -34,8 +34,7 @@ def main():
             run_dir = os.path.join(RESULTS_DIR, mc.name)
             if not os.path.exists(run_dir):
                 os.mkdir(run_dir)
-            modes = ["posterior"]
-            # modes = ["posterior", "fake", "prior"]
+            modes = ["posterior", "fake", "prior"]
             for mode in modes:
                 input_json = os.path.join(
                     mc.data_dir, f"stan_input_{mode}.json"

### 2022-01-29 prepare_data.py

@@ -10,6 +10,7 @@ from cmdstanpy import CmdStanModel, write_stan_json
 
 from src.data_preparation import (
     PrepareDataOutput,
+    check_is_df,
     prepare_data_brenda_km,
     prepare_data_sabio_km,
     prepare_hmdb_concs,
@@ -26,7 +27,11 @@ RAW_DATA_FILES = {
 
 PREPARED_DIR = os.path.join("data", "prepared")
 # Used to generate fake data
-TRUE_MODEL_FILE = os.path.join("src", "stan", "blk.stan")
+TRUE_MODEL_FILE = {
+    "brenda_km": os.path.join("src", "stan", "blk.stan"),
+    "sabio_km": os.path.join("src", "stan", "enz.stan"),
+}
+FAKE_DATA_SOURCE_DIR = os.path.join("data", "prepared", "sabio_km")
 HARDCODED_PARAMS = {
     "nu": 4,
     "mu": -2,
@@ -48,7 +53,8 @@ def generate_prepared_data():
     """Save prepared data in the PREPARED_DATA_DIR."""
     print("Reading raw data...")
     raw_data = {
-        k: pd.read_csv(v, index_col=0) for k, v in RAW_DATA_FILES.items()
+        k: check_is_df(pd.read_csv(v, index_col=0))
+        for k, v in RAW_DATA_FILES.items()
     }
 
     hmdb_output_file = os.path.join(PREPARED_DIR, "hmdb_concs.csv")
@@ -108,14 +114,14 @@ def generate_fake_data(pos: List[PrepareDataOutput]):
         fake_param_file = os.path.join(directory, "fake_data_params.json")
         with open(original_input_file, "r") as f:
             input_orig = json.load(f)
-        rng_params = {
-            f"a_{suff}": np.random.normal(
-                0, HARDCODED_PARAMS[f"tau_{suff}"], input_orig[f"N_{suff}"]
-            ).tolist()
-            for suff in ["substrate", "ec4_sub", "org_sub"]
-        }
+        rng_params = {}
+        for suff in ["substrate", "ec4_sub", "org_sub", "enz_sub"]:
+            if f"N_{suff}" in input_orig.keys():
+                rng_params[f"a_{suff}"] = np.random.normal(
+                    0, HARDCODED_PARAMS[f"tau_{suff}"], input_orig[f"N_{suff}"]
+                ).tolist()
         params = {**HARDCODED_PARAMS, **rng_params}
[…]

### 2022-01-29 src/data_preparation.py

@@ -88,6 +88,12 @@ class PrepareDataOutput:
         ]
 
 
+def check_is_df(maybe_df) -> pd.DataFrame:
+    """Shut up the type checker!"""
+    assert isinstance(maybe_df, pd.DataFrame)
+    return maybe_df
+
+
 def process_temperature_column(t: pd.Series) -> pd.Series:
     """Convert a series of string temperatures to floats.
 

### 2022-01-29 paper.tex
Update paper

@@
-% Created 2022-01-28 Fri 15:03
+% Created 2022-01-28 Fri 15:49
 % Intended LaTeX compiler: pdflatex
@@
 \usepackage{parskip}
+\usepackage{color,soul}
 \author{Teddy Groves and Areti Tsigkinopoulou}
 \date{\today}
-\title{Statistical analysis of kinetic parameter data from online databases}
+\title{Statistical analysis of Michaelis constant data from online databases}
 \hypersetup{
  pdfauthor={Teddy Groves and Areti Tsigkinopoulou},
- pdftitle={Statistical analysis of kinetic parameter data from online databases},
+ pdftitle={Statistical analysis of Michaelis constant data from online databases},
  pdfkeywords={},
@@
 \begin{abstract}
-We fit a range of Bayesian regression models to kinetic parameter data from the
-BRENDA and SABIO-RK databases, report the results and provide tools for
-reproducing the results and using them in a systems biology workflow. We
-illustrate the intended use of our results with case studies.
+We fit a range of Bayesian regression models to analyse data from the BRENDA and
+SABIO-RK databases pertaining to Michalis constants or Km parameters. We report
+the results and provide tools for reproducing them independently and using them
+in a systems biology workflow. We illustrate the intended use of our work with case studies.
 \end{abstract}
@@
 \section{Introduction}
-\label{sec:orge672a8a}
+\label{sec:org064f21e}
 The choice of appropriate parameters for biological models has been a recurring
@@
 \section{State of the art}
-\label{sec:org03911f4}
+\label{sec:orgf8424f9}
 The problem of modelling kinetic parameter data from online databases has
@@
 \section{Method}
-\label{sec:org39772cb}
+\label{sec:org84fb08a}
 \subsection{Data fetching}
-\label{sec:orgb256582}
+\label{sec:orgd1d255a}
 We fetched data from the SABIO-RK \citeprocitem{21}{[21]}
@@
 
+In total we fetched \hl{EXACT NUMBER HERE} raw BRENDA Km reports, \hl{EXACT NUMBER HERE} SABIO-RK Km reports and \hl{EXACT NUMBER HERE} Expasy EC numbers.
+
 \subsection{Data processing}
-\label{sec:org892ca55}
+\label{sec:org83c2bf8}
 We made several significant data processing choices. See the library file
@@
 \subsubsection{Filtering}
-\label{sec:orgc974c1b}
+\label{sec:org3e7de03}
 For each dataset and kinetic parameter, we removed all reports which failed to
@@
 \subsubsection{Grouping}
-\label{sec:orga809439}
+\label{sec:org3d149b3}
 Instead of modelling reports directly, we chose to group together reports with
@@
 \subsection{Statistical model}
-\label{sec:org880e178}
-We used a Bayesian hierarchical regression model to describe all data. Since
-both Km and Kcat parameters are constrained to be positive, we modelled them on
-natural logarithmic scale, using the same approach taken in
+\label{sec:org3ef838a}
+We used a Bayesian hierarchical regression model to describe all data. Since Km
+parameters are constrained to be positive, we modelled them on natural
+logarithmic scale, using the same approach taken in
 \citeprocitem{19}{[19]}. Our model includes a global mean
@@
 
-The full model specification in tilde notation is as follows:
+The full model specification in tilde notation is shown below.
+In these lines the symbol \(\tilde\) should be understood as meaning "has the distribution".
+\(ST\), \(N\) and \(\Gamma\) represent the student-t, normal and gamma distributions respectively.
+Square brackets represent truncated probability distributions.
+For example \(X \tilde N(0,1)[0, \infty]\) means that the variable \(X\) has a normal distribution truncated below at zero, also known as a "half-normal" distribution.
 
 \begin{align*}
-\ln y &\sim ST(\nu,
-               \mu + a^{sub} + a^{enz:sub} + a^{ec4:sub} + a^{org:sub},
-               \sigma) \\
+\ln{y} &\sim ST(\nu, \hat{\ln{y}}, \sigma) \\
+\hat{\ln{y}} &= \begin{cases}
+\textbf{enz available : } \mu + a^{sub} + a^{org:sub} + a^{ec4:sub} + a^{enz:sub} \\
+\textbf{otherwise : } \mu + a^{sub} + a^{org:sub} + a^{ec4:sub} \\
+\end{cases} \\
+a^{sub} &\sim N(0, \tau^{sub}) \\
+a^{org:sub} &\sim N(0, \tau^{org:sub}) \\
+a^{ec4:sub} &\sim N(0, \tau^{ec4:sub}) \\
+a^{enz:sub} &\sim N(0, \tau^{enz:sub}) \\ 
+\mu &\sim N(-1, 2) \\
 \nu &\sim \Gamma(2, 0.1)[1, \infty] \\
 \sigma &\sim N(0, 2)[0, \infty] \\
-\mu &\sim N(-1, 2) \\
-a^{sub} &\sim N(0, \tau^{sub}) \\
-a^{enz:sub} &\sim N(0, \tau^{enz:sub}) \\
-a^{ec4:sub} &\sim N(0, \tau^{ec4:sub}) \\
-a^{org:sub} &\sim N(0, \tau^{org:sub}) \\
-\tau^{sub} &\sim N(0, 1) \\
-\tau^{enz:sub} &\sim N(0, 1) \\
-\tau^{ec4:sub} &\sim N(0, 1) \\
-\tau^{org:sub} &\sim N(0, 1) \\
+\tau^{sub} &\sim N(0, 1)[0, \infty] \\
+\tau^{org:sub} &\sim N(0, 1)[0, \infty] \\
+\tau^{ec4:sub} &\sim N(0, 1)[0, \infty] \\
+\tau^{enz:sub} &\sim N(0, 1)[0, \infty] \\
 \end{align*}
 
-This model can be expressed as the following Stan program (see
-\citeprocitem{25}{[25]}):
+This model can be expressed as the following program in the probabilistic programming language Stan program \citeprocitem{25}{[25]}:
 
 \begin{verbatim}
+/* Extends the BLK model for more specific data from the SABIO-rk database */
+
 data {
@@
   real<lower=0> tau_substrate;
+  real<lower=0> tau_org_sub;
   real<lower=0> tau_ec4_sub;
   real<lower=0> tau_enz_sub;
-  real<lower=0> tau_org_sub;
   vector<multiplier=tau_substrate>[N_substrate] a_substrate;
+  vector<multiplier=tau_org_sub>[N_org_sub] a_org_sub;
   vector<multiplier=tau_ec4_sub>[N_ec4_sub] a_ec4_sub;
   vector<multiplier=tau_enz_sub>[N_enz_sub] a_enz_sub;
-  vector<multiplier=tau_org_sub>[N_org_sub] a_org_sub;
 }
 transformed parameters {
-  vector[N_biology] log_km = mu + a_substrate[substrate] + a_ec4_sub[ec4_sub] + a_org_sub[org_sub];
+  vector[N_biology] log_km = mu
+    + a_substrate[substrate]
+    + a_ec4_sub[ec4_sub]
+    + a_org_sub[org_sub];
   for (b in 1:N_biology){
-    log_km[b] += enz_sub[b] > 0 ? a_enz_sub[enz_sub[b]] : 0;
+    if (enz_sub[b] > 1) log_km[b] += a_enz_sub[enz_sub[b]];
   }
@@
 \subsection{Model validation}
-\label{sec:org9deef6f}
+\label{sec:org2229895}
 We used a number of standard methods to test our models' validity, aiming to
@@
 \subsubsection{Computation}
-\label{sec:orga7a514d}
+\label{sec:org63aeab0}
 For each MCMC run we used arviz \citeprocitem{27}{[27]} to
@@
 \subsubsection{Comparison models}
-\label{sec:org658e9aa}
+\label{sec:org381d7f1}
 Ideally we would compare the results of our models with previously published
@@
 \subsubsection{Graphical posterior predictive checks}
-\label{sec:org817e718}
-Figure \ref{fig:org0bd517d} below shows our main model's marginal posterior
+\label{sec:orgde7e8bc}
+Figure \ref{fig:orgf2681e2} below shows our main model's marginal posterior
 predictive 1\%-99\% interval for each observation, alongside the observed
@@
 \includegraphics[width=.9\linewidth]{./results/plots/ppc_km_sabio_enz.png}
-\caption{\label{fig:org0bd517d}Marginal posterior predictive distributions for main model fit to SABIO Km data}
+\caption{\label{fig:orgf2681e2}Marginal posterior predictive distributions for main model fit to SABIO Km data}
 \end{figure}
 
-In contrast figure \ref{fig:org0eeb67b} are graphical posterior predictive checks for the really simple
+In contrast figure \ref{fig:org4c94701} are graphical posterior predictive checks for the really simple
 and simple models.
@@
 \includegraphics[width=.9\linewidth]{./results/plots/ppc_km_sabio_enz_simple.png}
-\caption{\label{fig:org0eeb67b}Marginal posterior predictive distributions for simple and really simple models fit to SABIO Km data}
+\caption{\label{fig:org4c94701}Marginal posterior predictive distributions for simple and really simple models fit to SABIO Km data}
 \end{figure}
@@
 \subsubsection{Cross validation}
-\label{sec:org545da5c}
+\label{sec:org5361885}
 For a quantitative compliment to the graphical posterior predictive checks, we
@@
 \subsection{Web app}
-\label{sec:org90ac148}
+\label{sec:orga822afd}
 \section{Results}
-\label{sec:org0dac0a5}
+\label{sec:org5aba7b0}
 \subsection{Marginal distributions of interesting parameters}
-\label{sec:org7d66e87}
+\label{sec:orga6687c0}
 We begin by looking at the marginal distributions of the model's standard
 deviation parameters. Comparing these indicates which effects varied the
-most. The comparison in figure \ref{fig:org87fd0cc} shows that the most important
+most. The comparison in figure \ref{fig:org186a847} shows that the most important
 hierarchical parameter is the substrate effect, followed by the ec4/substrate
@@
 \includegraphics[width=.9\linewidth]{./results/plots/sd_posteriors.png}
-\caption{\label{fig:org87fd0cc}Comparison of hierarchical standard deviation parameters}
+\caption{\label{fig:org186a847}Comparison of hierarchical standard deviation parameters}
 \end{figure}
 
-Figure \ref{fig:org9c32565} compares the marginal distributions of km
+Figure \ref{fig:orgd13462e} compares the marginal distributions of km
 parameters from our best SABIO-RK and BRENDA models. The marginal posterior
@@
 \includegraphics[width=.9\linewidth]{./results/plots/log_km_comparison.png}
-\caption{\label{fig:org9c32565}Comparison of measured and modelled Km parameter distributions in BRENDA and SABIO-RK}
+\caption{\label{fig:orgd13462e}Comparison of measured and modelled Km parameter distributions in BRENDA and SABIO-RK}
 \end{figure}
@@
 \subsection{Comparison of estimated Kms with physiological metabolite concentrations}
-\label{sec:orgf96003e}
+\label{sec:org7a2d906}
 Several sources suggest that the Km parameter for a particular organism,
 substrate and enzyme should have the same order of magnitude as the typical
-physiological concentration of that substrate. [ANOTHER REFERENCE HERE?] For example
+physiological concentration of that substrate. \hl{ANOTHER REFERENCE HERE?} For example
 \citeprocitem{19}{[19]} write
@@
 corresponding Km measurements and marginal posterior distributions from our
-model. Figure \ref{fig:orga8d7e8c} shows the results.
+model. Figure \ref{fig:org58d4e0d} shows the results.
 
@@
 \includegraphics[width=.9\linewidth]{./results/plots/hmdb_comparison.png}
-\caption{\label{fig:orga8d7e8c}Comparison of model predictions with physiological metabolite concentrations from HMDB}
+\caption{\label{fig:org58d4e0d}Comparison of model predictions with physiological metabolite concentrations from HMDB}
 \end{figure}
@@
 parameter \(\tau^{org:sub}\), which concentrates closer to zero than any of the
-other hierarchical standard deviation parameters [CHECK THIS IS TRUE], as can be
-seen in figure \ref{fig:org87fd0cc}
+other hierarchical standard deviation parameters \hl{CHECK THIS IS TRUE}, as can be
+seen in figure \ref{fig:org186a847}
 
 \subsection{NADH vs NADPH}
-\label{sec:orgbf6cfd7}
+\label{sec:org5dbee81}
 
-Figure \ref{fig:org5160100} shows histograms of measured Km parameters from the SABIO-RK
+Figure \ref{fig:org5972ca6} shows histograms of measured Km parameters from the SABIO-RK
 dataset for NADH and NADPH binding enzymes alongside histograms of our final
@@
 \includegraphics[width=.9\linewidth]{./results/plots/nadh.png}
-\caption{\label{fig:org5160100}Comparison of modelled and measured Km parameters for NADH and NADPH}
+\caption{\label{fig:org5972ca6}Comparison of modelled and measured Km parameters for NADH and NADPH}
 \end{figure}
@@
 \subsection{Cofactor vs substrate}
-\label{sec:org17699d8}
+\label{sec:orgf252161}
 We tested whether there tended to be a systematic difference between the km of
@@
 
-We XXX as
-
 \subsection{Which results to use?}
-\label{sec:org604b259}
+\label{sec:orgb07a856}
 
@@
 The SABIO dataset is worse than the BRENDA dataset in one respect, however: at
-the time of writing there are far more reports in the BRENDA dataset. [EXACT
-NUMBERS HERE]. The model fit to the BRENDA dataset is therefore able to take
-into account more information than is available to the SABIO-RK model.
+the time of writing there are far more reports in the BRENDA dataset. The model
+fit to the BRENDA dataset is therefore able to take into account more
+information than is available to the SABIO-RK model.
 
@@
 \section{Case studies}
-\label{sec:org0d6b013}
+\label{sec:org61cb031}
 \section{Discussion}
-\label{sec:orgf4e5edd}
+\label{sec:org6be712f}
 \section{References}
-\label{sec:orgb6ee6b6}
+\label{sec:org6bb4446}
 

### 2022-01-31 analyse.py
Improve data preparation and add concentration analysis

@@ -6,8 +6,13 @@ import arviz as az
 import pandas as pd
 from xarray.core.dataset import Dataset
 
-from prepare_data import PREPARED_DIR as PD
-from src.analysis import plot_log_km_comparison, plot_nadh_comparison, plot_vars
+from prepare_data import PREPARED_DIR
+from src.analysis import (
+    plot_concentration_comparison,
+    plot_log_km_comparison,
+    plot_nadh_comparison,
+    plot_vars,
+)
 
 RUN_DIR_PARENT = os.path.join("results", "runs")
 RUN_DIRS = list(
@@ -25,23 +30,21 @@ def flatten_columns(df: pd.DataFrame, sep="_") -> pd.DataFrame:
     annoying.
 
     """
-    n_level = len(df.columns.levels)
-    if n_level < 2:
-        raise ValueError("Not enough levels to flatten!")
-    new = df.copy()
-    new_columns = df.columns.get_level_values(0)
+    assert isinstance(
+        df.columns, pd.MultiIndex
+    ), "Not enough levels to flatten!"
+    new_column_values = df.columns.get_level_values(0)
     new_columns_name = df.columns.names[0]
-    for i in range(n_level - 1):
+    for i in range(len(df.columns.levels) - 1):
         iplusone_vals = df.columns.get_level_values(i + 1)
         iplusone_name = df.columns.names[i + 1]
-        new_columns = [
+        new_column_values = [
             f"{v_i}{sep}{v_iplusone}"
-            for v_i, v_iplusone in zip(new_columns, iplusone_vals)
+            for v_i, v_iplusone in zip(new_column_values, iplusone_vals)
         ]
         new_columns_name = f"{new_columns_name}{sep}{iplusone_name}"
-    new.columns = new_columns
-    new.rename_axis(new_columns_name, axis="columns")
-    return new
+    new_columns = pd.Index(new_column_values, name=new_columns_name)
+    return pd.DataFrame(df, columns=new_columns)
 
 
 def load_df(f):
@@ -56,9 +59,10 @@ def main():
         os.path.basename(d): az.from_netcdf(os.path.join(d, "posterior.nc"))
         for d in RUN_DIRS
     }
-    input_dfs = {
-        "sabio-km": load_df(os.path.join(PD, "sabio_km", "lits.csv")),
-        "brenda-km": load_df(os.path.join(PD, "brenda_km", "lits.csv")),
+    dfs = {
[…]

### 2022-01-31 prepare_data.py

@@ -13,14 +13,13 @@ from src.data_preparation import (
     check_is_df,
     prepare_data_brenda_km,
     prepare_data_sabio_km,
-    prepare_hmdb_concs,
+    prepare_sabio_concentrations,
 )
 
 RAW_DIR = os.path.join("data", "raw")
 RAW_DATA_FILES = {
     "brenda_kcats": os.path.join(RAW_DIR, "brenda_kcat_reports.csv"),
     "brenda_kms": os.path.join(RAW_DIR, "brenda_km_reports.csv"),
-    "hmdb_concs": os.path.join(RAW_DIR, "hmdb_metabolite_concentrations.csv"),
     "sabio_reports": os.path.join(RAW_DIR, "sabio_reports.csv"),
     "brenda_nat_subs": os.path.join(RAW_DIR, "brenda_natural_substrates.csv"),
 }
@@ -57,10 +56,12 @@ def generate_prepared_data():
         for k, v in RAW_DATA_FILES.items()
     }
 
-    hmdb_output_file = os.path.join(PREPARED_DIR, "hmdb_concs.csv")
-    hmdb_concs = prepare_hmdb_concs(raw_data["hmdb_concs"])
-    hmdb_concs.to_csv(hmdb_output_file)
-    print(f"Saved hmdb metabolite concentrations to {hmdb_output_file}")
+    sabio_conc_output_file = os.path.join(PREPARED_DIR, "sabio_concs.csv")
+    sabio_concs = prepare_sabio_concentrations(raw_data["sabio_reports"])
+    sabio_concs.to_csv(sabio_conc_output_file)
+    print(
+        f"Saved SABIO-RK metabolite concentrations to {sabio_conc_output_file}."
+    )
 
     print("Preparing data...")
     brenda_km_data = prepare_data_brenda_km(
@@ -117,8 +118,12 @@ def generate_fake_data(pos: List[PrepareDataOutput]):
         rng_params = {}
         for suff in ["substrate", "ec4_sub", "org_sub", "enz_sub"]:
             if f"N_{suff}" in input_orig.keys():
-                rng_params[f"a_{suff}"] = np.random.normal(
-                    0, HARDCODED_PARAMS[f"tau_{suff}"], input_orig[f"N_{suff}"]
+                rng_params[f"a_{suff}"] = np.array(
+                    np.random.normal(
+                        0,
+                        HARDCODED_PARAMS[f"tau_{suff}"],
+                        input_orig[f"N_{suff}"],
[…]

### 2022-01-31 src/analysis.py

@@ -10,6 +10,8 @@ from matplotlib import pyplot as plt
 from matplotlib.figure import Figure
 from xarray.core.dataset import Dataset
 
+from src.data_preparation import check_is_df
+
 
 def plot_vars(posterior: Dataset, vars: List[str]) -> Figure:
     var_regex = "|".join(vars)
@@ -58,7 +60,7 @@ def plot_nadh_comparison(posterior: Dataset, lits: pd.DataFrame) -> Figure:
         ax.set_title(title)
         ax.hist(df.loc[is_nadh, ycol], alpha=0.6, label="nadh", bins=bins)
         ax.hist(df.loc[is_nadph, ycol], alpha=0.6, label="nadph", bins=bins)
-        ax.set_xlabel("Km value ($\ln$ scale)")
+        ax.set_xlabel("Km value ($\\ln$ scale)")
         ax.set_ylabel("Frequency")
         ax.legend(frameon=False)
     plt.tight_layout()
@@ -76,12 +78,14 @@ def plot_log_km_comparison(
     bins = np.linspace(0, 20, 50)
     xlow, xhigh = axes[1].get_xlim()
     for posterior, name, color in zip(posteriors, names, colors):
-        qs = (
-            posterior["log_km"]
-            .quantile([0.01, 0.5, 0.99], dim=["chain", "draw"])
-            .to_series()
-            .unstack("quantile")
-            .rename(columns={0.01: "low", 0.5: "med", 0.99: "high"})
+        qs = check_is_df(
+            check_is_df(
+                posterior["log_km"]
+                .quantile([0.01, 0.5, 0.99], dim=["chain", "draw"])
+                .to_series()
+                .unstack("quantile")
+                .rename(columns={0.01: "low", 0.5: "med", 0.99: "high"})
+            )
             .assign(width=lambda df: df["high"].subtract(df["low"]))
             .sort_values("med")
         )
@@ -99,14 +103,70 @@ def plot_log_km_comparison(
             linewidths=0.05,
         )
     axes[0].set_title(
-        "Histograms of 1%-99% $\ln$ scale marginal Km distribution widths"
+        "Histograms of 1%-99% $\\ln$ scale marginal Km distribution widths"
     )
-    axes[1].set_title("1%-99% $\ln$ scale marginal Km distributions")
+    axes[1].set_title("1%-99% $\\ln$ scale marginal Km distributions")
[…]

### 2022-01-31 src/data_preparation.py

@@ -132,18 +132,20 @@ def get_natural_ligands_col(r: pd.DataFrame, nat: pd.DataFrame):
 
 
 def correct_brenda_colnames(raw: pd.DataFrame) -> pd.DataFrame:
-    return raw.rename(
-        columns={
-            "ecNumber": "ec4",
-            "kmValue": "km",
-            "turnoverNumber": "kcat",
-            "ligandStructureId": "ligand_structure_id",
-        }
+    return check_is_df(
+        raw.rename(
+            columns={
+                "ecNumber": "ec4",
+                "kmValue": "km",
+                "turnoverNumber": "kcat",
+                "ligandStructureId": "ligand_structure_id",
+            }
+        )
     ).pipe(make_columns_lower_case)
 
 
 def correct_brenda_nulls(reports: pd.DataFrame) -> pd.DataFrame:
-    return reports.replace(["more", -999], np.nan)
+    return check_is_df(reports.replace(["more", -999], np.nan))
 
 
 def add_columns_to_brenda_reports(
@@ -361,8 +363,21 @@ def prepare_hmdb_concs(raw: pd.DataFrame) -> pd.DataFrame:
     return raw.loc[cond].copy().assign(concentration_uM=conc)
 
 
-def _contains(s1: pd.Series, s2: pd.Series) -> pd.Series:
-    return pd.Series([t in v for t, v in zip(s2, s1)])
+def prepare_sabio_concentrations(raw: pd.DataFrame) -> pd.DataFrame:
+    cond = (
+        raw["parameter.type"].eq("concentration")
+        & raw["parameter.startValue"].notnull()
+        & raw["parameter.startValue"].gt(0)
+        & ~raw["parameter.associatedSpecies"].eq("Enzyme")
+        & raw["parameter.unit"].eq("M")
+    )
+    out = raw.loc[cond].copy()
+    out["concentration_mM"] = np.exp(
+        np.log(
+            out[["parameter.startValue", "parameter.endValue"]].multiply(1000)
+        ).mean(axis=1)
+    )
+    return out
 
 
 def prepare_data_sabio_km(
@@ -371,25 +386,28 @@ def prepare_data_sabio_km(
     number_of_cv_folds: int = 10,
 ) -> PrepareDataOutput:
     assert isinstance(raw_reports, pd.DataFrame)
-    reports = raw_reports.rename(
-        columns={
-            "Substrate": "reaction_substrates",
-            "EnzymeType": "enzyme_type",
[…]

### 2022-01-31 paper.tex
Tiny paper update

@@
-% Created 2022-01-28 Fri 15:49
+% Created 2022-01-31 Mon 09:59
 % Intended LaTeX compiler: pdflatex
@@
 \section{Introduction}
-\label{sec:org064f21e}
+\label{sec:org528b2a7}
 The choice of appropriate parameters for biological models has been a recurring
@@
 \section{State of the art}
-\label{sec:orgf8424f9}
+\label{sec:org96f3597}
 The problem of modelling kinetic parameter data from online databases has
@@
 \section{Method}
-\label{sec:org84fb08a}
+\label{sec:orged9da89}
 \subsection{Data fetching}
-\label{sec:orgd1d255a}
+\label{sec:orged7a9f2}
 We fetched data from the SABIO-RK \citeprocitem{21}{[21]}
@@
 \subsection{Data processing}
-\label{sec:org83c2bf8}
+\label{sec:orgc838b21}
 We made several significant data processing choices. See the library file
@@
 \subsubsection{Filtering}
-\label{sec:org3e7de03}
+\label{sec:orgb184998}
 For each dataset and kinetic parameter, we removed all reports which failed to
@@
 \subsubsection{Grouping}
-\label{sec:org3d149b3}
+\label{sec:orge612070}
 Instead of modelling reports directly, we chose to group together reports with
@@
 \subsection{Statistical model}
-\label{sec:org3ef838a}
+\label{sec:orgec5a53f}
 We used a Bayesian hierarchical regression model to describe all data. Since Km
@@
 \subsection{Model validation}
-\label{sec:org2229895}
+\label{sec:org77f9da4}
 We used a number of standard methods to test our models' validity, aiming to
@@
 \subsubsection{Computation}
-\label{sec:org63aeab0}
+\label{sec:org9fedc54}
 For each MCMC run we used arviz \citeprocitem{27}{[27]} to
@@
 \subsubsection{Comparison models}
-\label{sec:org381d7f1}
+\label{sec:org134ceba}
 Ideally we would compare the results of our models with previously published
@@
 \subsubsection{Graphical posterior predictive checks}
-\label{sec:orgde7e8bc}
-Figure \ref{fig:orgf2681e2} below shows our main model's marginal posterior
+\label{sec:orge66fcab}
+Figure \ref{fig:org45dd8d2} below shows our main model's marginal posterior
 predictive 1\%-99\% interval for each observation, alongside the observed
@@
 \includegraphics[width=.9\linewidth]{./results/plots/ppc_km_sabio_enz.png}
-\caption{\label{fig:orgf2681e2}Marginal posterior predictive distributions for main model fit to SABIO Km data}
+\caption{\label{fig:org45dd8d2}Marginal posterior predictive distributions for main model fit to SABIO Km data}
 \end{figure}
 
-In contrast figure \ref{fig:org4c94701} are graphical posterior predictive checks for the really simple
+In contrast figure \ref{fig:org80f5a54} are graphical posterior predictive checks for the really simple
 and simple models.
@@
 \includegraphics[width=.9\linewidth]{./results/plots/ppc_km_sabio_enz_simple.png}
-\caption{\label{fig:org4c94701}Marginal posterior predictive distributions for simple and really simple models fit to SABIO Km data}
+\caption{\label{fig:org80f5a54}Marginal posterior predictive distributions for simple and really simple models fit to SABIO Km data}
 \end{figure}
@@
 \subsubsection{Cross validation}
-\label{sec:org5361885}
+\label{sec:orgb754a9c}
 For a quantitative compliment to the graphical posterior predictive checks, we
@@
 \subsection{Web app}
-\label{sec:orga822afd}
+\label{sec:orgb84573d}
 \section{Results}
-\label{sec:org5aba7b0}
+\label{sec:org2a0d786}
 \subsection{Marginal distributions of interesting parameters}
-\label{sec:orga6687c0}
+\label{sec:org3eae566}
 We begin by looking at the marginal distributions of the model's standard
 deviation parameters. Comparing these indicates which effects varied the
-most. The comparison in figure \ref{fig:org186a847} shows that the most important
+most. The comparison in figure \ref{fig:org6d5a538} shows that the most important
 hierarchical parameter is the substrate effect, followed by the ec4/substrate
@@
 \includegraphics[width=.9\linewidth]{./results/plots/sd_posteriors.png}
-\caption{\label{fig:org186a847}Comparison of hierarchical standard deviation parameters}
+\caption{\label{fig:org6d5a538}Comparison of hierarchical standard deviation parameters}
 \end{figure}
 
-Figure \ref{fig:orgd13462e} compares the marginal distributions of km
+Figure \ref{fig:org39dd938} compares the marginal distributions of km
 parameters from our best SABIO-RK and BRENDA models. The marginal posterior
@@
 \includegraphics[width=.9\linewidth]{./results/plots/log_km_comparison.png}
-\caption{\label{fig:orgd13462e}Comparison of measured and modelled Km parameter distributions in BRENDA and SABIO-RK}
+\caption{\label{fig:org39dd938}Comparison of measured and modelled Km parameter distributions in BRENDA and SABIO-RK}
 \end{figure}
@@
 \subsection{Comparison of estimated Kms with physiological metabolite concentrations}
-\label{sec:org7a2d906}
+\label{sec:org4b904a8}
 Several sources suggest that the Km parameter for a particular organism,
@@
 corresponding Km measurements and marginal posterior distributions from our
-model. Figure \ref{fig:org58d4e0d} shows the results.
+model. Figure \ref{fig:org0dc3ba8} shows the results.
 
@@
 \includegraphics[width=.9\linewidth]{./results/plots/hmdb_comparison.png}
-\caption{\label{fig:org58d4e0d}Comparison of model predictions with physiological metabolite concentrations from HMDB}
+\caption{\label{fig:org0dc3ba8}Comparison of model predictions with physiological metabolite concentrations from HMDB}
 \end{figure}
@@
 other hierarchical standard deviation parameters \hl{CHECK THIS IS TRUE}, as can be
-seen in figure \ref{fig:org186a847}
+seen in figure \ref{fig:org6d5a538}
 
 \subsection{NADH vs NADPH}
-\label{sec:org5dbee81}
+\label{sec:orgd69ea28}
 
-Figure \ref{fig:org5972ca6} shows histograms of measured Km parameters from the SABIO-RK
+Figure \ref{fig:org9ec833e} shows histograms of measured Km parameters from the SABIO-RK
 dataset for NADH and NADPH binding enzymes alongside histograms of our final
@@
 \includegraphics[width=.9\linewidth]{./results/plots/nadh.png}
-\caption{\label{fig:org5972ca6}Comparison of modelled and measured Km parameters for NADH and NADPH}
+\caption{\label{fig:org9ec833e}Comparison of modelled and measured Km parameters for NADH and NADPH}
 \end{figure}
@@
 \subsection{Cofactor vs substrate}
-\label{sec:orgf252161}
+\label{sec:org96fe3d4}
 We tested whether there tended to be a systematic difference between the km of
@@
 \subsection{Which results to use?}
-\label{sec:orgb07a856}
+\label{sec:orgb01d466}
 
@@
 \section{Case studies}
-\label{sec:org61cb031}
+\label{sec:org960a2cc}
 \section{Discussion}
-\label{sec:org6be712f}
+\label{sec:orgcfc3d46}
 \section{References}
-\label{sec:org6bb4446}
+\label{sec:orge89aaa4}
 

### 2022-02-01 prepare_data.py
better logging and extra data processing filter

@@ -60,7 +60,7 @@ def generate_prepared_data():
     sabio_concs = prepare_sabio_concentrations(raw_data["sabio_reports"])
     sabio_concs.to_csv(sabio_conc_output_file)
     print(
-        f"Saved SABIO-RK metabolite concentrations to {sabio_conc_output_file}."
+        f"Saved {len(sabio_concs)} SABIO-RK metabolite concentrations to {sabio_conc_output_file}."
     )
 
     print("Preparing data...")
@@ -81,8 +81,13 @@ def generate_prepared_data():
             os.path.join(output_dir, f"stan_input_{s}.json")
             for s in ["prior", "posterior"]
         )
-        po.lits.to_csv(os.path.join(output_dir, "lits.csv"))
-        po.reports.to_csv(os.path.join(output_dir, "reports.csv"))
+        lit_file = os.path.join(output_dir, "lits.csv")
+        po.lits.to_csv(lit_file)
+        print(f"Saved {len(po.lits)} study/biology combinations to {lit_file}")
+        report_file = os.path.join(output_dir, "reports.csv")
+        po.reports.to_csv(report_file)
+        print(f"Saved {len(po.reports)} reports to {report_file}")
+        print("Saving handy jsons...")
         write_stan_json(input_file_posterior, po.standict_posterior)
         write_stan_json(input_file_prior, po.standict_prior)
         for i, si in enumerate(po.standicts_cv):

### 2022-02-01 src/data_preparation.py

@@ -415,6 +415,7 @@ def prepare_data_sabio_km(
         & reports["enzyme_type"].str.contains("wildtype")
         & reports["start_value"].notnull()
         & reports["start_value"].gt(0)
+        & reports["start_value"].lt(2000)
         & reports["unit"].eq("M")
         & (
             reports["temperature"].isnull()

### 2022-02-01 analyse.py
Improve analysis code

@@ -11,6 +11,7 @@ from src.analysis import (
     plot_concentration_comparison,
     plot_log_km_comparison,
     plot_nadh_comparison,
+    plot_ppc,
     plot_vars,
 )
 
@@ -70,6 +71,11 @@ def main():
     # Ensure that posterior_km has the right type (mainly to shut up the checker...)
     assert isinstance(posterior_km, Dataset)
     assert isinstance(posterior_km_brenda, Dataset)
+
+    # Draw ppc plots
+    for run_name, posterior in posteriors.items():
+        f = plot_ppc(posterior)
+        f.savefig(os.path.join("results", "plots", f"ppc_{run_name}.png"))
 
     # compare the sd parameters
     f = plot_vars(posterior_km, vars=["tau", "sigma"])

### 2022-02-01 src/analysis.py

@@ -5,6 +5,7 @@ from typing import List
 import arviz as az
 import numpy as np
 import pandas as pd
+from arviz.data.inference_data import InferenceData
 from arviz.plots.forestplot import plot_forest
 from matplotlib import pyplot as plt
 from matplotlib.figure import Figure
@@ -114,6 +115,50 @@ def plot_log_km_comparison(
     return f
 
 
+def plot_ppc(idata: InferenceData) -> Figure:
+    posterior = idata.get("posterior")
+    observed_data = idata.get("observed_data")
+    assert isinstance(posterior, Dataset)
+    assert isinstance(observed_data, Dataset)
+    df = check_is_df(
+        posterior["yrep"]
+        .quantile([0.01, 0.1, 0.5, 0.9, 0.99], dim=["chain", "draw"])
+        .to_series()
+        .unstack("quantile")
+        .add_prefix("q_")
+    )
+    df["obs"] = observed_data["y"].values
+    df = check_is_df(df.sort_values("q_0.5"))
+    f, ax = plt.subplots(figsize=[12, 5])
+    x = np.linspace(*ax.get_xlim(), num=len(df))
+    ax.vlines(
+        x,
+        df["q_0.01"],
+        df["q_0.99"],
+        alpha=0.6,
+        zorder=-1,
+        linewidth=0.1,
+        label="1%-99% posterior predictive interval",
+    )
+    ax.vlines(
+        x,
+        df["q_0.1"],
+        df["q_0.9"],
+        zorder=1,
+        linewidth=0.1,
+        label="10%-90% posterior predictive interval",
+    )
+    ax.scatter(x, df["obs"], color="black", s=1, zorder=2)
+    ax.set_ylabel("Km value ($\\ln$ scale)")
+    ax.set_xlabel("Measurement (ordered by posterior predictive median)")
+    ax.set_yticks(
+        ax.get_yticks(), ["%.2g" % t for t in np.exp(ax.get_yticks())]
+    )
+    ax.set_xticklabels([])
+    ax.set_xticklabels([])
+    return f
+
+
 def plot_concentration_comparison(
     posterior: Dataset, conc: pd.DataFrame
 ) -> Figure:
@@ -170,8 +215,7 @@ def plot_concentration_comparison(
     return f
 
 
-def generate_summary_df(idata: az.InferenceData):
-    posterior = idata.get("posterior")
[…]

### 2022-02-01 generate_results.py
Don't fit if the netcdf file already exists

@@ -7,6 +7,7 @@ import arviz as az
 import pandas as pd
 import toml
 import xarray
+from xarray.core.dataset import Dataset
 
 from src.analysis import generate_summary_df
 from src.model_configuration import ModelConfiguration
@@ -36,6 +37,9 @@ def main():
                 os.mkdir(run_dir)
             modes = ["posterior", "fake", "prior"]
             for mode in modes:
+                idata_file = os.path.join(run_dir, f"{mode}.nc")
+                if os.path.exists(idata_file):
+                    continue
                 input_json = os.path.join(
                     mc.data_dir, f"stan_input_{mode}.json"
                 )
@@ -49,18 +53,21 @@ def main():
                     diagnose=True,
                     biology_maps=biology_maps,
                 )
-                idata_file = os.path.join(run_dir, f"{mode}.nc")
                 print(f"\n***Writing inference data to {idata_file}***\n")
                 idata.to_netcdf(idata_file)
                 # Generate summary table
                 lits = pd.read_csv(os.path.join(mc.data_dir, "lits.csv"))
                 assert isinstance(lits, pd.DataFrame)
-                summary_file = os.path.join(run_dir, f"{mode}_summary.csv")
-                summary = generate_summary_df(idata)
-                print(f"\n***Writing summary table to {summary_file}***\n")
-                summary.to_csv(summary_file)
+                posterior = idata.get("posterior")
+                assert isinstance(posterior, Dataset)
+                if "log_km" in posterior.data_vars:
+                    summary_file = os.path.join(run_dir, f"{mode}_summary.csv")
+                    summary = generate_summary_df(posterior)
+                    print(f"\n***Writing summary table to {summary_file}***\n")
+                    summary.to_csv(summary_file)
 
-            if mc.run_cross_validation:
+            idata_file_cv = os.path.join(run_dir, "cv.nc")
+            if mc.run_cross_validation and not os.path.exists(idata_file_cv):
                 if mc.sample_kwargs_cross_validation is None:
                     sample_kwargs = mc.sample_kwargs
                 else:
@@ -84,10 +91,11 @@ def main():
                     lliks.append(llik)
                 full_llik = xarray.concat(lliks, dim="ix_test")
                 cv_idata = az.InferenceData(log_likelihood=full_llik)
-                idata_file = os.path.join(run_dir, "cv.nc")
-                print(f"\n***Writing inference data to {idata_file}***\n")
[…]

### 2022-02-01 paper.tex
Update paper

@@
-% Created 2022-01-31 Mon 09:59
+% Created 2022-02-01 Tue 14:40
 % Intended LaTeX compiler: pdflatex
@@
 \section{Introduction}
-\label{sec:org528b2a7}
+\label{sec:orgfc5797f}
 The choice of appropriate parameters for biological models has been a recurring
@@
 \section{State of the art}
-\label{sec:org96f3597}
+\label{sec:orgce18d1d}
 The problem of modelling kinetic parameter data from online databases has
@@
 \section{Method}
-\label{sec:orged9da89}
+\label{sec:org78f5d5e}
 \subsection{Data fetching}
-\label{sec:orged7a9f2}
+\label{sec:org583062f}
 We fetched data from the SABIO-RK \citeprocitem{21}{[21]}
@@
 \subsection{Data processing}
-\label{sec:orgc838b21}
+\label{sec:org69cb9df}
 We made several significant data processing choices. See the library file
@@
 \subsubsection{Filtering}
-\label{sec:orgb184998}
+\label{sec:org044ce9e}
 For each dataset and kinetic parameter, we removed all reports which failed to
@@
 \subsubsection{Grouping}
-\label{sec:orge612070}
+\label{sec:orga4cb87c}
 Instead of modelling reports directly, we chose to group together reports with
@@
 \subsection{Statistical model}
-\label{sec:orgec5a53f}
+\label{sec:org9456160}
 We used a Bayesian hierarchical regression model to describe all data. Since Km
@@
 \subsection{Model validation}
-\label{sec:org77f9da4}
+\label{sec:org95e5ff9}
 We used a number of standard methods to test our models' validity, aiming to
@@
 \subsubsection{Computation}
-\label{sec:org9fedc54}
+\label{sec:orga5a8cb0}
 For each MCMC run we used arviz \citeprocitem{27}{[27]} to
@@
 \subsubsection{Comparison models}
-\label{sec:org134ceba}
+\label{sec:org02fdbca}
 Ideally we would compare the results of our models with previously published
@@
 \subsubsection{Graphical posterior predictive checks}
-\label{sec:orge66fcab}
-Figure \ref{fig:org45dd8d2} below shows our main model's marginal posterior
+\label{sec:org65e677b}
+Figure \ref{fig:org9d62348} below shows our main model's marginal posterior
 predictive 1\%-99\% interval for each observation, alongside the observed
@@
 \centering
-\includegraphics[width=.9\linewidth]{./results/plots/ppc_km_sabio_enz.png}
-\caption{\label{fig:org45dd8d2}Marginal posterior predictive distributions for main model fit to SABIO Km data}
+\includegraphics[width=.9\linewidth]{./results/plots/ppc_sabio-km-enz.png}
+\caption{\label{fig:org9d62348}Marginal posterior predictive distributions for main model fit to SABIO Km data}
 \end{figure}
 
-In contrast figure \ref{fig:org80f5a54} are graphical posterior predictive checks for the really simple
-and simple models.
+Figures \ref{fig:org9e0149b} and \ref{fig:org387aabb} show
+graphical posterior predictive checks for the really simple and simple models
+fit to the same dataset.
 
@@
 \centering
-\includegraphics[width=.9\linewidth]{./results/plots/ppc_km_sabio_enz_simple.png}
-\caption{\label{fig:org80f5a54}Marginal posterior predictive distributions for simple and really simple models fit to SABIO Km data}
+\includegraphics[width=.9\linewidth]{./results/plots/ppc_sabio-km-simple.png}
+\caption{\label{fig:org9e0149b}Marginal posterior predictive distributions for simple model fit to SABIO Km data}
 \end{figure}
 
-\subsubsection{Cross validation}
-\label{sec:orgb754a9c}
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.9\linewidth]{./results/plots/ppc_sabio-km-really-simple.png}
+\caption{\label{fig:org387aabb}Marginal posterior predictive distributions for really simple model fit to SABIO Km data}
+\end{figure}
+
+From the continuous line of black dots in figure \ref{fig:org9e0149b} we can
+see that the posterior predictive intervals have approximately the same order as
+the observed data, suggesting that the model is likely overfitting, or at least
+not using any structural information from the data.
+
+In contrast figure \ref{fig:org387aabb} shows that the really simple
+model is clearly underfitting the observed data - there is no relationship
+between the order of the black dots and the blue lines.
+
+\subsubsection{{\bfseries\sffamily TODO} Cross validation}
+\label{sec:org57d7028}
 For a quantitative compliment to the graphical posterior predictive checks, we
@@
 
-\subsection{Web app}
-\label{sec:orgb84573d}
+\hl{RESULTS OF CROSS VALIDATION COMPARISON}
+
+\subsection{{\bfseries\sffamily DONE} Web app}
+\label{sec:orga4f63ea}
+In order to make our results accessible to the systems biology community, we wrote a web app that presents them in an easy to use and actionable format. This can be found at \hl{URL}.
+
+The user can choose a dataset, organism, substrate and enzyme and is then
+presented with a KDE summary of the corresponding marginal posterior
+distribution and some summary statistics describing it, which can be used to
+inform a choice of priors.
+
+For each biological category there is an option to choose "Unknown" - in this
+case the relevant marginal posterior is calculated dynamically.
+
+The app also provides links from which tables summarising all the marginal
+posteriors can be downloaded.
+
+The app was written using Streamlit \citeprocitem{30}{[30]}.
+
 \section{Results}
-\label{sec:org2a0d786}
+\label{sec:org7b65808}
 \subsection{Marginal distributions of interesting parameters}
-\label{sec:org3eae566}
+\label{sec:org5a237e8}
 We begin by looking at the marginal distributions of the model's standard
 deviation parameters. Comparing these indicates which effects varied the
-most. The comparison in figure \ref{fig:org6d5a538} shows that the most important
+most. The comparison in figure \ref{fig:org9d781a4} shows that the most important
 hierarchical parameter is the substrate effect, followed by the ec4/substrate
 interaction, the enzyme/substrate interaction and finally the organism/substrate
-interaction. The three larger taus are all clearly greater than the estimated
-measurement error \texttt{sigma}, whereas there is overlap between the marginal
-distributions for \texttt{sigma} and \texttt{tau\_org\_sub}.
+interaction. All of the distributions tend to be greater than that of the
+estimated measurement error \texttt{sigma}, suggesting that our model is able to
+distinguish all the effects from measurement noise. The substrate and
+ec4:substrate taus tend to be the greatest, suggesting that these are the most important predictors.
 
@@
 \includegraphics[width=.9\linewidth]{./results/plots/sd_posteriors.png}
-\caption{\label{fig:org6d5a538}Comparison of hierarchical standard deviation parameters}
+\caption{\label{fig:org9d781a4}Comparison of hierarchical standard deviation parameters}
 \end{figure}
 
-Figure \ref{fig:org39dd938} compares the marginal distributions of km
+Figure \ref{fig:org42ddc1d} compares the marginal distributions of km
 parameters from our best SABIO-RK and BRENDA models. The marginal posterior
@@
 \includegraphics[width=.9\linewidth]{./results/plots/log_km_comparison.png}
-\caption{\label{fig:org39dd938}Comparison of measured and modelled Km parameter distributions in BRENDA and SABIO-RK}
+\caption{\label{fig:org42ddc1d}Comparison of measured and modelled Km parameter distributions in BRENDA and SABIO-RK}
 \end{figure}
 
+\hl{What else to put here? Distributions of $\ln$ kms?}
 \subsection{Comparison of estimated Kms with physiological metabolite concentrations}
-\label{sec:org4b904a8}
+\label{sec:org4c1309c}
 Several sources suggest that the Km parameter for a particular organism,
@@
 In order to test whether this is the case, we obtained data about some
-physiological metabolite concentrations and compared these with the
-corresponding Km measurements and marginal posterior distributions from our
-model. Figure \ref{fig:org0dc3ba8} shows the results.
+physiological metabolite concentrations from SABIO-RK and compared these with
+the corresponding Km measurements and marginal posterior distributions from our
+model. Figure \ref{fig:org7c195f7} shows the results.
 
@@
 \centering
-\includegraphics[width=.9\linewidth]{./results/plots/hmdb_comparison.png}
-\caption{\label{fig:org0dc3ba8}Comparison of model predictions with physiological metabolite concentrations from HMDB}
+\includegraphics[width=.9\linewidth]{./results/plots/concentration_comparison.png}
+\caption{\label{fig:org7c195f7}Comparison of model predictions with physiological metabolite concentrations from HMDB}
 \end{figure}
 
-From the graph it is clear that the first part of the hypothesis is correct, at
-least for the metabolites for which we were able to obtain both physiological
-concentrations and Km measurements: both are generally in the same order of
-magnitude. However, while enzymes with higher physiological concentrations tend
-to have higher Km values, this relationship is fairly weak.
+The graph on the left shows that our data somewhat support the first part of the
+hypothesis: the physiological concentrations are distributed similarly to the
+posterior Km samples.
+The graph on the right suggests that \(\ln\) scale Kms tend to be 1-2 \(\ln\) mM lower than their corresponding physiological concentration reports, give or take about 5 \(\ln\) mM.
+In other words, there is only a fairly weak relationship between a particular Km and its corresonding physiological concentration.
 
@@
 other hierarchical standard deviation parameters \hl{CHECK THIS IS TRUE}, as can be
-seen in figure \ref{fig:org6d5a538}
+seen in figure \ref{fig:org9d781a4}
 
 \subsection{NADH vs NADPH}
-\label{sec:orgd69ea28}
+\label{sec:org1ec26e0}
 
-Figure \ref{fig:org9ec833e} shows histograms of measured Km parameters from the SABIO-RK
+Figure \ref{fig:orga4246f3} shows histograms of measured Km parameters from the SABIO-RK
 dataset for NADH and NADPH binding enzymes alongside histograms of our final
@@
 \includegraphics[width=.9\linewidth]{./results/plots/nadh.png}
-\caption{\label{fig:org9ec833e}Comparison of modelled and measured Km parameters for NADH and NADPH}
+\caption{\label{fig:orga4246f3}Comparison of modelled and measured Km parameters for NADH and NADPH}
 \end{figure}
 
-\subsection{Cofactor vs substrate}
-\label{sec:org96fe3d4}
+\subsection{{\bfseries\sffamily TODO} Cofactor vs substrate}
+\label{sec:org7111380}
 We tested whether there tended to be a systematic difference between the km of
@@
 \subsection{Which results to use?}
-\label{sec:orgb01d466}
+\label{sec:orgfb8accf}
 
@@
 
-\section{Case studies}
-\label{sec:org960a2cc}
-\section{Discussion}
-\label{sec:orgcfc3d46}
+\section{{\bfseries\sffamily TODO} Case studies}
+\label{sec:org093574d}
+\section{{\bfseries\sffamily TODO} Discussion}
+\label{sec:org745f9fe}
 \section{References}
-\label{sec:orge89aaa4}
+\label{sec:orgf7cebe3}
 
@@
 2017, doi: \href{https://doi.org/10.1007/s11222-016-9696-4}{10.1007/s11222-016-9696-4}.}
+
+\hypertarget{citeproc_bib_item_30}{[30] “Streamlit The fastest way to build and share data apps.” Accessed: Feb. 01, 2022. [Online]. Available: \url{https://streamlit.io/}}
 \end{document}

### 2022-02-03 generate_results.py
Split up code in generate_results script

@@ -2,11 +2,13 @@
 
 import json
 import os
+from typing import Tuple
 
 import arviz as az
 import pandas as pd
 import toml
 import xarray
+from arviz.data.inference_data import InferenceData
 from xarray.core.dataset import Dataset
 
 from src.analysis import generate_summary_df
@@ -15,6 +17,90 @@ from src.sampling import sample
 
 MODEL_CONFIGURATION_DIR = "model_configurations"
 RESULTS_DIR = os.path.join("results", "runs")
+MODES = ["posterior", "fake", "prior"]
+
+
+def load_jsons(mc) -> Tuple[dict, dict, dict]:
+    with open(os.path.join(mc.data_dir, "coords.json"), "r") as f:
+        coords = json.load(f)
+    with open(os.path.join(mc.data_dir, "dims.json"), "r") as f:
+        dims = json.load(f)
+    with open(os.path.join(mc.data_dir, "biology_maps.json"), "r") as f:
+        biology_maps = json.load(f)
+    return coords, dims, biology_maps
+
+
+def run_in_mode(mc: ModelConfiguration, mode: str) -> None:
+    coords, dims, biology_maps = load_jsons(mc)
+    run_dir = os.path.join(RESULTS_DIR, mc.name)
+    idata_file = os.path.join(run_dir, f"{mode}.nc")
+    input_json = os.path.join(mc.data_dir, f"stan_input_{mode}.json")
+    print(f"\n***Fitting model {mc.name} in {mode} mode...***\n")
+    idata = sample(
+        stan_file=mc.stan_file,
+        input_json=input_json,
+        coords=coords,
+        dims=dims,
+        sample_kwargs=mc.sample_kwargs,
+        diagnose=True,
+        biology_maps=biology_maps,
+    )
+    summary = az.summary(
+        idata, filter_vars="regex", var_names="^(?!(a_.*|yrep|llik|log_.*)).*"
+    )
+    print(summary)
+    print(f"\n***Writing inference data to {idata_file}***\n")
+    idata.to_netcdf(idata_file)
+    # Generate summary table
+    lits = pd.read_csv(os.path.join(mc.data_dir, "lits.csv"))
+    assert isinstance(lits, pd.DataFrame)
+    posterior = idata.get("posterior")
+    assert isinstance(posterior, Dataset)
+    if "log_km" in posterior.data_vars:
+        summary_file = os.path.join(run_dir, f"{mode}_summary.csv")
[…]

### 2022-02-03 analyse.py
More analysis

@@ -4,6 +4,7 @@ import os
 
 import arviz as az
 import pandas as pd
+from arviz.data.inference_data import InferenceData
 from xarray.core.dataset import Dataset
 
 from prepare_data import PREPARED_DIR
@@ -11,6 +12,7 @@ from src.analysis import (
     plot_concentration_comparison,
     plot_log_km_comparison,
     plot_nadh_comparison,
+    plot_oos_cv,
     plot_ppc,
     plot_vars,
 )
@@ -54,10 +56,26 @@ def load_df(f):
     return out
 
 
+def load_idata(f):
+    out = az.from_netcdf(f)
+    assert isinstance(out, az.InferenceData)
+    return out
+
+
+def get_log_likelihood(idata: InferenceData) -> Dataset:
+    out = idata.get("log_likelihood")
+    assert isinstance(out, Dataset)
+    return out
+
+
 def main():
     # load data
+    priors = {
+        os.path.basename(d): load_idata(os.path.join(d, "prior.nc"))
+        for d in RUN_DIRS
+    }
     posteriors = {
-        os.path.basename(d): az.from_netcdf(os.path.join(d, "posterior.nc"))
+        os.path.basename(d): load_idata(os.path.join(d, "posterior.nc"))
         for d in RUN_DIRS
     }
     dfs = {
@@ -65,32 +83,70 @@ def main():
         "brenda": load_df(os.path.join(PREPARED_DIR, "brenda_km", "lits.csv")),
         "sabio_concs": load_df(os.path.join(PREPARED_DIR, "sabio_concs.csv")),
     }
-    posterior_km = posteriors["sabio-km-enz"].get("posterior")
-    posterior_km_brenda = posteriors["brenda-km-blk"].get("posterior")
+    prior_enz = priors["sabio-km-enz"].get("posterior")
+    prior_rs = priors["sabio-km-really-simple"].get("posterior")
+    posterior_enz = posteriors["sabio-km-enz"].get("posterior")
+    posterior_brenda = posteriors["brenda-km-blk"].get("posterior")
+    posterior_blk = posteriors["sabio-km-blk"].get("posterior")
 
     # Ensure that posterior_km has the right type (mainly to shut up the checker...)
-    assert isinstance(posterior_km, Dataset)
-    assert isinstance(posterior_km_brenda, Dataset)
+    assert isinstance(posterior_enz, Dataset)
+    assert isinstance(posterior_blk, Dataset)
[…]

### 2022-02-03 src/analysis.py

@@ -1,6 +1,6 @@
 """Functions for analysis some results."""
 
-from typing import List
+from typing import Dict, List
 
 import arviz as az
 import numpy as np
@@ -18,12 +18,14 @@ def plot_vars(posterior: Dataset, vars: List[str]) -> Figure:
     var_regex = "|".join(vars)
     plot_forest(
         posterior,
+        kind="ridgeplot",
         filter_vars="regex",
         var_names=var_regex,
-        ridgeplot_kind="hist",
+        ridgeplot_quantiles=None,
+        ridgeplot_truncate=False,
         combined=True,
-        kind="ridgeplot",
         hdi_prob=0.999,
+        figsize=[12, 8],
     )
     f = plt.gcf()
     ax = plt.gca()
@@ -68,11 +70,9 @@ def plot_nadh_comparison(posterior: Dataset, lits: pd.DataFrame) -> Figure:
     return f
 
 
-def plot_log_km_comparison(
-    posterior_sabio: Dataset, posterior_brenda: Dataset
-) -> Figure:
-    posteriors = posterior_sabio, posterior_brenda
-    names = ["SABIO-RK", "BRENDA"]
+def plot_log_km_comparison(ds_dict: Dict[str, Dataset]) -> Figure:
+    posteriors = ds_dict.values()
+    names = ds_dict.keys()
     colors = ["tab:blue", "tab:orange"]
     f, axes = plt.subplots(1, 2, figsize=[15, 5])
     axes = axes.ravel()
@@ -99,9 +99,10 @@ def plot_log_km_comparison(
             qs["low"],
             qs["high"],
             color=color,
-            zorder=0,
+            zorder=-1,
             label=name,
             linewidths=0.05,
+            alpha=0.5,
         )
     axes[0].set_title(
         "Histograms of 1%-99% $\\ln$ scale marginal Km distribution widths"
@@ -233,3 +234,29 @@ def generate_summary_df(posterior: Dataset):
     )
     assert isinstance(mean, pd.DataFrame)
     return mean.join(quantiles).reset_index()
+
+
+def plot_oos_cv(idatas: Dict[str, InferenceData]) -> Figure:
+    f, ax = plt.subplots(figsize=[12, 6])
+    bins = np.linspace(-10, 0, 100)
+    for name, idata in idatas.items():
+        llik = idata.get("log_likelihood")
+        assert isinstance(llik, Dataset)
+        if "llik_oos" in llik.data_vars:
+            oos = az.extract_dataset(
[…]

### 2022-02-03 prepare_data.py
Tweak data preparation

@@ -68,8 +68,11 @@ def generate_prepared_data():
         name="brenda_km",
         raw_reports=raw_data["brenda_kms"],
         natural_ligands=raw_data["brenda_nat_subs"],
+        number_of_cv_folds=5,
     )
-    sabio_km_data = prepare_data_sabio_km("sabio_km", raw_data["sabio_reports"])
+    sabio_km_data = prepare_data_sabio_km(
+        "sabio_km", raw_data["sabio_reports"], number_of_cv_folds=5
+    )
     prepare_data_outputs = [brenda_km_data, sabio_km_data]
     for po in prepare_data_outputs:
         output_dir = os.path.join(PREPARED_DIR, po.name)

### 2022-02-03 src/data_preparation.py

@@ -185,7 +185,7 @@ def prepare_data_brenda_km(
     name: str,
     raw_reports: pd.DataFrame,
     natural_ligands: pd.DataFrame,
-    number_of_cv_folds: int = 10,
+    number_of_cv_folds,
 ) -> PrepareDataOutput:
     """get dataframe of study/km combinations ("lits")
 
@@ -383,7 +383,7 @@ def prepare_sabio_concentrations(raw: pd.DataFrame) -> pd.DataFrame:
 def prepare_data_sabio_km(
     name: str,
     raw_reports: pd.DataFrame,
-    number_of_cv_folds: int = 10,
+    number_of_cv_folds,
 ) -> PrepareDataOutput:
     assert isinstance(raw_reports, pd.DataFrame)
     reports = check_is_df(

### 2022-02-03 paper.tex
Update paper

@@
-% Created 2022-02-01 Tue 14:40
+% Created 2022-02-03 Thu 14:55
 % Intended LaTeX compiler: pdflatex
@@
 \section{Introduction}
-\label{sec:orgfc5797f}
+\label{sec:org9cc370f}
 The choice of appropriate parameters for biological models has been a recurring
@@
 \section{State of the art}
-\label{sec:orgce18d1d}
+\label{sec:org1ec53bd}
 The problem of modelling kinetic parameter data from online databases has
@@
 \begin{equation}
-y_ijk \sim N(\mu_i + \alpha_ij + \beta_ik, \sigma)
+y_{ijk} \sim N(\mu_i + \alpha_ij + \beta_ik, \sigma)
 \end{equation}
@@
 \section{Method}
-\label{sec:org78f5d5e}
+\label{sec:orgb68824a}
 \subsection{Data fetching}
-\label{sec:org583062f}
+\label{sec:orgcc0ecb4}
 We fetched data from the SABIO-RK \citeprocitem{21}{[21]}
@@
 \subsection{Data processing}
-\label{sec:org69cb9df}
+\label{sec:org2adc330}
 We made several significant data processing choices. See the library file
@@
 \subsubsection{Filtering}
-\label{sec:org044ce9e}
+\label{sec:org80f3330}
 For each dataset and kinetic parameter, we removed all reports which failed to
@@
 \subsubsection{Grouping}
-\label{sec:orga4cb87c}
+\label{sec:org41b73d8}
 Instead of modelling reports directly, we chose to group together reports with
@@
 \subsection{Statistical model}
-\label{sec:org9456160}
+\label{sec:org5d80a30}
 We used a Bayesian hierarchical regression model to describe all data. Since Km
@@
 logarithmic scale, using the same approach taken in
-\citeprocitem{19}{[19]}. Our model includes a global mean
-parameter \(\mu\), hierarchical substrate-specific intercept parameters \(a^{sub}\)
-and hierarchical intercept parameters \(a^{enz:sub}\), \(a^{ec4:sub}\) and
-\(a^{org:sub}\) specific to interactions of substrate and enzyme, ec4 number and
-organism respectively. In addition we used a student-T measurement model with
-latent standard deviation and degrees of freedom \(\sigma\) and \(\nu\).
+\citeprocitem{19}{[19]}, but replacing the organism:substrate
+interaction with enzyme:substrate, where this information is available.
+
+Our model includes a global mean parameter \(\mu\), hierarchical
+substrate-specific intercept parameters \(a^{sub}\) and hierarchical intercept
+parameters \(a^{enz:sub}\), \(a^{ec4:sub}\) and specific to interactions of
+substrate and enzyme and ec4 number respectively. In addition we used a
+student-T measurement model with latent standard deviation and degrees of
+freedom \(\sigma\) and \(\nu\).
 
@@
 \citeprocitem{24}{[24]} and is truncated below at
-one, reflecting our view that the measurement distribution should not be
+2, reflecting our view that the measurement distribution should not be
 excessively heavy-tailed.
 
-The full model specification in tilde notation is shown below.
-In these lines the symbol \(\tilde\) should be understood as meaning "has the distribution".
-\(ST\), \(N\) and \(\Gamma\) represent the student-t, normal and gamma distributions respectively.
-Square brackets represent truncated probability distributions.
-For example \(X \tilde N(0,1)[0, \infty]\) means that the variable \(X\) has a normal distribution truncated below at zero, also known as a "half-normal" distribution.
+The full model specification in tilde notation is shown below. In these lines
+the symbol \(\tilde\) should be understood as meaning "has the
+distribution". \(ST\), \(N\), \(LN\) and \(\Gamma\) represent the student-t, normal,
+log-normal and gamma distributions respectively. Square brackets represent
+truncated probability distributions. For example \(X \tilde N(0,1)[0, \infty]\)
+means that the variable \(X\) has a normal distribution truncated below at zero,
+also known as a "half-normal" distribution.
 
@@
 \hat{\ln{y}} &= \begin{cases}
-\textbf{enz available : } \mu + a^{sub} + a^{org:sub} + a^{ec4:sub} + a^{enz:sub} \\
-\textbf{otherwise : } \mu + a^{sub} + a^{org:sub} + a^{ec4:sub} \\
+\textbf{enz available : } \mu + a^{sub} + a^{ec4:sub} + a^{enz:sub} \\
+\textbf{otherwise : } \mu + a^{sub} + a^{ec4:sub} \\
 \end{cases} \\
 a^{sub} &\sim N(0, \tau^{sub}) \\
-a^{org:sub} &\sim N(0, \tau^{org:sub}) \\
 a^{ec4:sub} &\sim N(0, \tau^{ec4:sub}) \\
 a^{enz:sub} &\sim N(0, \tau^{enz:sub}) \\ 
-\mu &\sim N(-1, 2) \\
-\nu &\sim \Gamma(2, 0.1)[1, \infty] \\
-\sigma &\sim N(0, 2)[0, \infty] \\
-\tau^{sub} &\sim N(0, 1)[0, \infty] \\
-\tau^{org:sub} &\sim N(0, 1)[0, \infty] \\
-\tau^{ec4:sub} &\sim N(0, 1)[0, \infty] \\
-\tau^{enz:sub} &\sim N(0, 1)[0, \infty] \\
+\mu &\sim N(-2, 2) \\
+\nu &\sim \Gamma(2, 0.1)[2, \infty] \\
+\sigma &\sim LN(0, 0.2) \\
+\tau^{sub} &\sim LN(0, 0.3) \\
+\tau^{ec4:sub} &\sim LN(0, 0.2) \\
+\tau^{enz:sub} &\sim LN(-2, 0.2) \\
 \end{align*}
 
-This model can be expressed as the following program in the probabilistic programming language Stan program \citeprocitem{25}{[25]}:
+Prior distributions for parameters with log-normal priors are shown graphically
+in figure \ref{fig:org777eb93}.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.9\linewidth]{./results/plots/sd_priors_enz.png}
+\caption{\label{fig:org777eb93}KDE plots of prior distributions for log-normal parameters}
+\end{figure}
+
+Our model can be expressed as the following program in the probabilistic programming language Stan program \citeprocitem{25}{[25]}:
 
@@
 parameters {
-  real<lower=1> nu;
+  real<lower=2> nu;
   real mu;
   real<lower=0> sigma;
-  real<lower=0> tau_substrate;
-  real<lower=0> tau_org_sub;
-  real<lower=0> tau_ec4_sub;
-  real<lower=0> tau_enz_sub;
+  real tau_substrate;
+  real tau_ec4_sub;
+  real tau_enz_sub;
   vector<multiplier=tau_substrate>[N_substrate] a_substrate;
-  vector<multiplier=tau_org_sub>[N_org_sub] a_org_sub;
   vector<multiplier=tau_ec4_sub>[N_ec4_sub] a_ec4_sub;
@@
 transformed parameters {
-  vector[N_biology] log_km = mu
-    + a_substrate[substrate]
-    + a_ec4_sub[ec4_sub]
-    + a_org_sub[org_sub];
+  vector[N_biology] log_km = mu + a_substrate[substrate] + a_ec4_sub[ec4_sub];
   for (b in 1:N_biology){
@@
   nu ~ gamma(2, 0.1);
-  sigma ~ normal(0, 2);
-  mu ~ normal(-1, 2);
+  mu ~ normal(-2, 2);
+  sigma ~ lognormal(0, 0.2);
+  tau_substrate ~ lognormal(0, 0.3);
+  tau_ec4_sub ~ lognormal(0, 0.2);
+  tau_enz_sub ~ lognormal(-2, 0.2);
   a_substrate ~ normal(0, tau_substrate);
@@
   a_enz_sub ~ normal(0, tau_enz_sub);
-  a_org_sub ~ normal(0, tau_org_sub);
-  tau_org_sub ~ normal(0, 1);
-  tau_ec4_sub ~ normal(0, 1);
-  tau_enz_sub ~ normal(0, 1);
-  tau_substrate ~ normal(0, 1);
 }
@@
 \subsection{Model validation}
-\label{sec:org95e5ff9}
+\label{sec:orge1b4412}
 We used a number of standard methods to test our models' validity, aiming to
@@
 \subsubsection{Computation}
-\label{sec:orga5a8cb0}
+\label{sec:org0882078}
 For each MCMC run we used arviz \citeprocitem{27}{[27]} to
@@
 \subsubsection{Comparison models}
-\label{sec:org02fdbca}
+\label{sec:orga1407c9}
 Ideally we would compare the results of our models with previously published
@@
 \subsubsection{Graphical posterior predictive checks}
-\label{sec:org65e677b}
-Figure \ref{fig:org9d62348} below shows our main model's marginal posterior
+\label{sec:org9b775a6}
+Figure \ref{fig:orga252fa7} below shows our main model's marginal posterior
 predictive 1\%-99\% interval for each observation, alongside the observed
@@
 \includegraphics[width=.9\linewidth]{./results/plots/ppc_sabio-km-enz.png}
-\caption{\label{fig:org9d62348}Marginal posterior predictive distributions for main model fit to SABIO Km data}
+\caption{\label{fig:orga252fa7}Marginal posterior predictive distributions for main model fit to SABIO Km data}
 \end{figure}
 
-Figures \ref{fig:org9e0149b} and \ref{fig:org387aabb} show
+Figures \ref{fig:org976ad80} and \ref{fig:org5edb4ec} show
 graphical posterior predictive checks for the really simple and simple models
@@
 \includegraphics[width=.9\linewidth]{./results/plots/ppc_sabio-km-simple.png}
-\caption{\label{fig:org9e0149b}Marginal posterior predictive distributions for simple model fit to SABIO Km data}
+\caption{\label{fig:org976ad80}Marginal posterior predictive distributions for simple model fit to SABIO Km data}
 \end{figure}
@@
 \includegraphics[width=.9\linewidth]{./results/plots/ppc_sabio-km-really-simple.png}
-\caption{\label{fig:org387aabb}Marginal posterior predictive distributions for really simple model fit to SABIO Km data}
+\caption{\label{fig:org5edb4ec}Marginal posterior predictive distributions for really simple model fit to SABIO Km data}
 \end{figure}
 
-From the continuous line of black dots in figure \ref{fig:org9e0149b} we can
+From the continuous line of black dots in figure \ref{fig:org976ad80} we can
 see that the posterior predictive intervals have approximately the same order as
@@
 
-In contrast figure \ref{fig:org387aabb} shows that the really simple
+In contrast figure \ref{fig:org5edb4ec} shows that the really simple
 model is clearly underfitting the observed data - there is no relationship
@@
 
-\subsubsection{{\bfseries\sffamily TODO} Cross validation}
-\label{sec:org57d7028}
+Overall these graphical posterior predictive checks made us confident that our model was achieving a qualitatively satisfactory balance between overfitting and underfitting.
+However, from graphical posterior predictive checks it was not easy to distinguish our final model from the BLK model.
+Figure \ref{fig:org34bdfa4}
+shows the posterior predictive check graph for this model: there is no clear sign of overfitting or under-fitting.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.9\linewidth]{./results/plots/ppc_sabio-km-blk.png}
+\caption{\label{fig:org34bdfa4}Marginal posterior predictive distributions for BLK model fit to SABIO Km data}
+\end{figure}
+
+In order to evaluate the two models in more detail, we compared their out-of-sample predictive performance using a cross-validation approach.
+
+\subsubsection{Cross validati}
+\label{sec:org59da6d4}
 For a quantitative compliment to the graphical posterior predictive checks, we
-used cross-validation to assess our models' out-of-sample predictive
+used cross validation to assess our models' out-of-sample predictive
 performance. For exploratory comparison we calculated each model's approximate
@@
 \citeprocitem{29}{[29]} and implemented in
-\citeprocitem{27}{[27]}. To address cases where diagnostics
-suggested that the approximate leave-one-out algorithm was likely to be
-unreliable we also carried out exact tenfold cross-validation.
+\citeprocitem{27}{[27]}. Since the diagnostics suggested that the
+approximate leave-one-out algorithm was likely to be unreliable, we also performed
+exact five-fold cross validation.
 
@@
 
-\hl{RESULTS OF CROSS VALIDATION COMPARISON}
+\begin{table}[htbp]
+\caption{Average out-of-sample log likelihoods for the SABIO-RK dataset}
+\centering
+\begin{tabular}{lr}
+\hline
+Model & Average out-of-sample log likelihood\\
+\hline
+BLK & -2.744\\
+final & -2.693\\
+\hline
+\end{tabular}
+\end{table}
 
-\subsection{{\bfseries\sffamily DONE} Web app}
-\label{sec:orga4f63ea}
+The final model made slightly better out-of-sample predictions. The distribution of log likelihoods by posterior sample can be seen in figure \ref{fig:org758fc32} below.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.9\linewidth]{./results/plots/cv.png}
+\caption{\label{fig:org758fc32}Histograms of out-of-sample log likelihoods}
+\end{figure}
+
+This leave-one-fold-out cross validation procedure is somewhat flawed because,
+unlike with leave-one-observation-out cross validation or in a real-life
+prediction scenario, the test dataset contains many unseen parameter
+combinations. However, this problem is likely to favour the BLK model over the
+final model, as there are tend to be more unseen enzyme/substrate combinations
+than unseen organism/substrate combinations. This test therefore satisfied us
+that the final model was an improvement on the BLK model for modelling the
+SABIO-RK dataset.
+
+\subsection{Web app}
+\label{sec:org0963c7b}
 In order to make our results accessible to the systems biology community, we wrote a web app that presents them in an easy to use and actionable format. This can be found at \hl{URL}.
@@
 \section{Results}
-\label{sec:org7b65808}
+\label{sec:org073d2dc}
 \subsection{Marginal distributions of interesting parameters}
-\label{sec:org5a237e8}
+\label{sec:orgdb0154e}
 We begin by looking at the marginal distributions of the model's standard
 deviation parameters. Comparing these indicates which effects varied the
-most. The comparison in figure \ref{fig:org9d781a4} shows that the most important
+most. The comparison in figure \ref{fig:orgb52286b} shows that the most important
 hierarchical parameter is the substrate effect, followed by the ec4/substrate
@@
 \centering
-\includegraphics[width=.9\linewidth]{./results/plots/sd_posteriors.png}
-\caption{\label{fig:org9d781a4}Comparison of hierarchical standard deviation parameters}
+\includegraphics[width=.9\linewidth]{./results/plots/sd_posteriors_enz.png}
+\caption{\label{fig:orgb52286b}Comparison of hierarchical standard deviation parameters}
 \end{figure}
 
-Figure \ref{fig:org42ddc1d} compares the marginal distributions of km
+Figure \ref{fig:orgba68f81} compares the marginal distributions of km
 parameters from our best SABIO-RK and BRENDA models. The marginal posterior
@@
 \includegraphics[width=.9\linewidth]{./results/plots/log_km_comparison.png}
-\caption{\label{fig:org42ddc1d}Comparison of measured and modelled Km parameter distributions in BRENDA and SABIO-RK}
+\caption{\label{fig:orgba68f81}Comparison of measured and modelled Km parameter distributions in BRENDA and SABIO-RK}
 \end{figure}
@@
 \subsection{Comparison of estimated Kms with physiological metabolite concentrations}
-\label{sec:org4c1309c}
+\label{sec:org11eb774}
 Several sources suggest that the Km parameter for a particular organism,
@@
 the corresponding Km measurements and marginal posterior distributions from our
-model. Figure \ref{fig:org7c195f7} shows the results.
+model. Figure \ref{fig:org295e429} shows the results.
 
@@
 \includegraphics[width=.9\linewidth]{./results/plots/concentration_comparison.png}
-\caption{\label{fig:org7c195f7}Comparison of model predictions with physiological metabolite concentrations from HMDB}
+\caption{\label{fig:org295e429}Comparison of model predictions with physiological metabolite concentrations from HMDB}
 \end{figure}
@@
 other hierarchical standard deviation parameters \hl{CHECK THIS IS TRUE}, as can be
-seen in figure \ref{fig:org9d781a4}
+seen in figure \ref{fig:orgb52286b}
 
 \subsection{NADH vs NADPH}
-\label{sec:org1ec26e0}
+\label{sec:orge6cf3c7}
 
-Figure \ref{fig:orga4246f3} shows histograms of measured Km parameters from the SABIO-RK
+Figure \ref{fig:org9607c60} shows histograms of measured Km parameters from the SABIO-RK
 dataset for NADH and NADPH binding enzymes alongside histograms of our final
@@
 \includegraphics[width=.9\linewidth]{./results/plots/nadh.png}
-\caption{\label{fig:orga4246f3}Comparison of modelled and measured Km parameters for NADH and NADPH}
+\caption{\label{fig:org9607c60}Comparison of modelled and measured Km parameters for NADH and NADPH}
 \end{figure}
@@
 \subsection{{\bfseries\sffamily TODO} Cofactor vs substrate}
-\label{sec:org7111380}
+\label{sec:org67134da}
 We tested whether there tended to be a systematic difference between the km of
@@
 \subsection{Which results to use?}
-\label{sec:orgfb8accf}
+\label{sec:orga31f699}
 
@@
 \section{{\bfseries\sffamily TODO} Case studies}
-\label{sec:org093574d}
+\label{sec:org8f9d1e3}
 \section{{\bfseries\sffamily TODO} Discussion}
-\label{sec:org745f9fe}
+\label{sec:org7d06580}
 \section{References}
-\label{sec:orgf7cebe3}
+\label{sec:orgee11e8b}
 

### 2022-02-04 paper.tex
More analysis and paper update

@@
-% Created 2022-02-03 Thu 14:55
+% Created 2022-02-04 Fri 13:27
 % Intended LaTeX compiler: pdflatex
@@
 \section{Introduction}
-\label{sec:org9cc370f}
+\label{sec:org5ea4afe}
 The choice of appropriate parameters for biological models has been a recurring
@@
 \section{State of the art}
-\label{sec:org1ec53bd}
+\label{sec:org9c93615}
 The problem of modelling kinetic parameter data from online databases has
@@
 \section{Method}
-\label{sec:orgb68824a}
+\label{sec:org605976c}
 \subsection{Data fetching}
-\label{sec:orgcc0ecb4}
+\label{sec:org3d25f90}
 We fetched data from the SABIO-RK \citeprocitem{21}{[21]}
@@
 \subsection{Data processing}
-\label{sec:org2adc330}
+\label{sec:org8c373ca}
 We made several significant data processing choices. See the library file
@@
 \subsubsection{Filtering}
-\label{sec:org80f3330}
+\label{sec:org0465d20}
 For each dataset and kinetic parameter, we removed all reports which failed to
@@
 \subsubsection{Grouping}
-\label{sec:org41b73d8}
+\label{sec:org29a793a}
 Instead of modelling reports directly, we chose to group together reports with
@@
 \subsection{Statistical model}
-\label{sec:org5d80a30}
+\label{sec:orgb315466}
 We used a Bayesian hierarchical regression model to describe all data. Since Km
@@
 distribution degrees of freedom parameter \(\nu\) follows the recommendation in
-\citeprocitem{24}{[24]} and is truncated below at
-2, reflecting our view that the measurement distribution should not be
-excessively heavy-tailed.
+\citeprocitem{24}{[24]} and is truncated below at 2,
+reflecting our view that the measurement distribution should not be excessively
+heavy-tailed. The prior for the enzyme:substrate interaction effect is also put
+close to zero, reflecting our view that, though non-negligible, this effect is
+likely to be smaller than the others.
 
@@
 Prior distributions for parameters with log-normal priors are shown graphically
-in figure \ref{fig:org777eb93}.
+in figure \ref{fig:org304b949}.
 
@@
 \includegraphics[width=.9\linewidth]{./results/plots/sd_priors_enz.png}
-\caption{\label{fig:org777eb93}KDE plots of prior distributions for log-normal parameters}
+\caption{\label{fig:org304b949}KDE plots of prior distributions for log-normal parameters}
 \end{figure}
@@
 \subsection{Model validation}
-\label{sec:orge1b4412}
+\label{sec:orgbeeba56}
 We used a number of standard methods to test our models' validity, aiming to
@@
 \subsubsection{Computation}
-\label{sec:org0882078}
+\label{sec:org79f7617}
 For each MCMC run we used arviz \citeprocitem{27}{[27]} to
@@
 \subsubsection{Comparison models}
-\label{sec:orga1407c9}
+\label{sec:org8ee97f9}
 Ideally we would compare the results of our models with previously published
@@
 \subsubsection{Graphical posterior predictive checks}
-\label{sec:org9b775a6}
-Figure \ref{fig:orga252fa7} below shows our main model's marginal posterior
+\label{sec:org4a04c1f}
+Figure \ref{fig:org626fb1d} below shows our main model's marginal posterior
 predictive 1\%-99\% interval for each observation, alongside the observed
@@
 \includegraphics[width=.9\linewidth]{./results/plots/ppc_sabio-km-enz.png}
-\caption{\label{fig:orga252fa7}Marginal posterior predictive distributions for main model fit to SABIO Km data}
+\caption{\label{fig:org626fb1d}Marginal posterior predictive distributions for main model fit to SABIO Km data}
 \end{figure}
 
-Figures \ref{fig:org976ad80} and \ref{fig:org5edb4ec} show
+Figures \ref{fig:org570e7fb} and \ref{fig:orgbfddfe3} show
 graphical posterior predictive checks for the really simple and simple models
@@
 \includegraphics[width=.9\linewidth]{./results/plots/ppc_sabio-km-simple.png}
-\caption{\label{fig:org976ad80}Marginal posterior predictive distributions for simple model fit to SABIO Km data}
+\caption{\label{fig:org570e7fb}Marginal posterior predictive distributions for simple model fit to SABIO Km data}
 \end{figure}
@@
 \includegraphics[width=.9\linewidth]{./results/plots/ppc_sabio-km-really-simple.png}
-\caption{\label{fig:org5edb4ec}Marginal posterior predictive distributions for really simple model fit to SABIO Km data}
+\caption{\label{fig:orgbfddfe3}Marginal posterior predictive distributions for really simple model fit to SABIO Km data}
 \end{figure}
 
-From the continuous line of black dots in figure \ref{fig:org976ad80} we can
+From the continuous line of black dots in figure \ref{fig:org570e7fb} we can
 see that the posterior predictive intervals have approximately the same order as
@@
 
-In contrast figure \ref{fig:org5edb4ec} shows that the really simple
+In contrast figure \ref{fig:orgbfddfe3} shows that the really simple
 model is clearly underfitting the observed data - there is no relationship
@@
 However, from graphical posterior predictive checks it was not easy to distinguish our final model from the BLK model.
-Figure \ref{fig:org34bdfa4}
+Figure \ref{fig:orge1490d5}
 shows the posterior predictive check graph for this model: there is no clear sign of overfitting or under-fitting.
@@
 \includegraphics[width=.9\linewidth]{./results/plots/ppc_sabio-km-blk.png}
-\caption{\label{fig:org34bdfa4}Marginal posterior predictive distributions for BLK model fit to SABIO Km data}
+\caption{\label{fig:orge1490d5}Marginal posterior predictive distributions for BLK model fit to SABIO Km data}
 \end{figure}
@@
 \subsubsection{Cross validati}
-\label{sec:org59da6d4}
+\label{sec:org666e20c}
 For a quantitative compliment to the graphical posterior predictive checks, we
@@
 
-The final model made slightly better out-of-sample predictions. The distribution of log likelihoods by posterior sample can be seen in figure \ref{fig:org758fc32} below.
+The final model made slightly better out-of-sample predictions. The distribution of log likelihoods by posterior sample can be seen in figure \ref{fig:orgdbd4e2d} below.
 
@@
 \includegraphics[width=.9\linewidth]{./results/plots/cv.png}
-\caption{\label{fig:org758fc32}Histograms of out-of-sample log likelihoods}
+\caption{\label{fig:orgdbd4e2d}Histograms of out-of-sample log likelihoods}
 \end{figure}
@@
 \subsection{Web app}
-\label{sec:org0963c7b}
+\label{sec:org358097e}
 In order to make our results accessible to the systems biology community, we wrote a web app that presents them in an easy to use and actionable format. This can be found at \hl{URL}.
@@
 \section{Results}
-\label{sec:org073d2dc}
+\label{sec:org8931a47}
 \subsection{Marginal distributions of interesting parameters}
-\label{sec:orgdb0154e}
+\label{sec:orgc03ca64}
 We begin by looking at the marginal distributions of the model's standard
 deviation parameters. Comparing these indicates which effects varied the
-most. The comparison in figure \ref{fig:orgb52286b} shows that the most important
+most. The comparison in figure \ref{fig:orgc30fb62} shows that the most important
 hierarchical parameter is the substrate effect, followed by the ec4/substrate
@@
 \includegraphics[width=.9\linewidth]{./results/plots/sd_posteriors_enz.png}
-\caption{\label{fig:orgb52286b}Comparison of hierarchical standard deviation parameters}
+\caption{\label{fig:orgc30fb62}Comparison of hierarchical standard deviation parameters}
 \end{figure}
 
-Figure \ref{fig:orgba68f81} compares the marginal distributions of km
+Figure \ref{fig:org8ebf498} compares the marginal distributions of km
 parameters from our best SABIO-RK and BRENDA models. The marginal posterior
@@
 \includegraphics[width=.9\linewidth]{./results/plots/log_km_comparison.png}
-\caption{\label{fig:orgba68f81}Comparison of measured and modelled Km parameter distributions in BRENDA and SABIO-RK}
+\caption{\label{fig:org8ebf498}Comparison of measured and modelled Km parameter distributions in BRENDA and SABIO-RK}
 \end{figure}
@@
 \subsection{Comparison of estimated Kms with physiological metabolite concentrations}
-\label{sec:org11eb774}
+\label{sec:org5bbe315}
 Several sources suggest that the Km parameter for a particular organism,
@@
 the corresponding Km measurements and marginal posterior distributions from our
-model. Figure \ref{fig:org295e429} shows the results.
+model. Figure \ref{fig:org59254d8} shows the results.
 
@@
 \includegraphics[width=.9\linewidth]{./results/plots/concentration_comparison.png}
-\caption{\label{fig:org295e429}Comparison of model predictions with physiological metabolite concentrations from HMDB}
+\caption{\label{fig:org59254d8}Comparison of model predictions with physiological metabolite concentrations from HMDB}
 \end{figure}
@@
 other hierarchical standard deviation parameters \hl{CHECK THIS IS TRUE}, as can be
-seen in figure \ref{fig:orgb52286b}
+seen in figure \ref{fig:orgc30fb62}
 
 \subsection{NADH vs NADPH}
-\label{sec:orge6cf3c7}
+\label{sec:orgb9021ce}
 
-Figure \ref{fig:org9607c60} shows histograms of measured Km parameters from the SABIO-RK
+Figure \ref{fig:orgb28d979} shows histograms of measured Km parameters from the SABIO-RK
 dataset for NADH and NADPH binding enzymes alongside histograms of our final
@@
 \includegraphics[width=.9\linewidth]{./results/plots/nadh.png}
-\caption{\label{fig:org9607c60}Comparison of modelled and measured Km parameters for NADH and NADPH}
+\caption{\label{fig:orgb28d979}Comparison of modelled and measured Km parameters for NADH and NADPH}
 \end{figure}
 
-\subsection{{\bfseries\sffamily TODO} Cofactor vs substrate}
-\label{sec:org67134da}
+\subsection{Cofactor vs substrate}
+\label{sec:org26c3ec5}
 We tested whether there tended to be a systematic difference between the km of
@@
 
+Figure \ref{fig:org30a85d3} shows the 1\%-99\% posterior intervals for substrate
+effects of cofactors superimposed on those of all substrates. There does not
+seem to be a particular pattern
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.9\linewidth]{./results/plots/cofactor_effects.png}
+\caption{\label{fig:org30a85d3}Comparison of modelled and measured Km parameters for NADH and NADPH}
+\end{figure}
+
+Figure \ref{fig:org875fc6c} shows, for each reaction for which both a
+cofactor and substrates were measured in the SABIO-RK dataset, the posterior mean of the cofactor and that of the substrate. Again there is no particular pattern
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.9\linewidth]{./results/plots/cofactor_substrate_comparison.png}
+\caption{\label{fig:org875fc6c}Comparison of modelled and measured Km parameters for NADH and NADPH}
+\end{figure}
+
 \subsection{Which results to use?}
-\label{sec:orga31f699}
+\label{sec:orge9e6b7f}
 
@@
 \section{{\bfseries\sffamily TODO} Case studies}
-\label{sec:org8f9d1e3}
+\label{sec:org39443a4}
 \section{{\bfseries\sffamily TODO} Discussion}
-\label{sec:org7d06580}
+\label{sec:org987bc3d}
 \section{References}
-\label{sec:orgee11e8b}
+\label{sec:orgdd8d6ae}
 

### 2022-02-04 analyse.py

@@ -9,6 +9,8 @@ from xarray.core.dataset import Dataset
 
 from prepare_data import PREPARED_DIR
 from src.analysis import (
+    plot_cofactor_effects,
+    plot_cofactor_substrate_comparison,
     plot_concentration_comparison,
     plot_log_km_comparison,
     plot_nadh_comparison,
@@ -148,6 +150,15 @@ def main():
     # concentration comparison
     f = plot_concentration_comparison(posterior_enz, dfs["sabio_concs"])
     f.savefig(os.path.join("results", "plots", "concentration_comparison.png"))
+
+    # cofactors
+    f = plot_cofactor_effects(posterior_enz)
+    f.savefig(os.path.join("results", "plots", "cofactor_effects.png"))
+
+    f = plot_cofactor_substrate_comparison(posterior_enz, dfs["sabio"])
+    f.savefig(
+        os.path.join("results", "plots", "cofactor_substrate_comparison.png")
+    )
 
 
 if __name__ == "__main__":

### 2022-02-04 src/analysis.py

@@ -3,15 +3,18 @@
 from typing import Dict, List
 
 import arviz as az
+import matplotlib
 import numpy as np
 import pandas as pd
 from arviz.data.inference_data import InferenceData
 from arviz.plots.forestplot import plot_forest
+from matplotlib import patches
 from matplotlib import pyplot as plt
 from matplotlib.figure import Figure
+from numpy.core.multiarray import inner
 from xarray.core.dataset import Dataset
 
-from src.data_preparation import check_is_df
+from src.data_preparation import COFACTORS, check_is_df
 
 
 def plot_vars(posterior: Dataset, vars: List[str]) -> Figure:
@@ -258,5 +261,94 @@ def plot_oos_cv(idatas: Dict[str, InferenceData]) -> Figure:
         title="Comparison of out-of-sample log likelihoods",
         xlabel="log likelihood",
         ylabel="Relative frequency",
+    )
+    return f
+
+
+def plot_cofactor_effects(posterior):
+    a_sub_qs = (
+        az.extract_dataset(posterior, var_names="a_substrate", combined=True)
+        .quantile([0.01, 0.5, 0.99], dim="sample")
+        .to_dataframe()
+        .unstack("quantile")["a_substrate"]
+        .drop("unknown", axis=0)
+        .sort_values(0.5)
+    )
+    f, ax = plt.subplots(figsize=[12, 5])
+    x = pd.Series(
+        np.linspace(*ax.get_xlim(), num=len(a_sub_qs)), index=a_sub_qs.index
+    )
+    ax.vlines(
+        x.values,
+        a_sub_qs[0.01],
+        a_sub_qs[0.99],
+        linewidth=0.1,
+        label="Other substrate",
+    )
+    ax.vlines(
+        x.loc[COFACTORS].values,
+        a_sub_qs.loc[COFACTORS, 0.01],
+        a_sub_qs.loc[COFACTORS, 0.99],
+        color="red",
+        label="Cofactor",
+    )
+    for c in COFACTORS:
+        ax.annotate(c, [x.loc[c], a_sub_qs.loc[c, 0.01]], fontsize=7)
+    ax.set_xticks([])
+    ax.legend(frameon=False)
+    ax.set(
+        title="1%-99% posterior intervals for substrate effects.",
+        ylabel="Parameter values",
+    )
+    return f
+
+
+def plot_cofactor_substrate_comparison(posterior, lits):
+    log_kms = az.extract_dataset(posterior, var_names="log_km", combined=True)
+    means = (
[…]

### 2022-02-04 src/data_preparation.py

@@ -21,7 +21,7 @@ COFACTORS = [
     "NADH",
     "NAD+",
     "NADPH",
-    "acetyl-CoA",
+    "O2",
     "NADP+",
     "ADP",
 ]
@@ -442,7 +442,9 @@ def prepare_data_sabio_km(
     lits = (
         reports.loc[cond]
         .groupby(lit_cols, dropna=False)
-        .agg({"y": "median", "biology": "first"})
+        .agg(
+            {"y": "median", "biology": "first", "reaction_substrates": "first"}
+        )
         .reset_index()
         .loc[lambda df: df.groupby("organism")["y"].transform("size") > 50]
         .reset_index()

### 2022-02-07 paper.tex
Spell checking

--- 2022-02-03 paper.tex
@@
-% Created 2022-02-03 Thu 14:55
+% Created 2022-02-07 Mon 11:55
 % Intended LaTeX compiler: pdflatex
@@
 We fit a range of Bayesian regression models to analyse data from the BRENDA and
-SABIO-RK databases pertaining to Michalis constants or Km parameters. We report
+SABIO-RK databases pertaining to Michaelis constants or Km parameters. We report
 the results and provide tools for reproducing them independently and using them
@@
 \section{Introduction}
-\label{sec:org9cc370f}
+\label{sec:org38b72aa}
 The choice of appropriate parameters for biological models has been a recurring
@@
 strategies. Although the most popular approach is the estimation of parameters
-through optimization methods
+through optimisation methods
 \citeprocitem{4}{[4]}, \citeprocitem{5}{[5]},
@@
 By taking into the account the existing issues with model parameterisation and
-in an effort to address them, we have previously developed a standardized
-protocol for the definition of appropriate parameter lognormal distributions
+in an effort to address them, we have previously developed a standardised
+protocol for the definition of appropriate parameter log-normal distributions
 based on information retrieved from different sources and in accordance with the
@@
 \section{State of the art}
-\label{sec:org1ec53bd}
+\label{sec:orgd7d1f52}
 The problem of modelling kinetic parameter data from online databases has
@@
 \citeprocitem{19}{[19]} models logarithmic-scale measurements
-\(y_{ijk}\) of a Km value for substrate \(i\), ec number \(j\) and organism \(k\) using
+\(y_{ijk}\) of a Km value for substrate \(i\), EC number \(j\) and organism \(k\) using
 the following ANOVA-style regression model:
@@
 In this expression the term \(\mu\) represents substrate-specific effects, whereas
-the terms \(\alpha\) and \(\beta\) respectively represent substrate-ec number and
+the terms \(\alpha\) and \(\beta\) respectively represent substrate-EC number and
 substrate-organism interaction effects. \(\sigma\) represents the accuracy of the
@@
 \section{Method}
-\label{sec:orgb68824a}
+\label{sec:org81472b6}
 \subsection{Data fetching}
-\label{sec:orgcc0ecb4}
+\label{sec:org2d7f969}
 We fetched data from the SABIO-RK \citeprocitem{21}{[21]}
@@
 
-In total we fetched \hl{EXACT NUMBER HERE} raw BRENDA Km reports, \hl{EXACT NUMBER HERE} SABIO-RK Km reports and \hl{EXACT NUMBER HERE} Expasy EC numbers.
+In total we fetched \hl{EXACT NUMBER HERE} raw BRENDA Km reports, \hl\{EXACT
+NUMBER HERE\} SABIO-RK Km reports and \hl{EXACT NUMBER HERE} Expasy EC numbers.
 
 \subsection{Data processing}
-\label{sec:org2adc330}
+\label{sec:orgac5968f}
 We made several significant data processing choices. See the library file
@@
 \subsubsection{Filtering}
-\label{sec:org80f3330}
+\label{sec:org63b906c}
 For each dataset and kinetic parameter, we removed all reports which failed to
@@
 \item The organism must be have data from at least 50 separate study/biology
-(i.e. organism:substrate:ec4 for BRENDA or organism:substrate:enzyme for
+(i.e. organism:substrate:EC4 for BRENDA or organism:substrate:enzyme for
 SABIO-RK) combinations that satisfy all the other conditions.
@@
 \subsubsection{Grouping}
-\label{sec:org41b73d8}
+\label{sec:orgceac13d}
 Instead of modelling reports directly, we chose to group together reports with
@@
 \subsection{Statistical model}
-\label{sec:org5d80a30}
+\label{sec:org3fdde61}
 We used a Bayesian hierarchical regression model to describe all data. Since Km
@@
 parameters \(a^{enz:sub}\), \(a^{ec4:sub}\) and specific to interactions of
-substrate and enzyme and ec4 number respectively. In addition we used a
+substrate and enzyme and EC4 number respectively. In addition we used a
 student-T measurement model with latent standard deviation and degrees of
@@
 Prior distributions for parameters with log-normal priors are shown graphically
-in figure \ref{fig:org777eb93}.
+in figure \ref{fig:org81dd790}.
 
@@
 \includegraphics[width=.9\linewidth]{./results/plots/sd_priors_enz.png}
-\caption{\label{fig:org777eb93}KDE plots of prior distributions for log-normal parameters}
+\caption{\label{fig:org81dd790}KDE plots of prior distributions for log-normal parameters}
 \end{figure}
@@
 \subsection{Model validation}
-\label{sec:orge1b4412}
+\label{sec:orgb8bb985}
 We used a number of standard methods to test our models' validity, aiming to
@@
 \subsubsection{Computation}
-\label{sec:org0882078}
+\label{sec:org68e15d6}
 For each MCMC run we used arviz \citeprocitem{27}{[27]} to
@@
 \subsubsection{Comparison models}
-\label{sec:orga1407c9}
+\label{sec:orgc0f36f6}
 Ideally we would compare the results of our models with previously published
@@
 model, includes a single vector of hierarchical parameters at the finest
-available granularity, i.e. organism:substrate:ec4 for BRENDA data and
+available granularity, i.e. organism:substrate:EC4 for BRENDA data and
 organism:substrate:enzyme for SABIO-RK data. See supplemental information for
@@
 Our thinking was that the really simple model and simple models represent
-relative extremes of underfitting and overfitting. A well-specified model should
-be able to make better predictions than both, unless the data are very
+relative extremes of under-fitting and over-fitting. A well-specified model
+should be able to make better predictions than both, unless the data are very
 surprisingly uninformative.
@@
 \subsubsection{Graphical posterior predictive checks}
-\label{sec:org9b775a6}
-Figure \ref{fig:orga252fa7} below shows our main model's marginal posterior
+\label{sec:orgd9d5534}
+Figure \ref{fig:org9c97ea3} below shows our main model's marginal posterior
 predictive 1\%-99\% interval for each observation, alongside the observed
@@
 \includegraphics[width=.9\linewidth]{./results/plots/ppc_sabio-km-enz.png}
-\caption{\label{fig:orga252fa7}Marginal posterior predictive distributions for main model fit to SABIO Km data}
+\caption{\label{fig:org9c97ea3}Marginal posterior predictive distributions for main model fit to SABIO Km data}
 \end{figure}
 
-Figures \ref{fig:org976ad80} and \ref{fig:org5edb4ec} show
+Figures \ref{fig:orgb7e73ec} and \ref{fig:orgcd7bb6c} show
 graphical posterior predictive checks for the really simple and simple models
@@
 \includegraphics[width=.9\linewidth]{./results/plots/ppc_sabio-km-simple.png}
-\caption{\label{fig:org976ad80}Marginal posterior predictive distributions for simple model fit to SABIO Km data}
+\caption{\label{fig:orgb7e73ec}Marginal posterior predictive distributions for simple model fit to SABIO Km data}
 \end{figure}
@@
 \includegraphics[width=.9\linewidth]{./results/plots/ppc_sabio-km-really-simple.png}
-\caption{\label{fig:org5edb4ec}Marginal posterior predictive distributions for really simple model fit to SABIO Km data}
+\caption{\label{fig:orgcd7bb6c}Marginal posterior predictive distributions for really simple model fit to SABIO Km data}
 \end{figure}
 
-From the continuous line of black dots in figure \ref{fig:org976ad80} we can
+From the continuous line of black dots in figure \ref{fig:orgb7e73ec} we can
 see that the posterior predictive intervals have approximately the same order as
-the observed data, suggesting that the model is likely overfitting, or at least
+the observed data, suggesting that the model is likely over-fitting, or at least
 not using any structural information from the data.
 
-In contrast figure \ref{fig:org5edb4ec} shows that the really simple
-model is clearly underfitting the observed data - there is no relationship
+In contrast figure \ref{fig:orgcd7bb6c} shows that the really simple
+model is clearly under-fitting the observed data - there is no relationship
 between the order of the black dots and the blue lines.
 
-Overall these graphical posterior predictive checks made us confident that our model was achieving a qualitatively satisfactory balance between overfitting and underfitting.
-However, from graphical posterior predictive checks it was not easy to distinguish our final model from the BLK model.
-Figure \ref{fig:org34bdfa4}
-shows the posterior predictive check graph for this model: there is no clear sign of overfitting or under-fitting.
+Overall these graphical posterior predictive checks made us confident that our
+model was achieving a qualitatively satisfactory balance between over-fitting
+and under-fitting. However, from graphical posterior predictive checks it was not
+easy to distinguish our final model from the BLK model. Figure
+\ref{fig:orgd10c016} shows the posterior predictive check graph for this model:
+there is no clear sign of over-fitting or under-fitting.
 
@@
 \includegraphics[width=.9\linewidth]{./results/plots/ppc_sabio-km-blk.png}
-\caption{\label{fig:org34bdfa4}Marginal posterior predictive distributions for BLK model fit to SABIO Km data}
+\caption{\label{fig:orgd10c016}Marginal posterior predictive distributions for BLK model fit to SABIO Km data}
 \end{figure}
@@
 
-\subsubsection{Cross validati}
-\label{sec:org59da6d4}
+\subsubsection{Cross validation}
+\label{sec:org0929268}
 For a quantitative compliment to the graphical posterior predictive checks, we
@@
 
-The final model made slightly better out-of-sample predictions. The distribution of log likelihoods by posterior sample can be seen in figure \ref{fig:org758fc32} below.
+The final model made slightly better out-of-sample predictions. The distribution of log likelihoods by posterior sample can be seen in figure \ref{fig:orgbb23632} below.
 
@@
 \includegraphics[width=.9\linewidth]{./results/plots/cv.png}
-\caption{\label{fig:org758fc32}Histograms of out-of-sample log likelihoods}
+\caption{\label{fig:orgbb23632}Histograms of out-of-sample log likelihoods}
 \end{figure}
@@
 \subsection{Web app}
-\label{sec:org0963c7b}
+\label{sec:orga9bdb7f}
 In order to make our results accessible to the systems biology community, we wrote a web app that presents them in an easy to use and actionable format. This can be found at \hl{URL}.
@@
 \section{Results}
-\label{sec:org073d2dc}
+\label{sec:orgc2e91c0}
 \subsection{Marginal distributions of interesting parameters}
-\label{sec:orgdb0154e}
+\label{sec:orgbb5d0dd}
 We begin by looking at the marginal distributions of the model's standard
 deviation parameters. Comparing these indicates which effects varied the
-most. The comparison in figure \ref{fig:orgb52286b} shows that the most important
-hierarchical parameter is the substrate effect, followed by the ec4/substrate
+most. The comparison in figure \ref{fig:orgb7bd88b} shows that the most important
+hierarchical parameter is the substrate effect, followed by the EC4/substrate
 interaction, the enzyme/substrate interaction and finally the organism/substrate
@@
 distinguish all the effects from measurement noise. The substrate and
-ec4:substrate taus tend to be the greatest, suggesting that these are the most important predictors.
+EC4:substrate taus tend to be the greatest, suggesting that these are the most important predictors.
 
@@
 \includegraphics[width=.9\linewidth]{./results/plots/sd_posteriors_enz.png}
-\caption{\label{fig:orgb52286b}Comparison of hierarchical standard deviation parameters}
+\caption{\label{fig:orgb7bd88b}Comparison of hierarchical standard deviation parameters}
 \end{figure}
 
-Figure \ref{fig:orgba68f81} compares the marginal distributions of km
+Figure \ref{fig:orgdad634e} compares the marginal distributions of km
 parameters from our best SABIO-RK and BRENDA models. The marginal posterior
@@
 \includegraphics[width=.9\linewidth]{./results/plots/log_km_comparison.png}
-\caption{\label{fig:orgba68f81}Comparison of measured and modelled Km parameter distributions in BRENDA and SABIO-RK}
+\caption{\label{fig:orgdad634e}Comparison of measured and modelled Km parameter distributions in BRENDA and SABIO-RK}
 \end{figure}
@@
 \subsection{Comparison of estimated Kms with physiological metabolite concentrations}
-\label{sec:org11eb774}
+\label{sec:org2cc136e}
 Several sources suggest that the Km parameter for a particular organism,
@@
 the corresponding Km measurements and marginal posterior distributions from our
-model. Figure \ref{fig:org295e429} shows the results.
+model. Figure \ref{fig:org651e45f} shows the results.
 
@@
 \includegraphics[width=.9\linewidth]{./results/plots/concentration_comparison.png}
-\caption{\label{fig:org295e429}Comparison of model predictions with physiological metabolite concentrations from HMDB}
+\caption{\label{fig:org651e45f}Comparison of model predictions with physiological metabolite concentrations from HMDB}
 \end{figure}
@@
 hypothesis: the physiological concentrations are distributed similarly to the
-posterior Km samples.
-The graph on the right suggests that \(\ln\) scale Kms tend to be 1-2 \(\ln\) mM lower than their corresponding physiological concentration reports, give or take about 5 \(\ln\) mM.
-In other words, there is only a fairly weak relationship between a particular Km and its corresonding physiological concentration.
+posterior Km samples. The graph on the right suggests that \(\ln\) scale Kms tend
+to be 1-2 \(\ln\) mM lower than their corresponding physiological concentration
+reports, give or take about 5 \(\ln\) mM. In other words, there is only a fairly
+weak relationship between a particular Km and its corresponding physiological
+concentration.
 
@@
 other hierarchical standard deviation parameters \hl{CHECK THIS IS TRUE}, as can be
-seen in figure \ref{fig:orgb52286b}
+seen in figure \ref{fig:orgb7bd88b}
 
 \subsection{NADH vs NADPH}
-\label{sec:orge6cf3c7}
+\label{sec:org9fbc199}
 
-Figure \ref{fig:org9607c60} shows histograms of measured Km parameters from the SABIO-RK
+Figure \ref{fig:orgba3d2b1} shows histograms of measured Km parameters from the SABIO-RK
 dataset for NADH and NADPH binding enzymes alongside histograms of our final
@@
 \includegraphics[width=.9\linewidth]{./results/plots/nadh.png}
-\caption{\label{fig:org9607c60}Comparison of modelled and measured Km parameters for NADH and NADPH}
+\caption{\label{fig:orgba3d2b1}Comparison of modelled and measured Km parameters for NADH and NADPH}
 \end{figure}
 
-\subsection{{\bfseries\sffamily TODO} Cofactor vs substrate}
-\label{sec:org67134da}
+\subsection{Cofactor vs substrate}
+\label{sec:org212ae00}
 We tested whether there tended to be a systematic difference between the km of
@@
 
+Figure \ref{fig:orgec5d6be} shows the 1\%-99\% posterior intervals for substrate
+effects of cofactors superimposed on those of all substrates. There does not
+seem to be a particular pattern
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.9\linewidth]{./results/plots/cofactor_effects.png}
+\caption{\label{fig:orgec5d6be}Comparison of modelled and measured Km parameters for NADH and NADPH}
+\end{figure}
+
+Figure \ref{fig:org804e529} shows, for each reaction for which both a
+cofactor and substrates were measured in the SABIO-RK dataset, the posterior
+mean of the cofactor and that of the substrate. Again there is no particular
+pattern
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.9\linewidth]{./results/plots/cofactor_substrate_comparison.png}
+\caption{\label{fig:org804e529}Comparison of modelled and measured Km parameters for NADH and NADPH}
+\end{figure}
+
 \subsection{Which results to use?}
-\label{sec:orga31f699}
+\label{sec:orgabb80dc}
 
@@
 The SABIO-RK dataset provides more specific information than the BRENDA
-datase. UNIPROT ids are available for many Km measurements, whereas the most
+dataset. UNIPROT ids are available for many Km measurements, whereas the most
 specific enzyme-level information in the BRENDA dataset is the EC4 number. In
@@
 substrate, literature reference, enzyme, temperature, pH and strain type in
-interoperable formats and without the need for potentially error-prone string
+inter-operable formats and without the need for potentially error-prone string
 operations.
@@
 the exact enzyme they wish to model while still appropriately taking into
-account measurements of related enyzmes. In addition, the SABIO model excludes
+account measurements of related enzymes. In addition, the SABIO model excludes
 some potential sources of bias, such as reports from non-wildtype strains and
@@
 deciding which dataset to use in a given case. A further complication is that,
-due to the lack of interoperable identifiers in the BRENDA dataset, it is
-difficult to directly compare the which yields better predctions.
+due to the lack of inter-operable identifiers in the BRENDA dataset, it is
+difficult to directly compare the which yields better predictions.
 
-In our judgment the best approach given these circumstances is to favour the
+In our judgement the best approach given these circumstances is to favour the
 SABIO-RK results in cases where they provides a lot of information for a given
@@
 contains a lot of information about that parameter. In order to assess the
-liklihood of the BRENDA dataset being too coarse for a parameter whose enzyme
+likelihood of the BRENDA dataset being too coarse for a parameter whose enzyme
 shares an EC number, the marginal prior distributions of the common parameters
@@
 specific SABIO results. Finally, in order to assess the likelihood of bias in
-the BRENDA data, our webapp provides links to the source documents for any
+the BRENDA data, our web app provides links to the source documents for any
 parameter measurement.
@@
 \section{{\bfseries\sffamily TODO} Case studies}
-\label{sec:org8f9d1e3}
+\label{sec:org64cbb0a}
 \section{{\bfseries\sffamily TODO} Discussion}
-\label{sec:org7d06580}
+\label{sec:org1a5ed20}
 \section{References}
-\label{sec:orgee11e8b}
+\label{sec:org59ecc75}
 

### 2022-02-09 paper.tex
Add bibliography

@@
-% Created 2022-02-07 Mon 11:55
+% Created 2022-02-09 Wed 16:56
 % Intended LaTeX compiler: pdflatex
@@
 \section{Introduction}
-\label{sec:org38b72aa}
+\label{sec:org2887f91}
 The choice of appropriate parameters for biological models has been a recurring
@@
 \section{State of the art}
-\label{sec:orgd7d1f52}
+\label{sec:orga395b0e}
 The problem of modelling kinetic parameter data from online databases has
@@
 \section{Method}
-\label{sec:org81472b6}
+\label{sec:orga5dd2bb}
 \subsection{Data fetching}
-\label{sec:org2d7f969}
+\label{sec:orgb1e3bcf}
 We fetched data from the SABIO-RK \citeprocitem{21}{[21]}
@@
 \subsection{Data processing}
-\label{sec:orgac5968f}
+\label{sec:org5d73799}
 We made several significant data processing choices. See the library file
@@
 \subsubsection{Filtering}
-\label{sec:org63b906c}
+\label{sec:orgf83cf04}
 For each dataset and kinetic parameter, we removed all reports which failed to
@@
 \subsubsection{Grouping}
-\label{sec:orgceac13d}
+\label{sec:orgc5306b6}
 Instead of modelling reports directly, we chose to group together reports with
@@
 \subsection{Statistical model}
-\label{sec:org3fdde61}
+\label{sec:org74adbb3}
 We used a Bayesian hierarchical regression model to describe all data. Since Km
@@
 Prior distributions for parameters with log-normal priors are shown graphically
-in figure \ref{fig:org81dd790}.
+in figure \ref{fig:org021a284}.
 
@@
 \includegraphics[width=.9\linewidth]{./results/plots/sd_priors_enz.png}
-\caption{\label{fig:org81dd790}KDE plots of prior distributions for log-normal parameters}
+\caption{\label{fig:org021a284}KDE plots of prior distributions for log-normal parameters}
 \end{figure}
@@
 \subsection{Model validation}
-\label{sec:orgb8bb985}
+\label{sec:orgc0ff81d}
 We used a number of standard methods to test our models' validity, aiming to
@@
 \subsubsection{Computation}
-\label{sec:org68e15d6}
+\label{sec:org66e4d5a}
 For each MCMC run we used arviz \citeprocitem{27}{[27]} to
@@
 \subsubsection{Comparison models}
-\label{sec:orgc0f36f6}
+\label{sec:org2bacc28}
 Ideally we would compare the results of our models with previously published
@@
 \subsubsection{Graphical posterior predictive checks}
-\label{sec:orgd9d5534}
-Figure \ref{fig:org9c97ea3} below shows our main model's marginal posterior
+\label{sec:org40de5a1}
+Figure \ref{fig:org2dad7a6} below shows our main model's marginal posterior
 predictive 1\%-99\% interval for each observation, alongside the observed
@@
 \includegraphics[width=.9\linewidth]{./results/plots/ppc_sabio-km-enz.png}
-\caption{\label{fig:org9c97ea3}Marginal posterior predictive distributions for main model fit to SABIO Km data}
+\caption{\label{fig:org2dad7a6}Marginal posterior predictive distributions for main model fit to SABIO Km data}
 \end{figure}
 
-Figures \ref{fig:orgb7e73ec} and \ref{fig:orgcd7bb6c} show
+Figures \ref{fig:org6b01034} and \ref{fig:org3007bf8} show
 graphical posterior predictive checks for the really simple and simple models
@@
 \includegraphics[width=.9\linewidth]{./results/plots/ppc_sabio-km-simple.png}
-\caption{\label{fig:orgb7e73ec}Marginal posterior predictive distributions for simple model fit to SABIO Km data}
+\caption{\label{fig:org6b01034}Marginal posterior predictive distributions for simple model fit to SABIO Km data}
 \end{figure}
@@
 \includegraphics[width=.9\linewidth]{./results/plots/ppc_sabio-km-really-simple.png}
-\caption{\label{fig:orgcd7bb6c}Marginal posterior predictive distributions for really simple model fit to SABIO Km data}
+\caption{\label{fig:org3007bf8}Marginal posterior predictive distributions for really simple model fit to SABIO Km data}
 \end{figure}
 
-From the continuous line of black dots in figure \ref{fig:orgb7e73ec} we can
+From the continuous line of black dots in figure \ref{fig:org6b01034} we can
 see that the posterior predictive intervals have approximately the same order as
@@
 
-In contrast figure \ref{fig:orgcd7bb6c} shows that the really simple
+In contrast figure \ref{fig:org3007bf8} shows that the really simple
 model is clearly under-fitting the observed data - there is no relationship
@@
 easy to distinguish our final model from the BLK model. Figure
-\ref{fig:orgd10c016} shows the posterior predictive check graph for this model:
+\ref{fig:org72ba74d} shows the posterior predictive check graph for this model:
 there is no clear sign of over-fitting or under-fitting.
@@
 \includegraphics[width=.9\linewidth]{./results/plots/ppc_sabio-km-blk.png}
-\caption{\label{fig:orgd10c016}Marginal posterior predictive distributions for BLK model fit to SABIO Km data}
+\caption{\label{fig:org72ba74d}Marginal posterior predictive distributions for BLK model fit to SABIO Km data}
 \end{figure}
@@
 \subsubsection{Cross validation}
-\label{sec:org0929268}
+\label{sec:orgd639db8}
 For a quantitative compliment to the graphical posterior predictive checks, we
@@
 
-The final model made slightly better out-of-sample predictions. The distribution of log likelihoods by posterior sample can be seen in figure \ref{fig:orgbb23632} below.
+The final model made slightly better out-of-sample predictions. The distribution of log likelihoods by posterior sample can be seen in figure \ref{fig:orgb922b3e} below.
 
@@
 \includegraphics[width=.9\linewidth]{./results/plots/cv.png}
-\caption{\label{fig:orgbb23632}Histograms of out-of-sample log likelihoods}
+\caption{\label{fig:orgb922b3e}Histograms of out-of-sample log likelihoods}
 \end{figure}
@@
 \subsection{Web app}
-\label{sec:orga9bdb7f}
+\label{sec:org6636606}
 In order to make our results accessible to the systems biology community, we wrote a web app that presents them in an easy to use and actionable format. This can be found at \hl{URL}.
@@
 \section{Results}
-\label{sec:orgc2e91c0}
+\label{sec:org63c5ba9}
 \subsection{Marginal distributions of interesting parameters}
-\label{sec:orgbb5d0dd}
+\label{sec:orgdddb28e}
 We begin by looking at the marginal distributions of the model's standard
 deviation parameters. Comparing these indicates which effects varied the
-most. The comparison in figure \ref{fig:orgb7bd88b} shows that the most important
+most. The comparison in figure \ref{fig:org06f71e5} shows that the most important
 hierarchical parameter is the substrate effect, followed by the EC4/substrate
@@
 \includegraphics[width=.9\linewidth]{./results/plots/sd_posteriors_enz.png}
-\caption{\label{fig:orgb7bd88b}Comparison of hierarchical standard deviation parameters}
+\caption{\label{fig:org06f71e5}Comparison of hierarchical standard deviation parameters}
 \end{figure}
 
-Figure \ref{fig:orgdad634e} compares the marginal distributions of km
+Figure \ref{fig:org2d622a8} compares the marginal distributions of km
 parameters from our best SABIO-RK and BRENDA models. The marginal posterior
@@
 \includegraphics[width=.9\linewidth]{./results/plots/log_km_comparison.png}
-\caption{\label{fig:orgdad634e}Comparison of measured and modelled Km parameter distributions in BRENDA and SABIO-RK}
+\caption{\label{fig:org2d622a8}Comparison of measured and modelled Km parameter distributions in BRENDA and SABIO-RK}
 \end{figure}
@@
 \subsection{Comparison of estimated Kms with physiological metabolite concentrations}
-\label{sec:org2cc136e}
+\label{sec:orge8607e9}
 Several sources suggest that the Km parameter for a particular organism,
@@
 the corresponding Km measurements and marginal posterior distributions from our
-model. Figure \ref{fig:org651e45f} shows the results.
+model. Figure \ref{fig:orgd03c95e} shows the results.
 
@@
 \includegraphics[width=.9\linewidth]{./results/plots/concentration_comparison.png}
-\caption{\label{fig:org651e45f}Comparison of model predictions with physiological metabolite concentrations from HMDB}
+\caption{\label{fig:orgd03c95e}Comparison of model predictions with physiological metabolite concentrations from HMDB}
 \end{figure}
@@
 other hierarchical standard deviation parameters \hl{CHECK THIS IS TRUE}, as can be
-seen in figure \ref{fig:orgb7bd88b}
+seen in figure \ref{fig:org06f71e5}
 
 \subsection{NADH vs NADPH}
-\label{sec:org9fbc199}
+\label{sec:org1506f49}
 
-Figure \ref{fig:orgba3d2b1} shows histograms of measured Km parameters from the SABIO-RK
+Figure \ref{fig:org60fc650} shows histograms of measured Km parameters from the SABIO-RK
 dataset for NADH and NADPH binding enzymes alongside histograms of our final
@@
 \includegraphics[width=.9\linewidth]{./results/plots/nadh.png}
-\caption{\label{fig:orgba3d2b1}Comparison of modelled and measured Km parameters for NADH and NADPH}
+\caption{\label{fig:org60fc650}Comparison of modelled and measured Km parameters for NADH and NADPH}
 \end{figure}
@@
 \subsection{Cofactor vs substrate}
-\label{sec:org212ae00}
+\label{sec:orgbbe8b89}
 We tested whether there tended to be a systematic difference between the km of
@@
 
-Figure \ref{fig:orgec5d6be} shows the 1\%-99\% posterior intervals for substrate
+Figure \ref{fig:org50108f5} shows the 1\%-99\% posterior intervals for substrate
 effects of cofactors superimposed on those of all substrates. There does not
@@
 \includegraphics[width=.9\linewidth]{./results/plots/cofactor_effects.png}
-\caption{\label{fig:orgec5d6be}Comparison of modelled and measured Km parameters for NADH and NADPH}
+\caption{\label{fig:org50108f5}Comparison of modelled and measured Km parameters for NADH and NADPH}
 \end{figure}
 
-Figure \ref{fig:org804e529} shows, for each reaction for which both a
+Figure \ref{fig:org41eb8bc} shows, for each reaction for which both a
 cofactor and substrates were measured in the SABIO-RK dataset, the posterior
@@
 \includegraphics[width=.9\linewidth]{./results/plots/cofactor_substrate_comparison.png}
-\caption{\label{fig:org804e529}Comparison of modelled and measured Km parameters for NADH and NADPH}
+\caption{\label{fig:org41eb8bc}Comparison of modelled and measured Km parameters for NADH and NADPH}
 \end{figure}
@@
 \subsection{Which results to use?}
-\label{sec:orgabb80dc}
+\label{sec:org6b5ba09}
 
@@
 \section{{\bfseries\sffamily TODO} Case studies}
-\label{sec:org64cbb0a}
+\label{sec:org8b95ff0}
 \section{{\bfseries\sffamily TODO} Discussion}
-\label{sec:org1a5ed20}
+\label{sec:orgb9dae09}
 \section{References}
-\label{sec:org59ecc75}
+\label{sec:orgefab9d5}
 

### 2022-02-09 src/data_preparation.py
Update natural substrates and SABIO reports, fix small data prep bug

@@ -213,9 +213,9 @@ def prepare_data_brenda_km(
     reports["biology"] = (
         reports[biology_cols].fillna("").apply("|".join, axis=1)
     )
+    reports = reports.loc[cond].copy()
     lits = (
-        reports.loc[cond]
-        .groupby(lit_cols)
+        reports.groupby(lit_cols)
         .agg({"y": "median", "biology": "first"})
         .reset_index()
         .loc[lambda df: df.groupby("organism")["y"].transform("size") > 50]

### 2022-02-10 case_study/RunModel.m
Add files via upload

@@ -1,0 +1,225 @@
+tic;
+global PGAT3_g_k PGDH6_c_Vmax PGDH6_c_Keq PGDH6_c_Km6PG PGDH6_c_KmNADP PGDH6_c_KmRul5P PGDH6_c_KmNADPH PGDH6_g_Vmax PGDH6_g_Keq PGDH6_g_Km6PG PGDH6_g_KmNADP PGDH6_g_KmRul5P PGDH6_g_KmNADPH AK_c_k1 AK_c_k2 AK_g_k1 AK_g_k2
+global ALD_g_Vmax ALD_g_Keq ALD_g_KmFru16BP ALD_g_KiATP ALD_g_KiADP ALD_g_KiAMP ALD_g_KmGA3P ALD_g_KmDHAP ALD_g_KiGA3P ATPu_c_k ENO_c_Vmax ENO_c_Keq ENO_c_Km2PGA ENO_c_KmPEP G3PDH_g_Vmax G3PDH_g_Keq G3PDH_g_KmDHAP G3PDH_g_KmNADH G3PDH_g_KmGly3P G3PDH_g_KmNAD
+global G6PDH_c_Vmax G6PDH_c_Keq G6PDH_c_KmGlc6P G6PDH_c_KmNADP G6PDH_c_Km6PGL G6PDH_c_KmNADPH G6PDH_g_Vmax G6PDH_g_Keq G6PDH_g_KmGlc6P G6PDH_g_KmNADP G6PDH_g_Km6PGL G6PDH_g_KmNADPH G6PP_c_Vmax G6PP_c_Keq G6PP_c_KmGlc6P G6PP_c_KmGlc
+global GAPDH_g_Vmax GAPDH_g_Keq GAPDH_g_KmGA3P GAPDH_g_KmNAD GAPDH_g_Km13BPGA GAPDH_g_KmNADH GDA_g_k GK_g_Vmax GK_g_Keq GK_g_KmGly3P GK_g_KmADP GK_g_KmGly GK_g_KmATP GlcT_c_Vmax GlcT_c_KmGlc GlcT_c_alpha GPO_c_Vmax GPO_c_KmGly3P
[…]

### 2022-02-10 case_study/TrypMet.m

@@ -1,0 +1,262 @@
+function dydt = TrypMet(t, y)
+%Parameter values:
+global PGAT3_g_k PGDH6_c_Vmax PGDH6_c_Keq PGDH6_c_Km6PG PGDH6_c_KmNADP PGDH6_c_KmRul5P PGDH6_c_KmNADPH PGDH6_g_Vmax PGDH6_g_Keq PGDH6_g_Km6PG PGDH6_g_KmNADP PGDH6_g_KmRul5P PGDH6_g_KmNADPH AK_c_k1 AK_c_k2 AK_g_k1 AK_g_k2
+global ALD_g_Vmax ALD_g_Keq ALD_g_KmFru16BP ALD_g_KiATP ALD_g_KiADP ALD_g_KiAMP ALD_g_KmGA3P ALD_g_KmDHAP ALD_g_KiGA3P ATPu_c_k ENO_c_Vmax ENO_c_Keq ENO_c_Km2PGA ENO_c_KmPEP G3PDH_g_Vmax G3PDH_g_Keq G3PDH_g_KmDHAP G3PDH_g_KmNADH G3PDH_g_KmGly3P G3PDH_g_KmNAD
+global G6PDH_c_Vmax G6PDH_c_Keq G6PDH_c_KmGlc6P G6PDH_c_KmNADP G6PDH_c_Km6PGL G6PDH_c_KmNADPH G6PDH_g_Vmax G6PDH_g_Keq G6PDH_g_KmGlc6P G6PDH_g_KmNADP G6PDH_g_Km6PGL G6PDH_g_KmNADPH G6PP_c_Vmax G6PP_c_Keq G6PP_c_KmGlc6P G6PP_c_KmGlc
+global GAPDH_g_Vmax GAPDH_g_Keq GAPDH_g_KmGA3P GAPDH_g_KmNAD GAPDH_g_Km13BPGA GAPDH_g_KmNADH GDA_g_k GK_g_Vmax GK_g_Keq GK_g_KmGly3P GK_g_KmADP GK_g_KmGly GK_g_KmATP GlcT_c_Vmax GlcT_c_KmGlc GlcT_c_alpha k1 k2 GPO_c_Vmax GPO_c_KmGly3P
[…]

### 2022-02-10 case_study/RunModel.m
Update RunModel.m

@@ -211,7 +211,7 @@ figure (2)
     
  %% Plot everything
  
- for k=1:length(normal{j,3})
+ for k=1:length(normal{1,3})
  figure ()
     for j=1:countNormal
         plot(normal{j,2},normal{j,3}(:,k));
